### 2022-08-15 example_analysis/Daniel_LCA.Rmd
LCA Example Analysis

---
title: "LCA in R"
output: html_notebook
---

This is an example of a Latent Class Analysis (LCA).

In Part 1, I use the Lazarsfeld dataset and OpenMX.
In Part 2, I use the Drinking dataset and OpenMx.
In Part 3, I redo the analysis from Part 2 using tidySEM.

```{r}
library(tidySEM)
library(OpenMx)
```

Part 1: LCA using OpenMx and Lazarsfeld Dataset

Analysis based on: https://rdrr.io/cran/OpenMx/src/inst/models/passing/LCAlazarsfeld.R

```{r}
data("lazarsfeld", package ="OpenMx", verbose= TRUE)
data <- lazarsfeld

freq <- data[,5]
data[,1] <- as.ordered(data[,1])
data[,2] <- as.ordered(data[,2])
data[,3] <- as.ordered(data[,3])
data[,4] <- as.ordered(data[,4])
vars <- data[,1:4]

nclass <- 2
nvar <- 4
nthresh <- 1

nameList <- names(vars)

class1 <- mxModel("Class1", 
            mxMatrix("Iden", name = "R", nrow = nvar, ncol = nvar, free=FALSE),
            mxMatrix("Full", name = "M", nrow = 1, ncol = nvar, free=FALSE),
            mxMatrix("Full", name = "ThresholdsClass1", nrow = 1, ncol = nvar, 
            	dimnames = list("Threshold",nameList), free=TRUE),
		  mxExpectationNormal(covariance="R", means="M", dimnames=nameList, 
				      thresholds="ThresholdsClass1"),
		  mxFitFunctionML(vector=TRUE))

class2 <- mxModel("Class2", 
            mxMatrix("Iden", name = "R", nrow = nvar, ncol = nvar, free=FALSE),
            mxMatrix("Full", name = "M", nrow = 1, ncol = nvar, free=FALSE),
            mxMatrix("Full", name = "ThresholdsClass2", nrow = 1, ncol = nvar, 
            	dimnames = list("Threshold",nameList), free=TRUE),
		  mxExpectationNormal(covariance="R", means="M", dimnames=nameList,
				      thresholds="ThresholdsClass2"),
		  mxFitFunctionML(vector=TRUE))

# Define the model

lcamodel <- mxModel("lcamodel", class1, class2, mxData(vars, type="raw"),

# Create class membership probabilities, constrain them to be in the range 0-1, and make their sum equal 1.0
            mxMatrix("Full", name = "ClassMembershipProbabilities", nrow = nclass, ncol = 1, free=TRUE, 
			        labels = c(paste("pclass", 1:nclass, sep=""))),
            mxBounds(c(paste("pclass", 1:nclass, sep="")),0,1),
		    mxMatrix("Iden", nrow = 1, name = "constraintLHS"),
		    mxAlgebra(sum(ClassMembershipProbabilities), name = "constraintRHS"),
            mxConstraint(constraintLHS == constraintRHS),

# Define the objective function
            mxAlgebra(-2*sum(freq * log(pclass1%x%Class1.objective + pclass2%x%Class2.objective)), 
            	name="lca"),
            mxFitFunctionAlgebra("lca")
)

# Run the job
model <- mxRun(lcamodel, suppressWarnings=TRUE)
summary(model)
model$matrices

# Check results against those hard-coded from old Mx:
negativeThresholds <- as.vector(mxEval(Class1.ThresholdsClass1, model))
positiveThresholds <- as.vector(mxEval(Class2.ThresholdsClass2, model))
if (negativeThresholds[[1]] > 0) {
	temp <- negativeThresholds
	negativeThresholds <- positiveThresholds
	positiveThresholds <- temp
}

omxCheckCloseEnough(negativeThresholds, c(-1.3247, -0.2909, -0.1466, -0.0044),.01)
omxCheckCloseEnough(positiveThresholds, c(0.1805, 0.9071, 1.3169, 1.5869),.01)
omxCheckCloseEnough(sort(as.vector(mxEval(ClassMembershipProbabilities, model))), c(0.4440, 0.5560),.01)
omxCheckCloseEnough(model$output$Minus2LogLikelihood, 4696.444, 0.01)
```

Part 2: LCA Using OpenMx and Drinking Dataset,

Dataset from: https://stats.oarc.ucla.edu/sas/dae/latent-class-analysis/

```{r}
data <- read.csv('C:/Users/danie/OneDrive/Desktop/LCA/drinking_data.csv', header = T, sep = ',')

#check if two rows have the same response pattern
library(DAKS)
pattern(data)

dat1 <- c(1,0,0,0,0,0,0,0,0, 126)
dat2 <- c(0,0,0,0,0,0,0,0,0, 102)
dat3 <- c(1,0,0,0,0,0,0,0,1, 47)
dat4 <- c(1,1,0,0,0,0,0,0,0, 47)
dat5 <- c(1,0,0,0,0,1,0,0,0, 43)
dat6 <- c(126, 102, 47, 47, 43)

dat <- data.frame(dat1,dat2,dat3,dat4,dat5)
dat <- t(dat)
dat <- as.data.frame(dat)
names(dat) <- c("Q1", "Q2", "Q3", "Q4", "Q5", "Q6", "Q7", "Q8", "Q9", "freq")
rm("data", "dat1", "dat2", "dat3", "dat4", "dat5", "dat6")

dat[,1] <- as.ordered(dat[,1])
dat[,2] <- as.ordered(dat[,2])
dat[,3] <- as.ordered(dat[,3])
dat[,4] <- as.ordered(dat[,4])
dat[,5] <- as.ordered(dat[,5])
dat[,6] <- as.ordered(dat[,6])
dat[,7] <- as.ordered(dat[,7])
dat[,8] <- as.ordered(dat[,8])
dat[,9] <- as.ordered(dat[,9])

freq <- dat[,10]
vars <- dat[,1:9]

nclass <- 2
nvar <- 9
nthresh <- 1
nameList <- names(vars)

class1 <- mxModel("Class1", 
                  mxMatrix("Iden", name = "R", nrow = nvar, ncol = nvar, free=FALSE),
                  mxMatrix("Full", name = "M", nrow = 1, ncol = nvar, free=FALSE),
                  mxMatrix("Full", name = "ThresholdsClass1", nrow = 1, ncol = nvar, 
                           dimnames = list("Threshold",nameList), free=TRUE),
                  mxExpectationNormal(covariance="R", means="M", dimnames=nameList, 
                                      thresholds="ThresholdsClass1"),
                  mxFitFunctionML(vector=TRUE))

class2 <- mxModel("Class2", 
                  mxMatrix("Iden", name = "R", nrow = nvar, ncol = nvar, free=FALSE),
                  mxMatrix("Full", name = "M", nrow = 1, ncol = nvar, free=FALSE),
                  mxMatrix("Full", name = "ThresholdsClass2", nrow = 1, ncol = nvar, 
                           dimnames = list("Threshold",nameList), free=TRUE),
                  mxExpectationNormal(covariance="R", means="M", dimnames=nameList,
                                      thresholds="ThresholdsClass2"),
                  mxFitFunctionML(vector=TRUE))

# Define the model

lcamodel <- mxModel(model = "lcamodel", class1, class2, mxData(vars, type="raw"))

# Create class membership probabilities, constrain them to be in the range 0-1, and make their sum equal 1.0
mxMatrix("Full", name = "ClassMembershipProbabilities", nrow = nclass, ncol = 1, free=TRUE, labels = c(paste("pclass", 1:nclass, sep=""))) 
mxBounds(c(paste("pclass", 1:nclass, sep="")),0,1) 
mxMatrix("Iden", nrow = 1, name = "constraintLHS") 
mxAlgebra(sum(ClassMembershipProbabilities), name = "constraintRHS") 
mxConstraint(constraintLHS == constraintRHS)

# Define the objective function
mxAlgebra(-2*sum(freq * log(pclass1%x%Class1.objective + pclass2%x%Class2.objective)), name="lca")

mxFitFunctionAlgebra("lca")

# Run the job
model <- mxRun(lcamodel, suppressWarnings=TRUE)
summary(model)
model$matrices

# Check results against those hard-coded from old Mx:
negativeThresholds <- as.vector(mxEval(Class1.ThresholdsClass1, model))
positiveThresholds <- as.vector(mxEval(Class2.ThresholdsClass2, model))
if (negativeThresholds[[1]] > 0) {
  temp <- negativeThresholds
  negativeThresholds <- positiveThresholds
  positiveThresholds <- temp
}

```

Part 3: LCA using tidySEM

# Part 1: Using mclust [incomplete]

```{r}
library(mclust)
```

```{r}
# Fit 1-9 class model
model <- Mclust(data)
```