### 2022-10-28 manuscript.tex
edits from Mauricio

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  ,man,floatsintext]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{latent class analysis, mixture models, best practices, free open source software, tidySEM\newline\indent Word count: 8206}
\usepackage{csquotes}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Best Practices in Latent Class Analysis using Free Open Source Software},
  pdfauthor={Caspar J. van Lissa1,2 \& Mauricio Garnier-Villarreal3},
  pdflang={en-EN},
  pdfkeywords={latent class analysis, mixture models, best practices, free open source software, tidySEM},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Best Practices in Latent Class Analysis using Free Open Source Software}
\author{Caspar J. van Lissa\textsuperscript{1,2} \& Mauricio Garnier-Villarreal\textsuperscript{3}}
\date{}

\shorttitle{BEST PRACTICES LCA}

\authornote{

Correspondence concerning this article should be addressed to Caspar J. van Lissa, Padualaan 14, 3584CH Utrecht, The Netherlands. E-mail: \href{mailto:c.j.vanlissa@gmail.com}{\nolinkurl{c.j.vanlissa@gmail.com}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} Utrecht University, Methodology \& Statistics\\\textsuperscript{2} Open Science Community Utrecht\\\textsuperscript{3} Vrije Universiteit Amsterdam, Sociology}

\abstract{%
Latent class analysis is a family of techniques for identifying groups in data based
on a parametric model. Its examples include mixture models, latent class
analysis (with continuous and/or ordinal indicators), and latent class (mixture)
growth analysis. Despite the popularity of this technique, there is limited
guidance with respect to best practices in estimating and reporting mixture
models. Moreover, although user-friendly interfaces for advanced mixture
modeling have long been available in commercial software packages, open
source alternatives have remained somewhat inaccessible. This tutorial
describes best practices for the estimation and reporting of latent class analysis,
using free and open source software in R. To this end, this tutorial
introduces new functionality for estimating and reporting mixture
models in the \texttt{tidySEM} R-package whose backend relies on
the \texttt{OpenMx} R-package.
}

\begin{document}
\maketitle

Latent class analysis (LCA) is an umbrella term that refers to a number
of techniques for estimating unobserved group membership based on a
parametric model of one or more observed indicators of group membership.
The types of LCA have become quite popular across scientific fields,
most notably finite Gaussian mixture modeling and latent profile
analysis. Vermunt, J.K. et al. (2004) defined LCA more generally as
virtually any statistical model where ``some of the parameters {[}\ldots{]}
differ across unobserved subgroups''.

Despite its popularity, there is a lack of standards for estimating and
reporting LCA. While Van De Schoot, Sijbrandij, Winter, Depaoli, and Vermunt (2017) developed
reporting guidelines for a specific type of LCA known as latent growth
models, general reporting guidelines for all LCA are still lacking. This
complicates manuscript review and assessment of the quality of published
studies, and introduces a risk of misapplications of the technique. The
present paper seeks to address this gap in the literature by suggesting
updated guidelines for estimation and reporting of LCA, based on current
best practices. Importantly, in order to lower the barrier of entry and
ensure reproducibility of all examples, this paper exclusively relies on
free, open source software in R. Our goal is to make best-practices in
LCA widely accessible.

\hypertarget{defining-latent-class-analysis}{%
\subsection{Defining Latent Class Analysis}\label{defining-latent-class-analysis}}

Latent class analysis is a group of methods for estimating unobserved
groups based on a parametric model of observed indicators of group
membership. The concept of LCA can be understood in different ways. A
mixture model assumes that the study population is composed of \(K\)
subpopulations or classes. It further assumes that the observed data are
a mixture of data generated by class-specific models. The simplest
univariate model is a normal distribution, which can be described with
two parameters: the mean and the variance. Commonly, the same model is
estimated across all classes, but with different parameters for each
class (i.e., class-specific means and variances). Mixture modeling then
estimates both the parameters for each class, and the probability that
an individual belongs to each class. This way, describing the whole
sample with a single mean and variance is a special case where we assume
the sample is homogeneous in function of these parameters.

As an illustrative example, imagine that a detective wants to know if it
would be possible to use mixture modeling to identify the sex of a
suspect, based on footprints found at the crime scene. To test the
feasibility of this approach, the detective records the shoe sizes and
sex of 100 volunteers. The resulting observed data look like this:

\begin{figure}
\includegraphics[width=7in]{shoedens} \caption{Kernel density plot of shoe sizes.}\label{fig:shoedens}
\end{figure}

The distribution is evidently bimodal, which bodes well for the intended
mixture model. In this case, the number of classes is known a-priori.
When estimating a two-class mixture model, the detective observes that
the model estimates the mean shoe size of the two groups are equal to
9.22 and
7.25, which is close to the true
means of the two groups, namely 9.04
and 6.93. When tabulating estimated
group membership against observed (known) group membership, it can be
seen that women are classified with a high degree of accuracy, but men
are not:

\begin{table}

\caption{\label{tab:tabshoe}Observed group membership by estimated class membership.}
\centering
\begin{tabular}[t]{l|r|r}
\hline
Observed & Class 1 & Class 2\\
\hline
Man & 28 & 21\\
\hline
Woman & 0 & 51\\
\hline
\end{tabular}
\end{table}

LCA is like confirmatory factor analysis, except that the continuous
latent variable is substituted with a categorical latent variable. Both
assume that the indicators are related, and these relations can be
represented by the underlying latent variable. One difference between
the two techniques is that factor analysis can be considered as a way to
group observed \emph{variables} into latent constructs, and LCA groups
\emph{individuals} into classes.While factor analysis seeks to describe the
relations between variables, LCA seeks to describe each unobserved group
and the differences between them. In line with this distinction, LCA
(mixture models) is referred to as a ``person-centered'' technique, and
factor analysis as a ``variable-centered'' technique
(Masyn, 2013; Nylund-Gibson \& Choi, 2018).

When the focus is on the model parameters in each group, LCA can be
thought of as similar to a multi-group structural equation model. The
main distinction is that group membership is not known a-priori, but is
instead estimated -- with measurement error -- based on the data.
Whereas in a multi-group model, the data are split by group and treated
as independent samples, in LCA, all cases contribute to the estimation
of all parameters in all groups. The relative contribution of each case
to the parameters of each group is determined by that case's posterior
probability of belonging to that group.

In factor analysis, the latent variable is continuous, so the relation
between indicators is define by the communalities between them, and the
latent variable represents the error corrected approximation of the
sample. While in LCA, the relation between indicators is define by how
well they separate the classes, as a multinomial regression between
indicators and latent variable (Masyn, 2013)

When the focus is on each individual's estimated class membership, LCA
can be thought of as a type of clustering algorithm. In line with this
perspective, LCA is sometimes described as ``model-based clustering''
(Hennig, Meila, Murtagh, \& Rocci, 2015; Scrucca, Fop, Murphy, \& Raftery, 2016). Many clustering
algorithms apply some recursive splitting algorithm to the data. By
contrast ``model-based'' clustering refers to the fact that LCA estimates
cluster membership based on a parametric model. Specifically, the
posterior class probability that an individual belongs to a latent class
can be computed from the likelihood of that individual's observed data
under given the class-specific model. For these differences, LCA is a clustering method, that corrects for (random) measurement error, while clustering algorithms do not correct for. This is something LCA has in common with factor analysis, as both model an underlying latent variable, which accounts for the communalities between indicators and correcting for measurement error. 0

Finally, in the context of machine learning, LCA can be considered as an
\emph{unsupervised classification} problem (Figueiredo \& Jain, 2002).
The term \emph{unsupervised} refers to the fact that the outcome variable --
true class membership -- is not known, and the term \emph{classification}
refers to the fact that the algorithm is predicting a categorical
outcome -- class membership.

\hypertarget{a-taxonomy-of-latent-class-analysis-methods}{%
\subsection{A Taxonomy of Latent Class Analysis Methods}\label{a-taxonomy-of-latent-class-analysis-methods}}

In this paper, we use the term latent class analysis to refer to a
family of techniques that estimate latent class membership based on a
parametric model of observed indicators. From a historical perspective,
the term was initially conceived to refer to analyses with categorical,
usually binary indicators (Vermunt, J.K. et al., 2004). Nowadays, there are
a number of related techniques, known by distinct names, that serve a
similar purpose. The term ``latent class analysis'' seems most appropriate
as an umbrella term for this broader class of models, as it only refers
to the purpose of the analysis, and does not imply restrictions to the
model used, or the level of measurement of the indicators. Given the
abundance of terms in use for closely related classes of models, we will
provide a rudimentary taxonomy of LCA methods.

A common type of LCA is the \emph{finite Gaussian mixture model}; a
univariate analysis where the observed distribution of a single variable
is assumed to result from a mixture of a known number of normal
(Gaussian) distributions. The parameters of a finite Gaussian mixture
model are the means and variances of these underlying normal
distributions. The analysis of shoe sizes presented earlier is a
canonical example of this type of analysis. In the multivariate case,
with more than one indicator variable, the parameters of a mixture model
are the means, variances, and covariances between the indicators (which
can be standardized to obtain correlations). These parameters can be
estimated freely, or set to be constrained across classes (Collins \& Lanza, 2010).

The technique known as \emph{latent profile analysis (LPA)} is a special case
of the mixture model, which assumes conditional independence of the
indicators. Conditional independence means that, after class membership
is accounted for, the covariances/correlations between indicators are
assumed to be zero. This can be conceived of as a restricted mixture
model with covariances fixed to zero. Stating that the covariances/correlations are accounted by the latent variable. In some cases, such constraints
will be inappropriate; for instance when the cohesion between indicators
is expected to differ between classes (Collins \& Lanza, 2010). As an example, a LPA of ocean
plastic particles found two classes of particles based on length and
width: a class of smaller particles with a high correlation between
length and width, meaning that these particles were approximately round
or square in shape, and a class of larger particles with a low
correlation between length and width, meaning that these particles were
heterogenous in shape. From a theoretical perspective, this makes sense,
because the smaller particles have been ground down to a more uniform
shape by the elements.

It is also possible to estimate a mixture model based on latent
indicators. This means that, within each class, one or more continuous
latent variables are estimated based on the observed indicators.
Categorical latent variable membership is then estimated based on these
continuous latent variables. A common application of this approach is in
longitudinal research, where the indicators reflect one construct
assessed at different time points. Examples of this approach include
\emph{growth mixture models} (GMM) and \emph{latent class growth analyses} (LCGA).
These techniques estimate a latent growth model to describe grouped
trajectories over time. The growth mixture model is a latent class model
where the parameters that indicate class membership are the intercepts
and variances, and typically covariances of the latent growth variables,
e.g., a latent intercept and slope. This technique assumes that
individuals within a class can have heterogenous trajectories. If the
variance of the growth parameters is fixed to zero, it is known as a
latent class growth analysis. This latter approach assumes that all
individuals within a class share the same identical trajectory, and that
any variance in the indicators not explained by the class-specific
latent trajectories is due to residual error variance.

The term latent class analysis originally referred to cases where the
observed indicators were categorical (Collins \& Lanza, 2010).
Nowadays, it is more commonly used
as an umbrella term. To prevent ambiguity, the special case where
indicators are of binary or ordinal measurement level might be described
as \emph{latent class analysis with ordinal indicators}. Latent class models
with ordinal indicators are parameterized differently from mixture
models. One common parameterization assumes that each categorical
variable reflects an underlying standard normal distribution. The
parameters are ``thresholds'' that correspond to quantiles of a standard
normal distribution (with \(N(\mu = 0, \sigma = 1)\)). These thresholds
are estimated based on the proportion of individuals in each of the
response categories of the indicator variable. For example, a binary
indicator has a single threshold that distinguishes the two response
categories. If responses are distributed 50/50, then the corresponding
threshold would be \(t_1 = 0.00\). If the responses are distributed 60/40,
then the resulting threshold would be \(t_1 = 0.25\).

This paper will primarily focus on mixture models and special cases
thereof, although most of the suggested guidelines are applicable to all
LCA methods.

\hypertarget{use-cases-for-latent-class-analysis}{%
\subsection{Use Cases for Latent Class Analysis}\label{use-cases-for-latent-class-analysis}}

There are several use cases for which LCA methods are suitable. One
example is to test a theory that postulates the existence of a
categorical latent variable. For example, \emph{identity status theory}
posits that, at any given point in time, adolescents reside in one of
four identity statuses. Latent class analysis can be used to identify
these four statuses based on observed indicators (e.g., self-reported
identity exploration and commitment). If results indicate that the data
are better described by a different number of classes, or that the
four-class solution does not correspond to the predicted pattern of
responses on the indicators, then the theory may be called into
question.

Another use case is unsupervised learning: when the goal is to restore
unobserved class membership based on observed indicators, or to classify
individuals. For example, a mixture model can be used as a diagnostic
aid when several clinical indicators can be used to distinguish between
a fixed number of physical (Baughman, Bisgard, Lynn, \& Meade, 2006), mental
(Wu, Woody, Yang, Pan, \& Blazer, 2011) health problems, or health outcomes after recovering from cancer (Hardy, Garnier-Villarreal, McCarthy, Anderson, \& Reynolds, 2018). The example of shoe size is a
rudimentary illustration of this type of application.

LCA can be used as a descriptive analysis where a researcher wishes to
describe their sample and identify a few prototypes based on many
variables. As an example, if a survey among all Dutch academics was
carried out with the goal of bringing about a funding reform, LCA could
be used to discover the types of publications that get funding.

With LCA, our goal could be to inductively identify the number of
classes. For instance, if we believe that a variable represents a group,
but don't know how many groups there are, LCA may be an appropriate
technique to answer this question. For example, Hopfer, Tan, and Wylie (2014)
studied substance use, sexual behavior, and mental health status of
urban population in Winnipeg, Canada. The underlying assumption was that
there were different risk profiles, but their number was not known. From
a collection of indicators, LCA provided evidence that there are four
distinct risk profiles in the Winnipeg area.

Another application of LCA is to classify individuals. In a peer
harassment study, Giang and Graham (2008) used LCA to classify over 2,000
sixth grade students into aggressor and victim latent classes. The
five-class solution comprised of classes of victims, aggressors, and
socially adjusted students. For instance, it revealed that there were
two types of victims: highly-victimized aggressive-victims and
highly-aggressive aggressive-victims.

LCA is also appropriate when we wish to identify indicators that capture
classes well. High quality indicators will be strongly related to the
latent variable and will lead to good class separation. This
relationship of high quality indicators to the latent variable is
reflected in very high or very low conditional response probabilities.
For a simulation study exploring the effects of indicator quality on
LCA, see Geiser and Wurpts (2014). Therefore, one could use conditional response
probabilities for each item to assess its quality with regards to how
well it helps separate the latent classes. From this, a theory about the
selection of indicators could be informed.

An extension of LCA is that containing covariates which can be used to
predict class membership. In this approach, we not only model the latent
class variable based on indicators, but we also relate the class
membership to other explanatory variables (Vermunt (2017)). An
example of using covariates comes from Nozadi et al. (2016) who
applied LCA to identify the probability of children's membership to an
anxiety class. The authors tested several covariates including
children's age, sex, and accuracy scores. Age and sex were not found to
be related to the children's latent class membership, hence these
covariates were excluded from the analysis. Accuracy scores, however,
were related to probabilities of being in anxiety and attention-anxiety
classes. Therefore this covariate was kept as a valuable predictor.

When our interest is the prediction of one or more outcomes, LCA can be
used to construct latent classes as categorical predictors.
Lanza, Tan, and Bray (2013) demonstrated how LCA can be used to classify
adolescents into depression classes, and subsequently these classes can
be used to predict smoking, grades, and delinquency. The study showed
that the outcomes predicted by class membership can be binary (regular
smoking), continuous (grades) or count (delinquency).

These mixture models can also be useful to evaluate quality of measurement across instrument. For example, employment status was collected across register and survey data, and there is disagreement between them. You could think that register data is without error, but it has been shown that is victim of measurement error as well. In this case mixture models was used to evaluate the underlying factor of ``error corrected employment status'', by using two indicators of the same question and identify random and systamtic measurement error in either indicator (Pavlopoulos \& Vermunt, 2015).

In addition to these applications, LCA can be used for dimensionality
reduction as the resulting groups summarize response patterns on a large
number of indicators. For example, MacGregor et al. (2021) investigated
symptom profiles among injured U.S. military personnel. They used
fifteen dichotomous items from the Post-Deployment Health Assessment
survey as LCA indicators. Combinatorics informs us that fifteen
dichotomous items have \(2^{15}\) or \(32,768\) unique symptom combinations.
Perhaps for this reason, MacGregor et al. (2021) incorporated LCA as a
method of dimensionality reduction. A five class solution was found to
have the best fit according to both statistical criteria and clinical
interpretability.

Finally, LCA can be used to deal with data which violate certain
assumptions. As discussed in the shoe size example, LCA can deal with
violations of normality. In fact, LCA assumes the population
distribution is a non-normal mixture of \(K\) normal distributions, and it
can discover the value of \(K\), i.e.~generate a \(K\)-class solution.

LCA and factor analysis share the approximation of an unobserved latent variable, the difference being if that variable is continuous of categorical. We encourage researchers to be agnostic about the \textbf{true} nature of the latent variable. And focus on the approach that fits the research question.

\hypertarget{best-practices}{%
\section{Best Practices}\label{best-practices}}

\hypertarget{best-practices-in-estimation}{%
\subsection{Best Practices in Estimation}\label{best-practices-in-estimation}}

The best practices in estimation, as outlined in Table
\ref{tab:checkest}, are rooted in existing recommendations for best
practices for estimating specific subtypes of LCA, including latent
class growth analysis (Van De Schoot et al., 2017) and latent
class analysis with ordinal indicators (e.g., Nylund-Gibson \& Choi, 2018).
These were generalized to be relevant to all types of LCA, and updated
to current best practices, as explained below.

\begin{tabular}[t]{l|l}
\hline
\# & Item\\
\hline
1. & Examining Observed Data\\
\hline
2. & Handling Missing Data\\
\hline
3. & Alternative Model Specifications\\
\hline
4. & Software\\
\hline
5. & Algorithm\\
\hline
6. & Class Enumeration\\
\hline
7. & Model Fit Indices\\
\hline
8. & Classification Diagnostics\\
\hline
9. & Interpreting Class Solutions\\
\hline
10. & Label switching\\
\hline
\end{tabular}

\hypertarget{examining-observed-data}{%
\subsubsection{Examining Observed Data}\label{examining-observed-data}}

Examining observed data is essential for any analysis as it may reveal
patterns and violations of assumptions that had not been considered
prior to data collection. Special attention should be paid to level of
measurement of the indicators. Finite Gaussian mixture models (including
LPA) are only suitable for continuous variables. Indicators with an
ordinal level of measurement are likely to violate the assumption of
within-class normal distributions of mixture models (see Vermunt, 2011). Personal experience consulting on LCA methods
and moderating the \texttt{tidyLPA} Google group suggest that the
misapplication of mixture models to ordinal (e.g., Likert-type)
indicators is the most common source of user error. Whereas it has been
argued that some parametric methods are robust when scales with 7+
indicators are treated as continuous (e.g., Norman, 2010) or
with 5+ categories (Rhemtulla, Brosseau-Liard, \& Savalei, 2012), this
certainly does not imply that all methods are. It is certainly unlikely
that such ordinal variables can be treated as a \emph{mixture} of multiple
normal distributions. The problem becomes egregious when the number of
classes estimated equals or exceeds the number of categories; in this
case, each class-specific mean could describe a single response
category, and a class-specific variance component would be nonsensical.
In sum, Likert-type scales are rarely suitable for Gaussian mixture modeling;
latent class analysis with ordinal indicators is more appropriate.

Relatedly, a recent publication claimed that an assumption of mixture
models is that observed indicators are normally distributed
(Spurk, Hirschi, Wang, Valero, \& Kauffeld, 2020). This is incorrect. When the number of classes is
greater than one, Gaussian mixture models assume that the observed indicators are
a mixture of multiple (multivariate) normal distributions. In our shoe
size example, it can be seen that the population distribution is
comprised of two normal distributions. When examined visually, the
population distribution is evidently bimodal. The Shapiro-Wilk normality
test (\(W = 0.971\), \(p < 0.05\) ) rejects the null hypothesis that the
sample comes from a normally distributed population. Yet, this is a
prototypical example of a mixed population distribution where LCA can
discover latent groups. If the population distribution were instead
normal, there would be no classes to extract as the whole population
would belong to a single (homogeneous) class.

Extensive descriptive statistics (including the number of unique values,
variance of categorical variables, and missingness; see next subsection)
can be obtained using the function \texttt{tidySEM::descriptives(data)}. Note,
however, that sample-level descriptive statistics are of limited value
when the goal of a study is to identify subsamples using latent class
analysis. Plots (density plots for continuous variables, and bar charts
for categorical ones) may be more diagnostic. Note that density plots
can also aid in the choice of the number of classes, as further
explained in the section on visualization. Descriptive statistics and
plots can be relegated to online supplements, provided that these are
readily accessible (consider using a GitHub repository as a
comprehensive public research archive, as explained in Van Lissa et al., 2021).

\hypertarget{handling-missing-data}{%
\subsubsection{Handling Missing Data}\label{handling-missing-data}}

Previous work has emphasized the importance of examining the pattern of
missing data and reporting how missingness was handled
(Van De Schoot et al., 2017). Three types of missingness have
been distinguished in the literature (Enders, 2010; Rubin, 1976): Missing
completely at random (MCAR), which means that missingness is random and unrelated to any observed or unobserved variables;
missing at random (MAR), which means that missingness is contingent on
the \emph{observed} data (and can thus be accounted for); and finally missing
not at random (MNAR), which means that missingness is related to
unobserved factors. It is possible to conduct a so-called ``MCAR'' test,
for example the non-parametric MCAR test (Jamshidian \& Jalal, 2010). But
note that the name ``MCAR test'' is somewhat misleading, as the
null-hypothesis of this test is that the data are not MAR, and a
significant test statistic indicates that missingness is related to the
observed data (MAR). A non-significant test statistic does not
distinguish between MCAR or MNAR. As Little's classic MCAR test relies
on the comparison of variances across groups with different patterns of
missing data, it assumes normality (Little, 1988). This assumption
is tenuous in the context of LCA. A non-parametric MCAR test, as
provided by Jamshidian and Jalal, may be more suitable
(Jamshidian \& Jalal, 2010). Unfortunately, this test was removed from the
central R-repository CRAN due to lack of maintenance. For this tutorial,
I have re-implemented it in the \texttt{mice} package as \texttt{mice::mcar()}, with a
fast backend in C++ and new printing and plotting methods.

While we concur that investigating missingness is due diligence, it is
important to emphasize that missingness is adequately handled by default
in many software packages for LCA, such as Mplus, Latet GOLD, and \texttt{OpenMx} which is
the backend of \texttt{tidySEM}. These packages use Full Information Maximum
Likelihood (FIML) estimation, which makes use of all available
information without imputing missing values. FIML is a best-practice
solution for handling missing data; on par with multiple imputation when the estimation model includes the same variables as the imputation model
(Lee \& Shi, 2021). FIML estimation assumes that missingness is
either MCAR or MAR. Thus, one would typically proceed with FIML,
regardless of the outcome of an MCAR test. Although FIML does not, by
default, handle missingness in exogenous variables -- all indicator
variables in LCA are endogenous, so this is not a concern.

Multiple imputation is less suitable to LCA for two reasons. First,
because LCA methods are often computationally expensive, and conducting
them on multiple imputed datasets may be unfeasible. Second, because
there is no straightforward way to integrate LCA results across multiple
datasets. To conclude; our recommendation is to inspect missingness
(e.g., using \texttt{mice::MCAR()}) and report the proportion of missingness
per variable (e.g., using \texttt{tidySEM::descriptives()}), before proceeding
with FIML. One minor concern is that the K-means algorithm, which
\texttt{tidySEM} uses for determining starting values, is \emph{not} robust to
missing values. When it fails, \texttt{tidySEM} automatically switches to
hierarchical clustering, unless the user specifies a different
clustering algorithm or uses manual starting values.

\hypertarget{alternative-model-specifications}{%
\subsubsection{Alternative Model Specifications}\label{alternative-model-specifications}}

In order to aid researchers working with latent trajectory models,
Van De Schoot et al. (2017) developed a protocol called
Guidelines for Reporting on Latent Trajectory Studies (GRoLTS). Two of
the GRoLTS guidelines (namely, 6a and 6b) refer to considering
alternative model specifications. Both are discussing specific cases in
latent trajectory model specification. The first is about whether the
variance of the growth parameter is estimated freely or fixed, and the
second is about whether conditional independence is assumed, as the
researcher might also want to free-up the variance-covariance structure.
In LCA, there are also many different ways to specify the model. Means,
variances, and covariances between the indicators can either be
constrained across classes, or estimated freely. Researchers using LCA
should transparently report their chosen parametrization as well as
discuss different parametrizations that were tested in the process.

Different types of latent class models have different parameters. For
example, mixture models and latent profile analyses typically have
class-specific means, variances, and covariances. Latent growth analyses
have the same parameters, but with respect to the latent growth
variables. Latent class analyses with ordinal indicators have
thresholds. All of these parameters can be freely estimated, constrained
to be equal across classes, or fixed to a certain value (e.g., to zero).
The total number of parameters thus scales with the number of estimated
classes. Consequently, LCA methods have a potentially very high number
of parameters. As any of these parameters could be misspecified, it is
important to consider alternative model specifications. However,
alternative model specifications may be approached differently depending
on whether an analysis is data driven (exploratory), or theory driven
(confirmatory). This distinction has remained underemphasized in prior
writing.

Prior literature on LCA has emphasized its exploratory applications
(Nylund, Asparouhov, \& Muthén, 2007). In exploratory LCA, a large number of models
are typically estimated in batch, with varying numbers of classes and
model specifications. The ``correct'' model specification is then
determined based on a combination of fit indices, significance tests,
and interpretability. For latent profile analysis, the function
\texttt{tidySEM::mx\_profiles(classes,\ variances,\ covariances)} largely
automates this process. The argument \texttt{classes} indicates which class
solutions should be estimated (e.g., 1 through 6). The argument
\texttt{variances} specifies whether variances should be \texttt{"equal"} or
\texttt{"varying"} across classes. The argument \texttt{covariances} specifies whether
covariances should be constrained to \texttt{"zero"}, \texttt{"equal"} or \texttt{"varying"}
across classes. The means are free to vary across classes by default,
although the more general function \texttt{tidySEM::mx\_mixture()} could be used
to circumvent this. After all models have been estimated, the function
\texttt{tidySEM::table\_fit()} can be used to obtain a model fit table suitable
for determining the optimal model according to best practices. Note,
however, that this table does not include the bootstrapped likelihood
ratio test (BLRT) by default, because this test is very computationally
expensive. It is recommended to use the function \texttt{tidySEM::BLRT()} to
compare a shortlist of likely candidate models based on other fit
indices. More on fit indices can be found in the Model Fit Indices
subsection of this paper.

Confirmatory LCA typically requires less comprehensive alternative model
specifications. For example, in the context of a preregistered analysis,
the main models of interest would have been specified a priori. Even in
this case, the theoretical model could be compared to a few others to
contextualize it.

\hypertarget{software}{%
\subsubsection{Software}\label{software}}

Many software packages are available for the estimation of LCA models.
Some of these packages have limited functionality, or implement specific
innovations. Other packages implement LCA in the context of a more
flexible structural equation modeling framework. The most notable
examples of the latter are the commercial programs Mplus and Latent
GOLD, and the free open source R-package OpenMx. The commercial packages
stand out because they offer relatively user-friendly interfaces and
implement sensible defaults for complex analyses, including LCA methods.
This lowers the threshold for applied researchers to adopt such methods.
Commercial software also has several downsides, however. One such
downside is that use of the software is restricted to those individuals
and institutions who can afford a license. Another downside is that the
source code, being proprietary, cannot be audited, debugged, or enhanced
by third parties. This incurs the risk that mistakes in the source code
may go unnoticed, and curbs progress as software developers cannot add
new functionality.

Conversely, the free open source program OpenMx (Neale et al., 2016) is very flexible, but
not very user-friendly. We directly address this limitation using the
\texttt{tidySEM} R-package. New functionality in \texttt{tidySEM} seeks to lower the
threshold for latent class analysis using \texttt{OpenMx}. It adheres to best
practices in estimation and reporting, as described in this paper. The
user interface is simple, making use of the model syntax of the widely
used \texttt{lavaan} R-package (Rosseel, 2012). This syntax offers a human-readable way to
specify latent variable models. Minor enhancements are made to simplify
the specification of LCA.

Because of the limitations of the aforementioned tools, we set out to
develop a free tool that provides sensible defaults and is easy to use,
but provides the option to access and modify all of the model inputs
(i.e., low barrier, high ceiling). \texttt{tidySEM} interfaces with existing
tools, and is able to translate between what existing tools are capable
of and what researchers and analysts carrying-out person-oriented
analyses would like to specify. Furthermore,\texttt{tidySEM} facilitates
fully-reproducible analyses and contributes to open science.

\hypertarget{algorithm}{%
\subsubsection{Algorithm}\label{algorithm}}

LCA parameters and model fit statistics can be estimated in a variety of
ways. The choice of the estimator depends on the presence of missing
values, sample size, number of indicators, and available computational
resources (Weller, Bowen, \& Faubert, 2020). A commonly used technique is maximum
likelihood (ML) estimation with the expectation-maximization (EM)
algorithm as a local optimizer. Imagine we are estimating two
parameters, e.g.~the class-specific means \(\mu_c\) on a continuous
indicator (ignoring the variance for now). The EM algorithm will attempt
to find a combination of values for these two parameters that maximizes
the likelihood (\(LL\)) of all observed data. In practice, instead of
maximizing \(LL\), often \(-2*LL\) is minimized, as this offers
computational advantages. We can think of this optimization problem as a
three-dimensional landscape: The X and Y dimensions are determined by
the class-specific means, so \(X = \mu_1\) and \(Y = \mu_2\) - and the
Z-dimension is determined by \(Z = -2*LL\). The optimizer must find the
deepest ``valley'' in this landscape, which reflects the combination of
\(\mu_1\) and \(\mu_2\) that maximizes the likelihood of the data. The EM
optimizer behaves somewhat like a marble dropped in this landscape. It
is dropped at some random point in space, and will roll into the nearest
valley. The problem is that, once EM rolls into a valley, it will settle
on the bottom of that valley (this is known as ``convergence''). It cannot
climb out again. Thus, if their are multiple valleys, the risk is that
the optimizer gets stuck in a shallower valley (a ``local optimum''), and
never discovers the deepest valley (the ``global optimum'', or best
solution). One solution to this problem is to drop many marbles at
random places, compare their final \(-2*LL\) values, choose the solution
with the lowest \(-2*LL\), and make sure that several marbles replicated
this solution. This is the ``random starts'' approach.

One problem with the random starts approach is that it is
computationally expensive to run this many replications. Moreover,
because the algorithm begins with random starting values, many of the
marbles are likely to be very far away from a ``good enough'' solution.
Two innovations may improve the estimation procedure. The first is that,
instead of picking random starting values, a ``reasonable solution'' may
be used for the starting values. For example, if we assume that the
different classes are likely to have different mean values on the
indicators, then the K-means clustering algorithm can be used to
determine these cluster centroids. We can compute the expected values of
all model parameters by treating the K-means solution as a known class
solution, and use these as starting values for a mixture model. One
remaining concern is that this approach may result in starting values
close to a local optimum, and that the EM algorithm will thus never find
the global optimum. A second innovation addresses this concern. Instead
of using EM, it is possible to use an optimizer that can climb out of a
valley. Simulated annealing iteratively considers some ``destination'' in
the landscape, and compares its likelihood to the current one. If the
destination likelihood is higher, the estimator moves there. If the
destination likelihood is \emph{lower}, the estimator still moves there
occasionally, based on probability. This latter property allows it to
escape local optima, and find the global optimum.

By default, \texttt{tidySEM} derives starting values using K-means clustering,
and identifies the global optimum solution using simulated annealing.
Once a solution has been found, simulated annealing is followed up with
a short run of the EM algorithm, as EM inherently produces an asymptotic
covariance matrix for the parameters that can be used to compute
standard errors. Note that these defaults can be manually overridden.

A recent paper suggested maximum likelihood with robust standard errors
should be used when the observed indicators are not normally distributed
(Spurk et al., 2020). This statement is incorrect, and may lead readers
to believe that they must use commercial software, as robust maximum
likelihood is currently only implemented in Mplus and LatentGOLD. As
explained before, mixture modeling assumes that observed data are a
mixture of (multivariate) normal distributions; thus, the observed
indicators will likely not be normally distributed.

\hypertarget{class-enumeration}{%
\subsubsection{Class Enumeration}\label{class-enumeration}}

As explained, LCA can be done in an exploratory or in a confirmatory
fashion. In exploratory LCA, a sequence of models is fitted to the data
with each additional model estimating one more class than the previous
model. These models are then compared and the best solution is selected
as the final class solution. In some cases, prior theory can inform the
researcher about the number of classes to expect. Even in such
confirmatory LCA cases, it is nonetheless useful to know if the
theoretical model is markedly better than those with differing numbers
of classes. Therefore, it may always be useful to compare different
class solutions.

From a sequence of models, the final class solution is chosen based on
both theoretical and statistical criteria. Theory should drive the
selection of indicator variables, inform the expectations and reflect on
the findings. In addition to this, there are several statistical
criteria to consider in model selection. These include but are not
limited to likelihood ratio tests, information criteria, and the Bayes
factor (Masyn, 2013; Weller et al., 2020).

Relative model fit can be examined using the likelihood ratio test. This
is only appropriate when the two models we wish to compare are nested.
The likelihood ratio test statistic is computed as the difference in
maximum log-likelihoods of the two models, with the test degrees of
freedom being the difference in the degrees of freedom of the two
compared models. The test statistic follows the \(\chi^2\) distribution,
and we want it to be non-significant in order to give support to the
simpler model. The likelihood ratio test can only compare two nested
models at a time (Lanza, Bray, \& Collins, 2003).

\hypertarget{model-fit-indices}{%
\subsubsection{Model Fit Indices}\label{model-fit-indices}}

Fit indices typically used for determining the optimal number of classes
include the Akaike Information Criterion (AIC) and the Bayesian
Information Criterion (BIC). Both information criteria are based on the
-2*log-likelihood (which is lower for better fitting models), and add a
penalty for the number of parameters (thus incentivizing simpler
models). This helps balance model fit and model complexity. The lower
the value of an information criterion, the better the overall fit of the
model.

The general objective of the Information Criteria (IC) is to evaluate the model's out-of-sample predictive accuracy, so adjusting for over fitting. Fit measures like \(R^2\) evaluate the in-sample predictive accuracy, meaning that they evaluate the models ability to predict the observed outcomes based on the same data that was use to build up the model. These metrics are positively bias, meaning that will present better model fit than it has in reality. While IC corrects for this positive bias by evaluating the models accuracy approximating the out-of-sample predictive accuracy, meaning that it is the ability to predict the outcome for observations that are not part of the training model. Ideally we would estimate the model a lot of time with less observations and predict their scores, but IC's approximate this by their different penalty metrics. For example \(R^2\) will increase even if an added predictors are unnecessary, while IC's will show worst fit when a predictor (complexity) is unnecessary (McElreath, 2020)

The BIC applies a stronger penalty for model complexity that
scales logarithmically with the sample size. The literature suggests the
BIC may be the most appropriate information criterion to use for model
comparison (Masyn, 2013; Nylund-Gibson \& Choi, 2018). Both the AIC and the BIC are
available in the \texttt{tidySEM} output.

Information criteria may occasionally contradict each other, so it is
important to identify a suitable strategy to reconcile them. One option
is to select a specific fit index before analyzing the data. Another
option is to always prefer the most parsimonious model that has best fit
according to any of the available fit indices. Yet another option is to
incorporate information from multiple fit indices using the analytic
hierarchy process (Akogul \& Erisoglu, 2016). Finally, one might make an
elbow plot and compare multiple information criteria
(Nylund-Gibson \& Choi, 2018).

Another common test of model fit is the likelihood ratio \(\chi^2\)
goodness-of-fit test. However, this test is not implemented in
\texttt{tidySEM}.

A disadvantage of the IC's is that we do not have standard errors for them, so we only have the absolute values without a measure of their variability. So, the difference between models IC can be very small. Still indicating that the model with the lower value is ``better'', but if this difference is very small can considered them ``functionally equal'', and you should take into consideration the interpretability of the model.

LCA studies commonly report -2*log likelihood of the final class
solution. This is a basic fit measure used to compute most information
criteria. However, since log likelihood is not penalized for model
complexity, it will continuously fall with the addition of more classes.

An alternative is using the bootstrapped likelihood ratio test which can
be run using \texttt{tidySEM::BLRT()}. Currently, this test is computationally
expensive and can be slow on most computers. A faster version of this
test, namely an implementation of \emph{the lazy bootstrap}
(Kollenburg, Mulder, \& Vermunt, 2018) to \texttt{tidySEM} is being developed.

\hypertarget{classification-diagnostics}{%
\subsubsection{Classification Diagnostics}\label{classification-diagnostics}}

Best models will divide the sample into subgroups which are internally
homogeneous and externally distinct. Classification diagnostics give us
a way to assess the degree to which this is the case. They are separate
from model fit indices as a model can fit the data well but show poor
latent class separation (Masyn, 2013). Classification diagnostics
should not be used for model selection, but they can be used to
disqualify certain solutions because they are uninterpretable.
Interpretability should always be a consideration when considering
different class solutions (Nylund-Gibson \& Choi, 2018).

Three important classification diagnostics provided by \texttt{tidySEM} are:
(1) the \emph{minimum} and \emph{maximum} percentage of the sample assigned to a
particular \emph{class}, (2) the \emph{range of the posterior class probabilities} by most likely class membership, and (3) \emph{entropy}. All three are based
on posterior class probabilities.

The posterior class probability is a measure of classification
uncertainty which can be computed for each individual, or averaged for
each latent class. When the posterior class probability is computed for
each individual in the dataset, it represents each person's probability
of belonging to each latent class. For each person, the highest
posterior class probability is then determined and the individual is
assigned to the corresponding class. We want each individual's posterior
class probabilities to be high for one and low for the remaining latent
classes. This is considered a high classification accuracy and means
that the classes are distinct. To obtain posterior class probabilities,
run \texttt{tidySEM::class\_prob()}. This function produces output comprised of
several elements:

\texttt{\$sum.posterior} is a summary table of the posterior class probabilities
indicating what proportion of the data contributes to each class.

\texttt{\$sum.mostlikely} is a summary table of the most likely class membership
based on the highest posterior class probability. From this table, we
compute the minimum and maximum percentage of the sample assigned to a
particular class, , i.e.~\textbf{n\_min} (the smallest class proportion based
on the posterior class probabilities) and \textbf{n\_max} (the largest class
proportion based on the posterior class probabilities). We are
especially interested in \textbf{n\_min} as if it is very small and comprised
of few observations, the model for that group might not be locally
identified. It may be impossible to calculate descriptive statistics for
such a small class. Estimating LCA parameters on small subsamples might
lead to bias in the results. Therefore, we advise caution when dealing
with small classes.

\texttt{\$mostlikely.class} is a table with rows representing the class the
person was assigned to, and the columns indicating the average posterior
probability. The diagonal represents the probability that observations
in each class will be correctly classified. If any of the values on the
diagonal of this table is low, we might consider not to interpret that
solution. In \texttt{tidySEM} we use the diagonal to compute the range of the
posterior class probabilities by most likely class membership which
consists of the lowest class posterior probability (\textbf{prob\_min}), and
the highest posterior probability (\textbf{prob\_max}). Both \textbf{prob\_min} and
\textbf{prob\_max} can be used to disqualify certain class solutions, and are
a convenient way to summarize class separation in LCA. We want both
\textbf{prob\_min} and \textbf{prob\_max} to be high as that means that for all
classes the people who were assigned to that class have a high
probability of being there. \textbf{prob\_min} is especially important as it
can diagnose if there is a class with low posterior probabilities which
could make one reconsider that class solution.

\texttt{\$avg.mostlikely} contains the average posterior probabilities for each
class, for the subset of observations with most likely class of 1:k,
where k is the number of classes.

\texttt{\$individual} is the individual posterior probability matrix, with
dimensions n (number of cases in the data) x k (number of classes). Additionally it includes the \texttt{predicted} class in function of the highest predicted probability.
Individual class probabilities and/or predicted class are often useful for researchers who wish to do follow up analyses.

Entropy is a summary measure of posterior class probabilities across
classes and individuals. It ranges from 0 (model classification no
better than random chance) to 1 (perfect classification). As a rule of
thumb, values above .80 are deemed acceptable and those approaching 1
are considered ideal. An appropriate use of entropy is that it can
disqualify certain solutions if class separability is too low. Entropy
was not built for nor should it be used for model selection during class
enumeration (Masyn, 2013).

\textbf{n\_min}, \textbf{n\_max}, \textbf{prob\_min}, \textbf{prob\_max}, and \textbf{entropy} and
can be obtained using \texttt{tidySEM::table\_fit()}.

\hypertarget{interpreting-class-solutions}{%
\subsubsection{Interpreting Class Solutions}\label{interpreting-class-solutions}}

An important outcome of LCA are conditional item probabilities, also
known as class-specific item probabilities (Masyn, 2013),
conditional response, or conditional solution probabilities
(Geiser, 2012). They indicate the probability of an item being
endorsed given that the observation belongs to a particular latent
class. Conditional item probabilities can be obtained using
\texttt{tidySEM::table\_prob()}. If a particular item is endorsed by two or more
classes at markedly different rates, it is said to discriminate well
between the classes and is consequently considered a good indicator.
Classes are considered highly homogeneous with respect to an item when
for a particular item there is a distinct difference in conditional item
probabilities for two or more classes. For instance, if an item is
endorsed below 30\% for one class and above 70\% for another class, the
classes have high homogeneity with respect to this item
(Masyn, 2013). Conditional item probabilities are the analogue of
mean and standard deviation when the indicators are binary or ordinal.

A problem which can occur is that of inadmissible solutions. With binary
indicators, LCA is modelling a cross-table with all the predictors. The
problem with such cross-tables is that they will often contain empty
cells, i.e.~combinations of responses that never occur together. This
problem is reflected by extreme conditional item probabilities (as in
exactly 0 or 1). Such boundary parameter estimates could indicate that
the solution is invalid (Geiser, 2012). Boundary parameter
estimates can also happen with continuous indicators. For instance, if
we have a zero-inflated normal distribution and a two class solution,
one class might have the mean of zero and its standard deviation cannot
be determined since there is little variance. This too could be a sign
of an invalid solution, warn that too many classes were extracted, or
indicate a local optimum (Geiser, 2012).

\hypertarget{label-switching}{%
\subsubsection{Label Switching}\label{label-switching}}

The final class solution will usually discover and enumerate several
classes. The class ordering, however, is completely arbitrary. The class
labeled as Class 1 in one solution may become Class 2 or Class 3 in
another model, even when the only difference between the models is in
their starting values. Label switching is something to be mindful of
when comparing different LCA models (Masyn, 2013).

The order of clusters is nondeterministic when using K-means in
\texttt{tidySEM}. Therefore label switching is still a consideration. A simple
solution to this is setting a random seed number one line prior to
fitting the model. We advise \texttt{tidySEM} users to always do so in order to
circumvent label switching.

Class names should be chosen to accurately reflect group membership, as theoretically relevant names according to the interpretation of the class solutions.
Overly simplified and generalized class names may prove misleading to
both audiences and researches alike leading to what is known as a naming
fallacy (Weller et al., 2020).

\hypertarget{best-practices-in-reporting}{%
\subsection{Best Practices in Reporting}\label{best-practices-in-reporting}}

Among studies using LCA, reporting practices vary significantly
(Weller et al., 2020). Various authors have tried to improve and
standardize ways of reporting LCA (e.g. Masyn, 2013; Weller et al., 2020), but more work is needed. Van Lissa et al. (2020)
developed WORCS, a workflow for open reproducible code in science. WORCS
consists of step-by-step guidelines for research projects based on the
TOP-guidelines developed by Nosek et al. (2015). WORCS workflow can be
easily implemented in R in form of an R package which facilitates
preregistration, article drafting, version control, citation and
formatting, among others (Van Lissa et al., 2020).

TOP-guidelines emphasize the use of comprehensive citation (including
referencing the software used in the analysis), as well as code and data
sharing wherever possible (Nosek et al., 2015). Van Lissa et al. (2020)
suggest sharing synthetic data in case the original data cannot be
shared, and provide functions to generate such synthetic data. Ideally,
the entire research project is made reproducible so that others may
download it and reproduce it with just one click; for guidance, see
Van Lissa et al. (2020).

As the open science movement is gaining momentum, researchers are
becoming increasingly aware how important it is that analyses can be
reproduced and audited. In line with open science principles, one of the
suggested reporting standards relates to reproducible code. In this
context, it is important to note that user-friendly methods for
estimating LCA models have predominantly been available in commercial
software packages (e.g., \emph{Mplus} and \emph{Latent GOLD}). A potential
downside of commercial software is that it restricts the ability to
reproduce analyses to license holders, and prevents auditing research
because the underlying source code is proprietary. To overcome these
limitations, the present paper introduces new user-friendly functions in
the \texttt{tidySEM} R-package that can be used to estimate a wide range of LCA
models using the free, open-source R-package \texttt{OpenMx} as the backend. The reporting
guidelines described in this paper are adopted in \texttt{tidySEM} by default.
The \texttt{tidySEM} R-package thus makes advanced mixture modeling based on
best practices widely accessible, and facilitates the adoption of the
estimation and reporting guidelines described in this paper.

\hypertarget{visualization}{%
\subsubsection{Visualization}\label{visualization}}

Plots can greatly improve the interpretability of LCA models. There are
several stages where this is an important consideration.

First, in class enumeration, we want to compare several competing class
solutions. This can be done by means of the AIC and BIC. Here, we
suggest using an elbow plot with the \(K-class\ solution\) on the X-axis,
and information criteria value on the Y-axis. We show an implementation
of such an elbow plot in our Tutorial subsection.

Second, once we have decided on the final class solution, we want to
interpret the response patterns on the indicators for each latent class.
While \texttt{tidySEM} can create a table showing the probability of each
item's response endorsement (using \texttt{tidySEM::table\_prob}), it may be
easier to inspect these probabilities visually. For this reason, we
created \texttt{ggplot2::plot\_prob()}. The resulting shows response patterns on
all indicators for each group. We give an example of this plot in the
Tutorial subsection.

\hypertarget{tutorial}{%
\section{Tutorial}\label{tutorial}}

This is an example of an exploratory LCA using \texttt{tidySEM}.

\hypertarget{loading-the-dataset}{%
\subsection{Loading the Dataset}\label{loading-the-dataset}}

In this example, we use \emph{data\_mix\_ordinal}, a simulated data for mixture
modelling with ordinal indicators. This dataset is built-in to
\texttt{tidySEM}. For more information about the dataset, type
\texttt{?tidySEM::data\_mix\_ordinal} into the R console after loading the
\texttt{tidySEM} using library().

We load the dataset and convert the indicators into ordered factors.

\hypertarget{exploring-the-data}{%
\subsection{Exploring the Data}\label{exploring-the-data}}

An important step to preceding any statistical analysis is data
exploration. Here we use \texttt{tidySEM::descriptives()} to describe the
dataset we are using.

As we can see, the output includes various descriptives of our dataset.
Special attention should be paid to examining the pattern of
missingness, as discussed in the \emph{Handling missing data} section of this
paper. In our example, we see that there are no missing values, hence we
proceed with our analysis.

\hypertarget{conducting-latent-class-analysis}{%
\subsection{Conducting Latent Class Analysis}\label{conducting-latent-class-analysis}}

Before we fit a series of LCA models, we set a random seed using
\texttt{set.seed()}. The seed number can be any digit. This is an important
step as there is some inherent randomness in the LCA computations, and
having the same seed number ensures that two separate researcher obtain
exactly the same results when fitting LCA models.

Finally, we reach the step of fitting LCA models. To do so, we use
\texttt{tidySEM::mx\_lca()} which takes data and number of classes as inputs. Notice that the analysis will be done on all the variable in the data set, so you have to make sure it only includes the relevant variables. In
our example, we want to fit 1 to 4 class solutions and compare their
output. Depending on your computer's computational power, this might
take a while.

Notice that the model was run 10 times for each \(K\) class solution. This because the functions runs a total of 10 \textbf{extra} times the model. This way we can choose the best fitting model across all the solutions. In this examples we see that all models converged without issues.

As \texttt{tidySEM::mx\_lca()} was run for several models at the same time, the resulting object \texttt{res} is a \texttt{mixture\_list} class where each \(K\) class solution is an object in the list, and each of them are \texttt{OpenMx} models (\texttt{MxModel}).

\hypertarget{class-enumeration-1}{%
\subsection{Class Enumeration}\label{class-enumeration-1}}

In class enumeration, we want to compare a sequence of LCA models fitted
to the data. To aid the process, we create a model fit table using
\texttt{tidySEM::table\_fit()} with the results object as the input. As the
output contains a lot of information on each of the four fitted models,
we select a subset of helpful model fit indices and classification
diagnostics.

Our selection of fit indices and classification diagnostics includes:

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.1325}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.8675}}@{}}
\caption{\label{tab:unnamed-chunk-7}Selection of Fit Indices and Classification Diagnostics}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Selection
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Description
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Selection
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Description
\end{minipage} \\
\midrule()
\endhead
Name & the \(K\)-class solution \\
LL & the -2*log-likelihood of each model \\
Parameters & Number of estimated parameters in each model \\
AIC & the Akaike Information Criterion value \\
BIC & the Bayesian Information Criterion value \\
prob\_min & the lowest posterior class probability by most likely class membership \\
prob\_max & the highest posterior class probability by most likely class membership \\
n\_min & the lowest class proportion based on the posterior class probabilities \\
n\_max & the highest class proportion based on the posterior class probabilities \\
\bottomrule()
\end{longtable}

We discussed several possible strategies to select the final class
solution. Here, we apply our own.

To aid our interpretation of the results, we create an elbow plot
showing the trends in information criteria across four models.

From the elbow plot, we see that AIC has a lower penalty for model
complexity. However, we are more interested in the BIC values, which are
similar for the one, two and three-class solutions, but the four-class
solution fits significantly worse. For this reason, we eliminate the
four-class solution from the selection process.

Then we examine the model fit table. As expected, the -2*log likelihood
falls successively with each added class. As previously stated,
classification diagnostics should not be used for model selection, but
they can be used to disqualify certain solutions because they are
uninterpretable. We see that prob\_min and n\_min for the four-class solution is
low, knowing that this solution also has a high BIC, we disqualify this
solution.

Out of the remaining three solutions, we notice that entropy is the
highest for the three-class solution, and it has a satisfactory prob\_min
and n\_min. Based on this, we retain the three class solution in model
selection. Note that entropy for the one-class solution will always
equal to one, as it is 100\% true that every case is in that class. Based
on the low entropy of the two-class solution, we eliminate this model.

Finally, when comparing the one and three-class solutions, we inspect
the information criteria. For BIC, the one-class solution fits better,
but the difference is marginal. AIC tells us that the added complexity
of having three classes still explains the data better than a one-class
solution. Therefore, we select the three-class solution as our
final-class solution.

\hypertarget{interpreting-the-final-class-solution}{%
\subsection{Interpreting the Final Class Solution}\label{interpreting-the-final-class-solution}}

To aid our understanding of the final class solution, we use
\texttt{ggplot2::plot\_prob()} with the results of the three-class model as the
input. The resulting graph shows response patterns on all the indicators
for each group.

If we want to know the probability of each response option's endorsement
for each class, we can use \texttt{tidySEM::table\_prob}. These are thresholds
for ordinal dependent variables in the probability scale.

In the plot, we can see the distributions of the response probabilities
on the indicators for each of the three classes. For instance, we see
that in Class 1 the most common response to u2 is 2, while in Class 2
and Class 3 this is 0. We can also see that response 1 is a rare
response not forming the majority in any class. Class 2 distinguishes
itself because the majority scores the response 0 category of u3 and u4,
while in Class 1 and 2 this is not the case. Class 3 distinguishes
itself because the most common response to u3 and u4 is 3.

We can also interpret the response patterns numerically. It is a matter of preferences on how to interpret these probabilities. Here is where you would \textbf{name} each class, such that each response pattern is theoretically meaningful.

\hypertarget{extracting-posterior-class-probabilities}{%
\subsection{Extracting Posterior Class Probabilities}\label{extracting-posterior-class-probabilities}}

Another step is to extract posterior class probabilities. This is done
by the use of \texttt{tidySEM::class\_prob} with the results of the final class
solution as the input.

\newpage

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-akogul_comparison_2016}{}}%
Akogul, S., \& Erisoglu, M. (2016). A {Comparison} of {Information} {Criteria} in {Clustering} {Based} on {Mixture} of {Multivariate} {Normal} {Distributions}. \emph{Mathematical and Computational Applications}, \emph{21}(3), 34. \url{https://doi.org/10.3390/mca21030034}

\leavevmode\vadjust pre{\hypertarget{ref-baughman_mixture_2006}{}}%
Baughman, A. L., Bisgard, K. M., Lynn, F., \& Meade, B. D. (2006). Mixture model analysis for establishing a diagnostic cut-off point for pertussis antibody levels. \emph{Statistics in Medicine}, \emph{25}(17), 2994--3010. \url{https://doi.org/10.1002/sim.2442}

\leavevmode\vadjust pre{\hypertarget{ref-collins_latent_2010}{}}%
Collins, L. M., \& Lanza, S. T. (2010). \emph{Latent {Class} and {Latent} {Transition} {Analysis}: {With} {Applications} in the {Social}, {Behavioral}, and {Health} {Science}}. New Jersey, USA: John Wiley \& Sons, Inc.

\leavevmode\vadjust pre{\hypertarget{ref-enders_applied_2010}{}}%
Enders, C. K. (2010). \emph{Applied missing data analysis}. New York: Guilford Press.

\leavevmode\vadjust pre{\hypertarget{ref-figueiredo_unsupervised_2002}{}}%
Figueiredo, M. A. T., \& Jain, A. K. (2002). Unsupervised learning of finite mixture models. \emph{IEEE Transactions on Pattern Analysis and Machine Intelligence}, \emph{24}(3), 381--396. \url{https://doi.org/10.1109/34.990138}

\leavevmode\vadjust pre{\hypertarget{ref-geiser_data_2012}{}}%
Geiser, C. (2012). \emph{Data {Analysis} with {Mplus}}. Guilford Press.

\leavevmode\vadjust pre{\hypertarget{ref-geiser_is_2014}{}}%
Geiser, C., \& Wurpts, I. (2014). Is adding more indicators to a latent class analysis beneficial or detrimental? {Results} of a {Monte} {Carlo} study. \emph{Frontiers in Psychology: Quantitative Psychology and Measurement}, \emph{5}. https://doi.org/\url{https://doi.org/10.3389/fpsyg.2014.00920}

\leavevmode\vadjust pre{\hypertarget{ref-giang_using_2008}{}}%
Giang, M. T., \& Graham, S. (2008). Using latent class analysis to identify aggressors and victims of peer harassment. \emph{Aggressive Behavior}, \emph{34}(2), 203--213. \url{https://doi.org/10.1002/ab.20233}

\leavevmode\vadjust pre{\hypertarget{ref-hardy_exploring_2018}{}}%
Hardy, T. M., Garnier-Villarreal, M., McCarthy, D. O., Anderson, R. A., \& Reynolds, R. M. (2018). Exploring the {Ovarian} {Reserve} {Within} {Health} {Parameters}: {A} {Latent} {Class} {Analysis}. \emph{Western Journal of Nursing Research}, 1--16. \url{https://doi.org/10.1177/0193945918792303}

\leavevmode\vadjust pre{\hypertarget{ref-hennig_handbook_2015}{}}%
Hennig, C., Meila, M., Murtagh, F., \& Rocci, R. (2015). \emph{Handbook of {Cluster} {Analysis}}. 28.

\leavevmode\vadjust pre{\hypertarget{ref-hopfer_social_2014}{}}%
Hopfer, S., Tan, X., \& Wylie, J. L. (2014). A {Social} {Network}--{Informed} {Latent} {Class} {Analysis} of {Patterns} of {Substance} {Use}, {Sexual} {Behavior}, and {Mental} {Health}: {Social} {Network} {Study} {III}, {Winnipeg}, {Manitoba}, {Canada}. \emph{American Journal of Public Health}, \emph{104}(5), 834--839. \url{https://doi.org/10.2105/AJPH.2013.301833}

\leavevmode\vadjust pre{\hypertarget{ref-jamshidian_tests_2010}{}}%
Jamshidian, M., \& Jalal, S. (2010). Tests of {Homoscedasticity}, {Normality}, and {Missing} {Completely} at {Random} for {Incomplete} {Multivariate} {Data}. \emph{Psychometrika}, \emph{75}(4), 649--674. \url{https://doi.org/10.1007/s11336-010-9175-3}

\leavevmode\vadjust pre{\hypertarget{ref-van_kollenburg_lazy_2018}{}}%
Kollenburg, G. H. van, Mulder, J., \& Vermunt, J. K. (2018). \emph{The {Lazy} {Bootstrap}. {A} {Fast} {Resampling} {Method} for {Evaluating} {Latent} {Class} {Model} {Fit}}. 23.

\leavevmode\vadjust pre{\hypertarget{ref-lanza_introduction_2003}{}}%
Lanza, S. T., Bray, B. C., \& Collins, L. M. (2003). An {Introduction} to {Latent} {Class} and {Latent} {Transition} {Analysis}. In \emph{Handbook of {Psychology}: {Research} {Methods} in {Psychology}} (2nd ed., Vol. 2, pp. 690--712). John Wiley \& Sons.

\leavevmode\vadjust pre{\hypertarget{ref-lanza_latent_2013}{}}%
Lanza, S. T., Tan, X., \& Bray, B. C. (2013). Latent {Class} {Analysis} {With} {Distal} {Outcomes}: {A} {Flexible} {Model}-{Based} {Approach}. \emph{Structural Equation Modeling : A Multidisciplinary Journal}, \emph{20}(1), 1--26. \url{https://doi.org/10.1080/10705511.2013.742377}

\leavevmode\vadjust pre{\hypertarget{ref-lee_comparison_2021}{}}%
Lee, T., \& Shi, D. (2021). A comparison of full information maximum likelihood and multiple imputation in structural equation modeling with missing data. \emph{Psychological Methods}, No Pagination Specified--No Pagination Specified. \url{https://doi.org/10.1037/met0000381}

\leavevmode\vadjust pre{\hypertarget{ref-little_test_1988}{}}%
Little, R. J. A. (1988). A {Test} of {Missing} {Completely} at {Random} for {Multivariate} {Data} with {Missing} {Values}. \emph{Journal of the American Statistical Association}, \emph{83}(404), pp. 1198--1202. \url{https://doi.org/10.2307/2290157}

\leavevmode\vadjust pre{\hypertarget{ref-macgregor_symptom_2021}{}}%
MacGregor, A. J., Dougherty, A. L., D'Souza, E. W., McCabe, C. T., Crouch, D. J., Zouris, J. M., \ldots{} Fraser, J. J. (2021). Symptom profiles following combat injury and long-term quality of life: A latent class analysis. \emph{Quality of Life Research}, \emph{30}(9), 2531--2540. \url{https://doi.org/10.1007/s11136-021-02836-y}

\leavevmode\vadjust pre{\hypertarget{ref-masyn_latent_2013}{}}%
Masyn, K. E. (2013). Latent {Class} {Analysis} and {Finite} {Mixture} {Modeling}. In \emph{The {Oxford} {Handbook} of {Quantitative} {Methods}}: \emph{Vol.} \emph{2: Statistical Analysis} (p. 551). Oxford University Press.

\leavevmode\vadjust pre{\hypertarget{ref-mcelreath_statistical_2020}{}}%
McElreath, R. (2020). \emph{Statistical rethinking: A {Bayesian} course with examples in {R} and {Stan}} (2nd ed.). Boca Raton: Taylor; Francis, CRC Press.

\leavevmode\vadjust pre{\hypertarget{ref-openmx}{}}%
Neale, M. C., Hunter, M. D., Pritikin, J. N., Zahery, M., Brick, T. R., Kirkpatrick, R. M., \ldots{} Boker, S. M. (2016). Open{M}x 2.0: {E}xtended structural equation and statistical modeling. \emph{Psychometrika}, \emph{81}(2), 535--549. \url{https://doi.org/10.1007/s11336-014-9435-8}

\leavevmode\vadjust pre{\hypertarget{ref-norman_likert_2010}{}}%
Norman, G. (2010). Likert scales, levels of measurement and the {``laws''} of statistics. \emph{Advances in Health Sciences Education}, \emph{15}(5), 625--632. \url{https://doi.org/10.1007/s10459-010-9222-y}

\leavevmode\vadjust pre{\hypertarget{ref-nosek_promoting_2015}{}}%
Nosek, B. A., Alter, G., Banks, G. C., Borsboom, D., Bowman, S. D., Breckler, S. J., \ldots{} Yarkoni, T. (2015). Promoting an open research culture. \emph{Science}, \emph{348}(6242), 1422--1425. \url{https://doi.org/10.1126/science.aab2374}

\leavevmode\vadjust pre{\hypertarget{ref-nozadi_moderating_2016}{}}%
Nozadi, S. S., Troller-Renfree, S., White, L. K., Frenkel, T., Degnan, K. A., Bar-Haim, Y., \ldots{} Fox, N. A. (2016). The {Moderating} {Role} of {Attention} {Biases} in understanding the link between {Behavioral} {Inhibition} and {Anxiety}. \emph{Journal of Experimental Psychopathology}, \emph{7}(3), 451--465. \url{https://doi.org/10.5127/jep.052515}

\leavevmode\vadjust pre{\hypertarget{ref-nylund_deciding_2007}{}}%
Nylund, K. L., Asparouhov, T., \& Muthén, B. O. (2007). Deciding on the {Number} of {Classes} in {Latent} {Class} {Analysis} and {Growth} {Mixture} {Modeling}: {A} {Monte} {Carlo} {Simulation} {Study}. \emph{Structural Equation Modeling: A Multidisciplinary Journal}, \emph{14}(4), 535--569. \url{https://doi.org/10.1080/10705510701575396}

\leavevmode\vadjust pre{\hypertarget{ref-nylund-gibson_ten_2018}{}}%
Nylund-Gibson, K., \& Choi, A. Y. (2018). Ten frequently asked questions about latent class analysis. \emph{Translational Issues in Psychological Science}, \emph{4}(4), 440--461. \url{https://doi.org/10.1037/tps0000176}

\leavevmode\vadjust pre{\hypertarget{ref-pavlopoulos_measuring_2015}{}}%
Pavlopoulos, D., \& Vermunt, J. K. (2015). Measuring {Temporary} {Employment}. {Do} {Survey} or {Register} {Data} {Tell} the {Truth}? \emph{Survey Methodology}, \emph{41}(1), 197--214. Retrieved from \url{https://www150.statcan.gc.ca/n1/pub/12-001-x/2015001/article/14151-eng.pdf}

\leavevmode\vadjust pre{\hypertarget{ref-rhemtulla_when_2012}{}}%
Rhemtulla, M., Brosseau-Liard, P. É., \& Savalei, V. (2012). When can categorical variables be treated as continuous? {A} comparison of robust continuous and categorical {SEM} estimation methods under suboptimal conditions. \emph{Psychological Methods}, \emph{17}(3), 354--373. \url{https://doi.org/10.1037/a0029315}

\leavevmode\vadjust pre{\hypertarget{ref-lavaan}{}}%
Rosseel, Y. (2012). {lavaan}: An {R} package for structural equation modeling. \emph{Journal of Statistical Software}, \emph{48}(2), 1--36. \url{https://doi.org/10.18637/jss.v048.i02}

\leavevmode\vadjust pre{\hypertarget{ref-rubin_inference_1976}{}}%
Rubin, D. B. (1976). Inference and {Missing} {Data}. \emph{Biometrika}, \emph{63}(3), 581--592. \url{https://doi.org/10.2307/2335739}

\leavevmode\vadjust pre{\hypertarget{ref-scrucca_mclust_2016}{}}%
Scrucca, L., Fop, M., Murphy, T. B., \& Raftery, A. E. (2016). \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5096736}{Mclust 5: {Clustering}, {Classification} and {Density} {Estimation} {Using} {Gaussian} {Finite} {Mixture} {Models}}. \emph{The R Journal}, \emph{8}(1), 289--317.

\leavevmode\vadjust pre{\hypertarget{ref-spurk_latent_2020}{}}%
Spurk, D., Hirschi, A., Wang, M., Valero, D., \& Kauffeld, S. (2020). Latent profile analysis: {A} review and {``how to''} guide of its application within vocational behavior research. \emph{Journal of Vocational Behavior}, \emph{120}, 103445. \url{https://doi.org/10.1016/j.jvb.2020.103445}

\leavevmode\vadjust pre{\hypertarget{ref-van_de_schoot_grolts-checklist_2017}{}}%
Van De Schoot, R., Sijbrandij, M., Winter, S. D., Depaoli, S., \& Vermunt, J. K. (2017). The {GRoLTS}-{Checklist}: {Guidelines} for {Reporting} on {Latent} {Trajectory} {Studies}. \emph{Structural Equation Modeling: A Multidisciplinary Journal}, \emph{24}(3), 451--467. \url{https://doi.org/10.1080/10705511.2016.1247646}

\leavevmode\vadjust pre{\hypertarget{ref-van_lissa_worcs_2020}{}}%
Van Lissa, C. J., Brandmaier, A. M., Brinkman, L., Lamprecht, A.-L., Peikert, A., Struiksma, M. E., \& Vreede, B. (2020). \emph{{WORCS}: {A} {Workflow} for {Open} {Reproducible} {Code} in {Science}}. \url{https://doi.org/10.17605/OSF.IO/ZCVBS}

\leavevmode\vadjust pre{\hypertarget{ref-van_lissa_worcs_2021}{}}%
Van Lissa, C. J., Brandmaier, A. M., Brinkman, L., Lamprecht, A.-L., Peikert, A., Struiksma, M. E., \& Vreede, B. M. I. (2021). {WORCS}: {A} workflow for open reproducible code in science. \emph{Data Science}, \emph{4}(1), 29--49. \url{https://doi.org/10.3233/DS-210031}

\leavevmode\vadjust pre{\hypertarget{ref-vermunt_k-means_2011}{}}%
Vermunt, J. K. (2011). K-means may perform as well as mixture model clustering but may also be much worse: {Comment} on {Steinley} and {Brusco} (2011). \emph{Psychological Methods}, \emph{16}(1), 82--88. \url{https://doi.org/10.1037/a0020144}

\leavevmode\vadjust pre{\hypertarget{ref-vermunt_latent_2017}{}}%
Vermunt, J. K. (2017). Latent {Class} {Modeling} with {Covariates}: {Two} {Improved} {Three}-{Step} {Approaches}. \emph{Political Analysis}, \emph{18}(4), 450--469. \url{https://doi.org/10.1093/pan/mpq025}

\leavevmode\vadjust pre{\hypertarget{ref-vermunt_jk_latent_2004}{}}%
Vermunt, J.K., Magidson, J., Lewis-Beck, M., Bryman, A., Liao, T.F., \& Department of Methodology and Statistics. (2004). Latent class analysis. In \emph{The {Sage} encyclopedia of social sciences research methods} (pp. 549--553). Sage. Retrieved from \url{https://research.tilburguniversity.edu/en/publications/0caedd00-27c1-42bd-bb4e-d1dcb0864956}

\leavevmode\vadjust pre{\hypertarget{ref-weller_latent_2020}{}}%
Weller, B. E., Bowen, N. K., \& Faubert, S. J. (2020). Latent {Class} {Analysis}: {A} {Guide} to {Best} {Practice}. \emph{Journal of Black Psychology}, \emph{46}(4), 287--311. \url{https://doi.org/10.1177/0095798420930932}

\leavevmode\vadjust pre{\hypertarget{ref-wu_abuse_2011}{}}%
Wu, L.-T., Woody, G. E., Yang, C., Pan, J.-J., \& Blazer, D. G. (2011). Abuse and dependence on prescription opioids in adults: A mixture categorical and dimensional approach to diagnostic classification. \emph{Psychological Medicine}, \emph{41}(3), 653--664. \url{https://doi.org/10.1017/S0033291710000954}

\end{CSLReferences}

\end{document}