### 2022-11-22 example_analysis/LCA_Tutorial.Rmd
Update November 23
- updated the literature list
- added use case examples
- corrected spelling/grammar mistakes
- adjusted spacing, also in the tutorial

*** things to resolve:
1) Caspar should add Well-being compass study to Zotero and update the in-text citation for that use case
2) the plot_density function doesn't work, also the example in the documentation;
I would like to show how to use it in class enumeration section of the tutorial
3) final proof-reading, anything to add?

---
title: "LCA in R using tidySEM"
output: html_notebook
---
# Tutorial

This is an example of an exploratory LCA using `tidySEM`.

## Loading the Dataset

In this example, we use *data_mix_ordinal*, a simulated data for mixture
modeling with ordinal indicators. This dataset is built-in to
`tidySEM`. For more information about the dataset, type
`?tidySEM::data_mix_ordinal` into the R console after loading the
`tidySEM` using library().

We load the dataset and convert the indicators into ordered factors.

```{r, include = TRUE, eval=F}
library(tidySEM) # loading tidySEM
df <- data_mix_ordinal  # dataset
df[1:4] <- lapply(df, ordered) # indicators as ordered factors
```

## Exploring the Data

An important step to preceding any statistical analysis is data
exploration. Here we use `tidySEM::descriptives()` to describe the
dataset we are using.

```{r, include = TRUE, eval=F}
tidySEM::descriptives(df) # descriptives
```

As we can see, the output includes various descriptives of our dataset.
Special attention should be paid to examining the pattern of
missingness, as discussed in the *Handling missing data* section of this
paper. In our example, we see that there are no missing values, hence we
proceed with our analysis.

## Conducting Latent Class Analysis

Before we fit a series of LCA models, we set a random seed using
`set.seed()`. The seed number can be any digit. This is an important
step as there is some inherent randomness in the LCA computations, and
having the same seed number ensures that two separate researcher obtain
exactly the same results when fitting LCA models.

Finally, we reach the step of fitting LCA models. To do so, we use
`tidySEM::mx_lca()` which takes data and number of classes as inputs. 
Notice that the analysis will be done on all the variable in the data set, 
so you have to make sure it only includes the relevant variables. 
In our example, we want to fit 1 to 4 class solutions and compare 
their output. Depending on your computer's computational power, 
this might take a while.

```{r fitting_LCA, include = TRUE, eval=F}
set.seed(123) # setting seed 
res <- mx_lca(data = df, classes = 1:4) # fitting LCA 1 to 4 class solutions
```

Notice that the model run 10 times for each $K$ class solution. 
This way we can choose the best fitting model across all the solutions. 
In this example, we see that all models converged without issues.

We can inspect the class of resulting objects.

```{r, include = TRUE, eval=F}
is(res)
is(res[[1]])
```

As `tidySEM::mx_lca()` run several models at a time, the resulting object `res` 
is a `mixture_list` class where each $K$ class solution is a list element, or
more specifically an `OpenMx` model (`MxModel`).

## Class Enumeration

In class enumeration, we want to compare a sequence of LCA models fitted
to the data. To aid the process, we create a model fit table using
`tidySEM::table_fit()` with the results object as the input. As the
output contains a lot of information on each of the four fitted models,
we select a subset of helpful model fit indices and classification
diagnostics.

```{r fit_table, include = TRUE, eval=F}
fit_table <- table_fit(res) # model fit table
fit_table[ , c("Name", "LL", "Parameters", 
               "AIC", "BIC", "Entropy", 
               "prob_min", "prob_max", 
               "n_min", "n_max")] # our selection
```

Our selection of fit indices and classification diagnostics includes:

```{r, echo=F}
Selection <- c(
    "Name", "LL", "Parameters", 
    "AIC", "BIC", "prob_min", 
    "prob_max", "n_min", "n_max")

Description <- c(
    "the $K$-class solution",
    "the -2*log-likelihood of each model",
    "Number of estimated parameters in each model",
    "the Akaike Information Criterion value",
    "the Bayesian Information Criterion value",
    "the lowest posterior class probability by most likely class membership",
    "the highest posterior class probability by most likely class membership",
    "the lowest class proportion based on the posterior class probabilities",
    "the highest class proportion based on the posterior class probabilities")

selection <- cbind(Selection, Description)

knitr::kable(selection, format = 'pipe', 
             caption = "Selection of Fit Indices and 
                        Classification Diagnostics")
```

We discussed several possible strategies to select the final class
solution. Here, we apply our own.

To aid our interpretation of the results, we create an elbow plot
showing the trends in information criteria across four models.

```{r elbow_plot, include = TRUE, message=F, warning=F, eval=F}
library(tidyverse) # for data-wrangling
library(ggplot2) # for plots

elbow_plot <- fit_table[ , c("Name", "AIC", "BIC")] # extract ICs
elbow_plot <- pivot_longer(elbow_plot, cols = c("AIC", "BIC"), names_to = "IC", values_to = "Value") # to long format

ggplot(elbow_plot, aes(x = Name, y = Value, group = IC))+
  geom_point(aes(color = IC))+
  geom_line(aes(color = IC))+
  labs(title="Elbow Plot of Information Criteria per Class Solution", x = "Class Solution", y = " Value")+
  scale_color_manual(name = "Information Criterion", values = c(AIC = 'blue', BIC = 'red'))+
  theme_minimal()

```

From the elbow plot, we see that AIC has a lower penalty for model
complexity. However, we are more interested in the BIC values, which are
similar for the one, two and three-class solutions, but the four-class
solution fits significantly worse. For this reason, we eliminate the
four-class solution from the selection process.

Then we examine the model fit table. As expected, the -2\*log likelihood
falls successively with each added class. As previously stated,
classification diagnostics should not be used for model selection, but
they can be used to disqualify certain solutions because they are
uninterpretable. We see that prob_min and n_min for the four-class solution is
low, knowing that this solution also has a high BIC, we disqualify this
solution.

Out of the remaining three solutions, we notice that entropy is the
highest for the three-class solution, and it has a satisfactory prob_min
and n_min. Based on this, we retain the three class solution in model
selection. Note that entropy for the one-class solution will always
equal to one, as it is 100% true that every case is in that class. Based
on the low entropy of the two-class solution, we eliminate this model.

Finally, when comparing the one and three-class solutions, we inspect
the information criteria. For BIC, the one-class solution fits better,
but the difference is marginal. AIC tells us that the added complexity
of having three classes still explains the data better than a one-class
solution. Therefore, we select the three-class solution as our
final-class solution.

## Interpreting the Final Class Solution

To aid our understanding of the final class solution, we use
`ggplot2::plot_prob()` with the results of the three-class model as the
input. The resulting graph shows response patterns on all the indicators
for each group.

If we want to know the probability of each response option's endorsement
for each class, we can use `tidySEM::table_prob`. These are thresholds
for ordinal dependent variables in the probability scale.

```{r plot_prob, include = TRUE, eval=F}
plot_prob(res[[3]]) # visualizing the response patterns for the final class solution
table_prob(res[[3]]) # tabulating the response patterns for the final class solution
```

In the plot, we can see the distributions of the response probabilities
on the indicators for each of the three classes. For instance, we see
that in Class 1 the most common response to u2 is 2, while in Class 2
and Class 3 this is 0. We can also see that response 1 is a rare
response not forming the majority in any class. Class 2 distinguishes
itself because the majority scores the response 0 category of u3 and u4,
while in Class 1 and 2 this is not the case. Class 3 distinguishes
itself because the most common response to u3 and u4 is 3.

We can also interpret the response patterns numerically. It is a matter of preferences on how to interpret these probabilities. Here is where you would **name** each class, such that each response pattern is theoretically meaningful.

## Extracting Posterior Class Probabilities

Another step is to extract posterior class probabilities. This is done
by the use of `tidySEM::class_prob` with the results of the final class
solution as the input.

```{r extract_Post_Class_Prob, include = TRUE, eval=F}
probs <- class_prob(res[[3]]) # extracting the posterior class probabilities of the final class solution
probs$mostlikely.class # posterior class probabilities by most likely class membership
probs$individual # individual posterior class probabilities
```