### 2023-01-23 appendix_c.Rmd

---
title             : "Appendix C: Latent Class Growth Analysis"
shorttitle        : "BEST PRACTICES LATENT CLASS ANALYSIS"
bibliography      : "r-references.bib"
floatsintext      : yes
linenumbers       : no
draft             : no
mask              : no
figurelist        : no
tablelist         : no
footnotelist      : no
classoption       : "man"
output            : papaja::apa6_pdf
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
library(yaml)
library(scales)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  tidy.opts = list(width.cutoff = 60), tidy = TRUE
)
options(scipen = 1, digits = 2)

run_everything = FALSE

pkgdown <- TRUE

```

This vignette illustrated `tidySEM`'s ability to perform latent class
growth analysis, or growth mixture modeling. The simulated data used for
this example are inspired by work in progress by Plas and colleagues, on
heterogeneity in depression trajectories among Dutch military personnel
who were deployed to Afghanistan. The original data were collected as
part of the *Prospection in Stress-related Military Research (PRISMO)*
study, which examined of psychological problems after deployment in more
than 1,000 Dutch military personnel from 2005-2019.

First, we load all required packages:

```{r echo = TRUE, eval = TRUE}
library(tidySEM)
library(ggplot2)
library(MASS)
```

## Data preprocessing

We first examined the descriptive statistics for the sum score scales:

```{r, results='asis'}
# Get descriptives
df <- plas_depression
desc <- descriptives(df)
desc <- desc[, c("name", "mean", "median", "sd", "min", "max", 
"skew_2se", "kurt_2se")
]
papaja::apa_table(desc, caption = "Item descriptives")
```

Note that all variables were extremely right-skewed due to censoring at
the lower end of the scale.

We can examine these distributions visually as well:

```{r echo = TRUE, eval = FALSE}
df_plot <- reshape(df, direction = "long", varying = names(df))
ggplot(df_plot, aes(x = scl)) +
  geom_density() +
  facet_wrap(~time) + theme_bw()
```

```{r, eval = run_everything, echo = FALSE}
df_plot <- reshape(df, direction = "long", varying = names(df))
p = ggplot(df_plot, aes(x = scl)) +
  geom_density() +
  facet_wrap(~time) + theme_bw()
ggsave("appendices/plot_dist.png", p, width = 150, height = 120, units = "mm")
```

```{r, echo = FALSE, out.width="50%", eval = pkgdown}
knitr::include_graphics("appendices/plot_dist.png")
```

As this type of skew can result in convergence problems in LCGA, we
compared several transformations to reduce skew: The square and cube
root, log, inverse, and Box-Cox transformations.

```{r echo = TRUE, eval = FALSE, results='hide'}
df_scores <- df_plot
# Store original range of SCL
rng_scl <- range(df_scores$scl)
# Log-transform
df_scores$log <- scales::rescale(log(df_scores$scl), to = c(0, 1))
# Square root transform
df_scores$sqrt <- scales::rescale(sqrt(df_scores$scl), to = c(0, 1))
# Cube root transform
df_scores$qrt <- scales::rescale(df_scores$scl^.33, to = c(0, 1))
# Reciprocal transform
df_scores$reciprocal <- scales::rescale(1/df_scores$scl, to = c(0, 1))
# Define function for Box-Cox transformation
bc <- function(x, lambda){
  (((x ^ lambda) - 1) / lambda)
}
# Inverse Box-Cox transformation
invbc <- function(x, lambda){
  ((x*lambda)+1)^(1/lambda)
}
# Box-Cox transform
b <- MASS::boxcox(lm(df_scores$scl ~ 1), plotit = FALSE)
lambda <- b$x[which.max(b$y)]
df_scores$boxcox <- bc(df_scores$scl, lambda)
# Store range of Box-Cox transformed data
rng_bc <- range(df_scores$boxcox)
df_scores$boxcox <- scales::rescale(df_scores$boxcox, to = c(0, 1))
# Rescale SCL
df_scores$scl <- scales::rescale(df_scores$scl, to = c(0, 1))
```

We can plot these transformations:

```{r, eval = FALSE, echo = TRUE}
# Make plot data
df_plot <- do.call(rbind, lapply(c("scl", "log", "sqrt", "qrt", "boxcox"), function(n){
  data.frame(df_scores[c("time", "id")],
             Value = df_scores[[n]],
             Transformation = n)
}))
# Plot
ggplot(df_plot, aes(x = Value, colour = Transformation)) +
  geom_density() +
  facet_wrap(~time) +
  scale_y_sqrt() +
  xlab("scl (rescaled to 0-1)") +
  theme_bw()
```

```{r, eval = run_everything, echo = FALSE}
df_plot <- reshape(df, direction = "long", varying = names(df))
df_scores <- df_plot
# Store original range of SCL
rng_scl <- range(df_scores$scl)
# Log-transform
df_scores$log <- scales::rescale(log(df_scores$scl), to = c(0, 1))
# Square root transform
df_scores$sqrt <- scales::rescale(sqrt(df_scores$scl), to = c(0, 1))
# Cube root transform
df_scores$qrt <- scales::rescale(df_scores$scl^.33, to = c(0, 1))
# Reciprocal transform
df_scores$reciprocal <- scales::rescale(1/df_scores$scl, to = c(0, 1))
# Define function for Box-Cox transformation
bc <- function(x, lambda){
  (((x ^ lambda) - 1) / lambda)
}
# Inverse Box-Cox transformation
invbc <- function(x, lambda){
  ((x*lambda)+1)^(1/lambda)
}
# Box-Cox transform
b <- MASS::boxcox(lm(df_scores$scl ~ 1), plotit = FALSE)
lambda <- b$x[which.max(b$y)]
df_scores$boxcox <- bc(df_scores$scl, lambda)
# Store range of Box-Cox transformed data
rng_bc <- range(df_scores$boxcox)
df_scores$boxcox <- scales::rescale(df_scores$boxcox, to = c(0, 1))
# Rescale SCL
df_scores$scl <- scales::rescale(df_scores$scl, to = c(0, 1))

df_plot <- do.call(rbind, lapply(c("scl", "log", "sqrt", "qrt", "boxcox"), function(n){
  data.frame(df_scores[c("time", "id")],
             Value = df_scores[[n]],
             Transformation = n)
}))
# Plot
p <- ggplot(df_plot, aes(x = Value, colour = Transformation)) +
  geom_density() +
  facet_wrap(~time) +
  scale_y_sqrt() +
  xlab("scl (rescaled to 0-1)") +
  theme_bw()
ggsave("appendices/plot_trans.png", p, width = 150, height = 120, units = "mm")
```

```{r, echo = FALSE, eval = pkgdown, out.width="50%"}
knitr::include_graphics("appendices/plot_trans.png")
```

Evidently, the Box-Cox transformation reduced skew the most.
Consequently, we proceeded with the Box-Cox transformed scores for
analysis.

```{r, echo = TRUE, eval = run_everything}
dat <- df_scores[, c("id", "time", "boxcox")]
dat <- reshape(dat, direction = "wide", v.names = "boxcox", timevar = "time", idvar = "id")
names(dat) <- gsub("boxcox.", "scl", names(dat))
```

# Latent Class Growth Analysis

Next, we estimated a latent class growth model for SCL. The model
included an overall intercept, centered at T1, `i`. To model the
potential effect of deployment on depresion, we also included a dummy
variable that was zero before deployment, and 1 after deployment,
`step`. Finally, to model potential change (or recovery) in depression
post-deployment, we included a linear slope from T2-T6, `s`. All
variances of growth parameters were fixed to zero due to the sparse
nature of the data. In this vignette, we do not consider more than 5
classes, because the analyses are computationally very intensive and the
data were simulated from a 3-class model.

It is important to not that in LCGA, the subgroups will be limited by
the specify growth structure, meaning that LCA will identify distinctive
growth patterns within the intercept, step, and slope growth. For
example, if there is a subgroup that follows a quadratic growth pattern
this models will not be able to identify it.

**NOTE: The time scales in this model are not correct; it currently
assumes that all measurements are equidistant. Feel free to experiment
with adjusting this.**

```{r lcga, echo = TRUE, eval = FALSE}
set.seed(27796)
dat[["id"]] <- NULL
res_step <- mx_growth_mixture(
  model =
  "
  i =~ 1*scl1 + 1*scl2 + 1*scl3 +1*scl4 +1*scl5 +1*scl6
  step =~ 0*scl1 + 1*scl2 + 1*scl3 +1*scl4 +1*scl5 +1*scl6
  s =~ 0*scl1 + 0*scl2 + 1*scl3 +2*scl4 +3*scl5 +4*scl6
  scl1 ~~ vscl1*scl1
  scl2 ~~ vscl2*scl2
  scl3 ~~ vscl3*scl3
  scl4 ~~ vscl4*scl4
  scl5 ~~ vscl5*scl5
  scl6 ~~ vscl6*scl6
  i ~~ 0*i
  step ~~ 0*step
  s ~~ 0*s
  i ~~ 0*s
  i ~~ 0*step
  s ~~ 0*step", classes = 1:5,
  data = dat)
# Additional iterations because of
# convergence problems for model 1:
res_step[[1]] <- mxTryHardWideSearch(res_step[[1]], extraTries = 50)
```

```{r echo = FALSE, eval = run_everything}
set.seed(27796)
dat[["id"]] <- NULL
res_step <- mx_growth_mixture(
  model =
"i =~ 1*scl1 + 1*scl2 + 1*scl3 +1*scl4 +1*scl5 +1*scl6
step =~ 0*scl1 + 1*scl2 + 1*scl3 +1*scl4 +1*scl5 +1*scl6
s =~ 0*scl1 + 0*scl2 + 1*scl3 +2*scl4 +3*scl5 +4*scl6
scl1 ~~ vscl1*scl1
scl2 ~~ vscl2*scl2
scl3 ~~ vscl3*scl3
scl4 ~~ vscl4*scl4
scl5 ~~ vscl5*scl5
scl6 ~~ vscl6*scl6
i ~~ 0*i
step ~~ 0*step
s ~~ 0*s
i ~~ 0*s
i ~~ 0*step
s ~~ 0*step", classes = 1:5,
  data = dat)
# Additional iterations because of
# convergence problems for model 1:
res_step[[1]] <- mxTryHardWideSearch(res_step[[1]], extraTries = 50)
saveRDS(res_step, "appendices/res_step.RData")
fit <- table_fit(res_step)
saveRDS(fit, "appendices/lcga_fit.RData")
```

```{r, echo = FALSE, eval = FALSE}
res_step <- readRDS("appendices/res_step.RData")
```

Note that the first model showed convergence problems, throwing the
error: *The model does not satisfy the first-order optimality conditions
to the required accuracy, and no improved point for the merit function
could be found during the final linesearch.* To address this problem, we
performed additional iterations to\
find a better solution, using `OpenMx::mxTryHardWideSearch()`. This also
illustrates that `tidySEM` mixture models inherit from `OpenMx`'s
`MxModel`, and thus, different `OpenMx` functions can be used to act on
models specified via `tidySEM`.

The fifth model also evidenced convergence problems, but this (as we
will see) is because the solution is overfitted.

## Class enumeration

To determine the correct number of classes, we considered the following
criteria:

1.  We do not consider classes with, on average, fewer than 5
    participants per parameter in a class due to potential local
    underidentification
2.  Lower values for information criteria (AIC, BIC, saBIC) indicate
    better fit
3.  Significant Lo-Mendell-Rubin LRT test indicates better fit for $k$
    vs $k-1$ classes
4.  We do not consider solutions with entropy \< .90 because poor class
    separability compromises interpretability of the results
5.  We do not consider solutions with minimum posterior classification
    probability \< .90 because poor class separability compromises
    interpretability of the results

```{r eval = FALSE, echo = TRUE}
# Get fit table fit
tab_fit <- table_fit(res_step)
# Select columns
tab_fit <- tab_fit[, c("Name", "Classes", "LL", "Parameters", "BIC", "Entropy", "prob_min", "n_min", "warning", "lmr_p")]
tab_fit
```

```{r tabfit, eval = TRUE, echo = FALSE, results = "asis"}
# Get fit table fit
tab_fit <- readRDS("appendices/lcga_fit.RData")
tab_fit <- tab_fit[, c("Name", "Classes", "LL", "Parameters", "BIC", "Entropy", "prob_min", "n_min", "warning", "lmr_p")]
names(tab_fit) <- c("Name", "Classes", "LL", "p", "BIC", "Ent.", "p_min", "n_min", "warn", "lmr_p")
papaja::apa_table(tab_fit, digits = 2, caption = "Fit of LCGA models")
```

According to the Table, increasing the number of classes keeps
increasing model fit according to all ICs except the BIC, which
increased after 3 classes.

The first two LMR tests are significant, indicating that a 2- and
3-class solution were a significant improvement over a 1- and 2-class
solution, respectively. However, solutions with \>3 classes had entropy
and minimum posterior classification probability below the pre-specified
thresholds. Models with \>3 solutions also had fewer than five
observations per parameter. This suggests that the preferred model
should be selected from 1-3 classes.

### Scree plot

A scree plot indicates that the largest decrease in ICs occurs from 1-2
classes, and the inflection point for all ICs is at 3 classes. Moreover,
the BIC increased after 3 classes. A three-class solution thus appears
to be the most parsimonious solution with good fit.

```{r plotscree, echo = TRUE, eval = FALSE}
plot(tab_fit, statistics = c("AIC", "BIC", "saBIC"))
```

```{r echo = FALSE, eval = run_everything}
p <- plot(tab_fit, statistics = c("AIC", "BIC", "saBIC"))
ggsave("appendices/lcga_plot_fit.png", p, width = 150, height = 120, units = "mm")
```

```{r echo = FALSE, eval = pkgdown, out.width="50%"}
knitr::include_graphics("appendices/lcga_plot_fit.png")
```

Based on the aforementioned criteria, we selected a 3-class model for
further analyses. First, to prevent label switching, we re-order these
classes by the value of the intercept `i`. Then, we report the estimated
parameters.

```{r echo = TRUE, eval = FALSE}
res_final <- mx_switch_labels(res_step[[3]], param = "M[1,7]", decreasing = FALSE)
tab_res <- table_results(res_final, columns = NULL)
# Select rows and columns
tab_res <- tab_res[tab_res$Category %in% c("Means", "Variances"), c("Category", "lhs", "est", "se", "pval", "confint", "name")]
tab_res
```

```{r echo = FALSE, eval = run_everything, results='asis'}
res_final <- mx_switch_labels(res_step[[3]], param = "M[1,7]", decreasing = FALSE)
tab_res <- table_results(res_final, columns = NULL)
write.csv(tab_res, "lcga_tab_res.csv", row.names = FALSE)
```

```{r tabres, echo=FALSE, eval = TRUE, results='asis'}
tab_res <- read.csv("appendices/lcga_tab_res.csv", stringsAsFactors = FALSE)
class(tab_res) <- c("tidy_results", "data.frame")
tab_res <- tab_res[tab_res$Category %in% c("Means", "Variances"), c("Category", "lhs", "est", "se", "pval", "confint", "name")]
papaja::apa_table(tab_res, digits = 2, caption = "Results from 3-class LCGA model", longtable = TRUE)
```

As evident from these results, Class 1 started at a relatively lower
level of depressive symptoms, experienced a decrease after deployment,
followed by increase over time. Class 2 started at a moderate level of
depressive symptoms, experienced an increase after deployment, followed
by significant increase over time from T2-T6. Class 3 started at a
relatively higher level, experienced an increase after deployment,
followed by stability.

## Wald tests

To test whether parameters are significantly different between classes,
we can use Wald tests. Wald tests can be specified for all parameters in
the model, using the hypothesis syntax from the `bain` package for
informative hypothesis testing.

To identify the names of parameters in the model, we can use the `name`
column of the results table above. Alternatively, to see all parameters
in the model, run:

```{r params, echo = TRUE, eval = FALSE}
names(coef(res_final))
```

```{r echo = FALSE, eval = TRUE}
c("mix3.weights[1,2]", "mix3.weights[1,3]", "vscl1", "vscl2", 
"vscl3", "vscl4", "vscl5", "vscl6", "class1.M[1,7]", "class1.M[1,8]", 
"class1.M[1,9]", "class2.M[1,7]", "class2.M[1,8]", "class2.M[1,9]", 
"class3.M[1,7]", "class3.M[1,8]", "class3.M[1,9]")
```

Next, specify equality constrained hypotheses. For example, a hypothesis
that states that the mean intercept is equal across groups is specified
as follows:

`"class1.M[1,7] = class2.M[1,7] & class1.M[1,7] = class3.M[1,7]`

It is also possible to consider comparisons between two classes at a
time. When conducting many significance tests, consider correcting for
multiple comparisons however.

```{r echo = TRUE, eval = FALSE}
wald_tests <- wald_test(res_final, 
                   "
                   class1.M[1,7] = class2.M[1,7]&
                   class1.M[1,7] = class3.M[1,7];
                   class1.M[1,8] = class2.M[1,8]&
                   class1.M[1,8] = class3.M[1,8];
                   class1.M[1,9] = class2.M[1,9]&
                   class1.M[1,9] = class3.M[1,9]")
# Rename the hypothesis
wald_tests$Hypothesis <- c("Mean i", "Mean step", "Mean slope")
papaja::apa_table(wald_tests, digits = 2, caption = "Wald tests")
```

```{r echo = FALSE, eval = run_everything}
wald_tests <- wald_test(res_final, 
                   "class1.M[1,7] = class2.M[1,7]&class1.M[1,7] = class3.M[1,7];class1.M[1,8] = class2.M[1,8]&class1.M[1,8] = class3.M[1,8];class1.M[1,9] = class2.M[1,9]&class1.M[1,9] = class3.M[1,9]")
# Rename the hypothesis
wald_tests$Hypothesis <- c("Mean i", "Mean step", "Mean slope")
write.csv(wald_tests, "lcga_wald_tests.csv", row.names = FALSE)
```

```{r waldtests, echo = FALSE, eval = TRUE, results='asis'}
wald_tests <- read.csv("appendices/lcga_wald_tests.csv", stringsAsFactors = FALSE)
class(wald_tests) <- c("wald_test", "data.frame")
papaja::apa_table(wald_tests, digits = 2, caption = "Wald tests")
```

All Wald tests are significant, indicating that there are significant
differences between the intercepts, step function, and slopes of the
three classes.

## Trajectory plot

Finally, we can plot the growth trajectories. This can help interpret
the results better, as well as the residual heterogeneity around class
trajectories.

```{r makelcgaplot, echo = TRUE, eval = FALSE}
p <- plot_growth(res_step[[3]], rawdata = TRUE, alpha_range = c(0, .05))
# Add Y-axis breaks in original scale
brks <- seq(0, 1, length.out = 5)
labs <- round(invbc(scales::rescale(brks, from = c(0, 1), to = rng_bc), lambda))
p <- p + scale_y_continuous(breaks = seq(0, 1, length.out = 5), labels = labs) + ylab("SCL (rescaled from Box-Cox)")
p
```

```{r echo = FALSE, eval = run_everything}
p <- plot_growth(res_step[[3]], rawdata = TRUE, alpha_range = c(0, .05))
# Add Y-axis breaks in original scale
brks <- seq(0, 1, length.out = 5)
labs <- round(invbc(scales::rescale(brks, from = c(0, 1), to = rng_bc), lambda))
p <- p + scale_y_continuous(breaks = seq(0, 1, length.out = 5), labels = labs) + ylab("SCL (rescaled from Box-Cox)")
ggsave("appendices/plot_traj.png", p, width = 150, height = 120, units = "mm")
```

```{r echo = FALSE, eval = pkgdown, out.width="50%"}
knitr::include_graphics("appendices/plot_traj.png")
```

Note that the observed individual trajectories show very high
variability within classes.

<!-- \newpage -->

<!-- # References -->

<!-- ::: {#refs custom-style="Bibliography"} -->

<!-- ::: -->

## R session

```{r echo = TRUE, eval=TRUE}
sessionInfo()
```