### 2023-01-23 manuscript.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  ,man,floatsintext]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{latent class analysis, mixture models, best practices, free open source software, tidySEM\newline\indent Word count: 9001}
\usepackage{csquotes}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Best Practices in Latent Class Analysis using Free Open Source Software},
  pdfauthor={Caspar J. van Lissa1,2, Mauricio Garnier-Villarreal3, \& Daniel Anadria1},
  pdflang={en-EN},
  pdfkeywords={latent class analysis, mixture models, best practices, free open source software, tidySEM},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Best Practices in Latent Class Analysis using Free Open Source Software}
\author{Caspar J. van Lissa\textsuperscript{1,2}, Mauricio Garnier-Villarreal\textsuperscript{3}, \& Daniel Anadria\textsuperscript{1}}
\date{}

\shorttitle{BEST PRACTICES LATENT CLASS ANALYSIS}

\authornote{

Correspondence concerning this article should be addressed to Caspar J. van Lissa, Padualaan 14, 3584CH Utrecht, The Netherlands. E-mail: \href{mailto:c.j.vanlissa@gmail.com}{\nolinkurl{c.j.vanlissa@gmail.com}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} Utrecht University, Methodology \& Statistics\\\textsuperscript{2} Open Science Community Utrecht\\\textsuperscript{3} Vrije Universiteit Amsterdam, Sociology}

\abstract{%
Latent class analysis (LCA) refers to a family of techniques for identifying groups
in data based on a parametric model.
Examples include mixture models, LCA with ordinal
indicators, and latent class growth analysis.
Despite the popularity of this technique, there is limited
guidance with respect to best practices in estimating and reporting mixture
models. Moreover, although user-friendly interfaces for advanced mixture
modeling have long been available in commercial software packages, open
source alternatives have remained somewhat inaccessible. This tutorial
describes best practices for the estimation and reporting of LCA,
using free and open source software in R. To this end, this tutorial
introduces new functionality for estimating and reporting mixture
models in the \texttt{tidySEM} R-package whose backend relies on \texttt{OpenMx}.
}

\begin{document}
\maketitle

Latent class analysis (LCA) is an umbrella term that refers to a number
of techniques for estimating unobserved group membership based on a
parametric model of one or more observed indicators of group membership.
The types of LCA have become quite popular across scientific fields,
most notably finite Gaussian mixture modeling and latent profile
analysis. Jeroen K. Vermunt et al. (2004) defined LCA more generally as virtually
any statistical model where ``some of the parameters {[}\ldots{]} differ across
unobserved subgroups''.

Despite the popularity of LCA, two challenges impede broader adoption
and correct application of the method. The first challenge is that most
existing software that implements these models is commercial and
closed-source. The second challenge is that there is a lack of standards
for estimating and reporting LCA, and some myths and misunderstandings
have been perpetuated in prior literature. This introduces a risk of
misapplications of the technique, and complicates manuscript review and
assessment of the quality of published studies. The present paper seeks
to address these challenges by introducing updated guidelines for
estimating and reporting LCA, based on current best practice.
Furthermore, meeting the increasing demand for open source research
software, this paper introduces new functionality in the R-package
\texttt{tidySEM} (\textbf{vanlissaTidySEMTidyStructural2022?}) to perform LCA using free
open source software (FOSS).

\hypertarget{the-argument-for-free-open-source-software}{%
\subsection{The Argument for Free Open Source Software}\label{the-argument-for-free-open-source-software}}

Despite the popularity of LCA, most existing software is commercial and
closed-source, for example, Mplus (\textbf{muthenMplusUserGuide1998?}), for which
the R-package \texttt{MplusAutomation} offers augmented LCA functionality
(\textbf{hallquistMplusAutomationPackageFacilitating2018?}), and LatentGOLD
{[}REF{]}.
A
downside of commercial software is that its use is restricted to
individuals and institutions who can afford a license. Another downside
is that the proprietary source code cannot be audited, debugged, or
enhanced by third parties. This goes against open science principles
(\textbf{lamprechtFAIRPrinciplesResearch2019?}), incurs a risk that mistakes in
the source code go unnoticed, and curbs progress as independent
developers cannot fix bugs or add functionality.

There are several compelling reasons to prefer free open source software
(FOSS) instead. FOSS is freely available for anyone to use, modify, and
distribute, which means that scientists can access and use a wide range
of powerful tools without financial impediments. This is invaluable for
reducing gatekeeping of researchers with limited budgets, such as those
from developing countries. Furthermore, FOSS promotes collaboration
within the scientific community, as its users can contribute to the
development, maintenance, and support of the software. Finally, FOSS
promotes transparency and reproducibility: Because the source code for
FOSS is accessible, researchers can review and verify the methods and
algorithms used, which helps to ensure the integrity and reliability of
research findings. Overall, FOSS has the potential to significantly
improve the efficiency and effectiveness of scientific research, while
also promoting transparency, collaboration, and inclusion within the
scientific community.

Notable FOSS for LCA includes mclust (Scrucca, Fop, Murphy, \& Raftery, 2016), depmixS4
(Visser \& Speekenbrink, 2010), and OpenMx (Neale et al., 2016). A crucial limitation of
these existing packages is that some have limited functionality (e.g.,
Mclust), or implement specific innovations (e.g., depmixS4). OpenMx is
unique in that, like Mplus, it implement LCA in the context of a fully
featured structural equation modeling framework. Remaining obstacles to
the use of \texttt{OpenMx} is that it is not very user-friendly and its
functionality for LCA is poorly documented.

These limitations are addressed by the \texttt{tidySEM} package, which provide
a user-friendly interface for LCA in \texttt{OpenMx}. The package also includes
convenience functions to obtain relevant statistics, tests and plots in
accordance with best practices in estimation and reporting LCA, as
described in this paper. The primary purpose of the \texttt{tidySEM} package is
to provide a tidy workflow for generating, estimating, reporting, and
plotting structural equation models in a software-agnostic way. The user
interface is simple, as models can be specified using the widely used
\texttt{lavaan} syntax (Rosseel, 2012). The package contains
user-friendly wrapper functions to conduct LCA with sensible defaults,
but the resulting models can still be fully customized (low barrier,
high ceiling). Specifically, LCA models constructed in \texttt{tidySEM} inherit
from \texttt{OpenMx}' format \texttt{MxModel}, and remain compatible with downstream
functions that process such models. Being FOSS, \texttt{tidySEM} facilitates
fully-reproducible analyses and contributes to open science. It was
developed with open science in mind, and adheres to the FAIR software
principles (Findable, Accessible, Interoperable, and Reusable, \textbf{lamprechtFAIRPrinciplesResearch2019?}), and the \href{https://bestpractices.coreinfrastructure.org/en/criteria}{OpenSSF Best
Practices} for
FOSS software.

\hypertarget{defining-latent-class-analysis}{%
\subsection{Defining Latent Class Analysis}\label{defining-latent-class-analysis}}

LCA refers to methods for estimating unobserved groups based on a
parametric model of observed indicators of group membership
(\textbf{vermuntLatentClassAnalysis2004?}). It has become popular across
scientific fields, under a number of different names; most notably
(finite Gaussian) mixture modeling and latent profile analysis. The
concept of LCA can be understood in different ways. Generally speaking,
LCA assumes that the study population is composed of \(K\) subpopulations.
It further assumes that the observed data are a mixture of data
originating from those subpopulations; hence the name mixture model.
Consider the simplest possible univariate ``model'', which is a normal
distribution. This model has two parameters: the mean and the variance.
LCA aims to estimate the values of those parameters across a known
number of classes \(K\), as well as the probability of every individual
\(i\) belonging to classes \(1 \ldots k\). Commonly, the same model is
estimated across all classes, but with different parameters for each
class (i.e., class-specific means and variances).

As an illustrative example, imagine that a detective wants to know if it
would be possible to use mixture modeling to identify the sex of a
suspect, based on footprints found at the crime scene. To test the
feasibility of this approach, the detective records the shoe sizes and
sex of 100 volunteers. The resulting observed data look like this:

\begin{figure}
\includegraphics[width=7in]{shoedens} \caption{Kernel density plot of shoe sizes.}\label{fig:shoedens}
\end{figure}

The distribution is evidently bimodal, which bodes well for the intended
mixture model. In this case, the number of classes is known a-priori.
When estimating a two-class mixture model, the detective observes that
the model estimates the mean shoe size of the two groups are equal to
9.22 and
7.25, which is close to the true
means of the two groups, namely 9.04
and 6.93. When tabulating estimated
group membership against observed (known) group membership, it can be
seen that women are classified with a high degree of accuracy, but men
are not:

\begin{table}

\caption{\label{tab:tabshoe}Observed group membership by estimated class membership.}
\centering
\begin{tabular}[t]{l|r|r}
\hline
Observed & Class 1 & Class 2\\
\hline
Man & 28 & 21\\
\hline
Woman & 0 & 51\\
\hline
\end{tabular}
\end{table}

Another way to conceptualize LCA is by analogy to a measurement model in
structural equation modeling (\textbf{molenaarArbitraryNatureLatent1994?}). A
mixture model is like confirmatory factor analysis, except with a
categorical instead of a continuous latent variable. One difference
between the two techniques is that factor analysis can be considered as
a way to group observed \emph{variables} into latent constructs, and LCA
groups \emph{individuals} into classes. While factor analysis seeks to
describe the relations between variables, LCA seeks to describe
unobserved groups and classify individual observations to those groups.
In
line with this distinction, LCA (mixture models) is referred to as a
``person-centered'' technique, and factor analysis as a
``variable-centered'' technique (Masyn, 2013; Nylund-Gibson \& Choi, 2018).

When the focus is on the model parameters in each group, LCA can be
thought of as similar to a multi-group structural equation model. The
main distinction is that group membership is not known a-priori, but
estimated from the data. Whereas in a multi-group model, the data are
split by group and treated as independent samples, in LCA, all cases
contribute to the estimation of all parameters in all groups. The
relative contribution of each case to the parameters of each group is
determined by that case's posterior probability of belonging to that
group.

When the focus is on each individual's estimated class membership, LCA
can be thought of as a type of clustering algorithm that corrects for
(random) measurement error. Specifically, the posterior class
probability that an individual belongs to a latent class can be computed
from the likelihood of that individual's observed data under given the
class-specific model. These posterior class probabilities can be used to
weight follow-up analyses, or individuals can be assigned to classes
based on their highest class probability. If the classes are clearly
distinct, measurement error is low, and the latter approach may be
acceptable. This perspective on mixture modeling is sometimes referred
to as ``model-based clustering'' (Hennig, Meila, Murtagh, \& Rocci, 2015; Scrucca et al., 2016). Many clustering algorithms apply some recursive
splitting algorithm to the data. By contrast ``model-based'' clustering
refers to the fact that LCA estimates cluster membership based on a
parametric model. Since a parametric model estimates only a fixed number
of parameters, it can be a relatively parsimonious solution compared to
non-parametric techniques.

Finally, in the context of machine learning, LCA can be considered as an
\emph{unsupervised} learning method for \emph{classification}
(Figueiredo \& Jain, 2002). The term \emph{unsupervised} refers to the
fact that the outcome variable -- true class membership -- is not known,
and the term \emph{classification} refers to the fact that the algorithm is
predicting a categorical outcome: class membership.

\hypertarget{a-taxonomy-of-latent-class-analysis-methods}{%
\subsection{A Taxonomy of Latent Class Analysis Methods}\label{a-taxonomy-of-latent-class-analysis-methods}}

The term LCA was initially conceived to refer to analyses with
categorical, usually binary indicators (Collins \& Lanza, 2009). It was
later generalized to refer to any technique that estimates a categorical
latent variable based on a parametric model of observed indicators, as
it only refers to the purpose of the analysis, and does not imply
restrictions to the model used, or the level of measurement of the
indicators (Jeroen K. Vermunt et al., 2004). To prevent ambiguity, we refer to LCA
with binary or ordinal indicators as \emph{LCA with ordinal indicators}.
Given the abundance of redundant terms in use for closely related types
of LCA models, we here provide a rudimentary taxonomy.

A common type of LCA is the \emph{finite Gaussian mixture model}; a
univariate analysis where the observed distribution of a single
continuous variable is assumed to result from a mixture of a known
number of normal (Gaussian) distributions. The parameters of a finite
Gaussian mixture model are the means and variances of these underlying
normal distributions. The analysis of shoe sizes presented earlier is a
canonical example of this type of analysis. In the multivariate case,
with more than one indicator variable, the parameters of a mixture model
are the means, variances, and covariances between the indicators (which
can be standardized to obtain correlations). These parameters can be
estimated freely, or set to be constrained across classes
(Collins \& Lanza, 2009).

The technique known as \emph{latent profile analysis (LPA)} is a special case
of the mixture model, which assumes conditional independence of the
indicators. Conditional independence means that, after class membership
is accounted for, the residual covariances/correlations between
indicators are assumed to be zero. In some cases, such constraints will
be inappropriate; for instance when the cohesion between indicators is
expected to differ between classes (Collins \& Lanza, 2009). Consider, for
example, a mixture model of ocean plastic particles, which found two
classes of particles based on length and width: A class of larger
particles with a low correlation between length and width, and a class
of smaller particles with a high correlation between length and width.
The reason for this difference in correlations makes theoretical sense:
the large particles were heterogenous in shape, whereas the smaller
particles had been polished to a more uniform (rounded) shape by the
elements (Alkema, Van Lissa, Kooi, \& Koelmans, 2022).

A longitudinal extension of LCA is the Hidden Markov Model (HMM, also
called Latent Transition Analysis), which estimates the change in class
membership over time with the markov assumption over time. The markov
assumption states that the class membership at time \(t\) is only affected
by the class membership at time \(t-1\). The structural part of the model
is comprise of the initial states (proportion of the sample in each
class at the first time point), and transition probabilities
(probabilities to switch class membership across time). The transition
probabilities are estimated as multinomial logistic regressions between
class memberships (Jeroen K. Vermunt, 2010).

Finally, some LCA models use latent variables as indicators. A common
application of this approach is in longitudinal research with repeated
measures of a construct. Examples of this approach include \emph{growth
mixture models} (GMM) and \emph{latent class growth analyses} (LCGA)
(Jung \& Wickrama, 2008). These techniques estimate a latent growth
model to describe grouped trajectories over time. The indicators of
class membership in a GMM are the intercepts, and variances, and
covariances of the latent growth variables that describe individual
trajectories on the observed variables. For linear trajectories, these
latent variables are an intercept and slope. The free variances allow
trajectories to vary within classes. The LCGA differs from the GMM in
that it assumes homogeneity of trajectories within classes. This is
achieved by restricting the (co)variances of the growth parameters to
zero. Any variance in the indicators not explained by the class-specific
latent trajectories is assumed to be error variance.

\hypertarget{use-cases-for-latent-class-analysis}{%
\subsection{Use Cases for Latent Class Analysis}\label{use-cases-for-latent-class-analysis}}

There are several use cases for which LCA methods are suitable.

\hypertarget{testing-theory}{%
\subsubsection{Testing theory}\label{testing-theory}}

Although LCA is often discussed as an exploratory analysis technique, it
can also be used in a confirmatory manner. Given that LCA is similar to
confirmatory factor analysis, it can be used to similar ends: To assess
whether a theoretical measurement model holds. This use case is relevant
when testing a theory that postulates the existence of a categorical
latent variable. For example, \emph{identity status theory} posits that, at
any given point in time, adolescents reside in one of four identity
statuses (Marcia, 1966). LCA can be used to identify these
four statuses based on observed indicators (e.g., self-reported identity
exploration and commitment)
(\textbf{luyckxDevelopmentalTypologiesIdentity2008?}). Similarly, \emph{personality
type theory} states that dimensional differences in personality can
largely be explained by an undercontrolled, overcontrolled, and
resilient type, which could also be restored using LCA
(\textbf{donnellanResilientOvercontrolledUndercontrolled2010?}).

One challenge is that, unlike confirmatory factor analysis, LCA does not
provide absolute fit indices (see Model Fit Indices). Thus, to test a
theory, one can ascertain that the data are better described by the
number of classes dictated by theory than by a different number of
classes. Furthermore, it would be possible to test whether the observed
pattern of class-specific means corresponds to a hypothesized pattern;
either qualitatively or quantitatively, using significance tests for
pre-specified values (see Inference).

Appendix A is a tutorial to confirmatory LCA in \texttt{tidySEM}, based on a
recent study that tested the theory that identities of Belgian
adolescents with migration backgrounds could be summarized as those
relating to their heritage, national and regional identity
(Maene, D'hondt, Van Lissa, Thijs, \& Stevens, 2022).

\hypertarget{exploring-heterogeneity}{%
\subsubsection{Exploring Heterogeneity}\label{exploring-heterogeneity}}

Another use case is to explore whether a population is comprised of
latent subpopulations, and if so, how many. The primary focus in this
use case is on class enumeration. For example,
one study explored heterogeneity in the population of
military service members and veterans with prior suicide attempts, and
identified two latent profiles characterized by high versus low suicide
risk based on self-reported suicidal ideation and information about
prior attempted methods and severity (Gromatsky et al., 2022).
Knowing that the population of suicidal service members is heterogeneous
can inform future research, policy, and care provided. Appendix B offers
a tutorial for exploratory LCA, based on the aforementioned study of the
distribution of ocean microplastics (Alkema et al., 2022).

\hypertarget{measurement-model}{%
\subsubsection{Measurement model}\label{measurement-model}}

LCA can also be used as a measurement model, similar to confirmatory
factor analysis. This is useful when class membership is measured
imperfectly by several indicators: LCA can partial out measurement error
and restore most likely class membership. For example, one study found
discrepancies between self-reported employment status and official
register data (Pavlopoulos \& Vermunt, 2015). While it might be tempting
to assume that register data is free of error, this assumption was shown
to be false. A mixture model was used to estimate employment status,
correcting for measurement error, from both self-report and registry
data. The study investigated random and systematic measurement error in
both indicators, which reveal strengths and weaknesses of both types of
assessment.

This example also illustrates that LCA allows us to evaluate the
reliability of different indicators (Geiser \& Wurpts, 2014).
A
continuous indicator that discriminates well between classes will show
large differences in class-specific means; an ordinal indicator will
show very high or very low conditional response probabilities for some
classes but not for others. This information can be used to select the
indicators that are most diagnostic of class membership. For an example
of LCA with a focus on differential item functioning, consider a study
on university admission tests of Saudi Arabian students
(Tsaousis, Sideridis, \& AlGhamdi, 2020).

\hypertarget{dimension-reduction}{%
\subsubsection{Dimension reduction}\label{dimension-reduction}}

LCA can also be used to reduce high-dimensional data to just a few
prototypes. Factor analysis is also sometimes used for dimension
reduction, but only accounts for linear covariance between indicators,
whereas LCA can accommodate complex - but discrete - patterns. For
example, imagine that there exist two indicators - \(x\) and \(y\), and some
individuals score high on both, some score low on both, and some score
high on \(x\) but low on \(y\). Factor analysis would only be able to
capture the differences between the first two groups well; LCA captures
differences between all three. Note that this is a pragmatic use of LCA;
there is no theory about the existence of a categorical latent variable
(cf.~Testing theory). Consequently, the class solution should be treated
as descriptive of the sample, and not be ``reified'' - treated as evidence
of the existence of a categorical latent variable.

The use of LCA for dimension reduction is common in longitudinal
research, where one developmental process is summarized using latent
class growth analysis (LCGA), and the resulting categories are then used
as a moderator of another developmental process in a multi-group model.
For example, one study conducted a LCGA of adolescent empathy
development on two dimensions of empathy and found three groups: high,
average, and low empathy (Van Lissa et al., 2015). These groups were
then used as moderator of a latent growth analysis of adolescent- and
parent-reported conflict. Results showed that the high-empathy group
evidenced greater adolescent-parent agreement about the incidence of
conflict than the other groups, and that the low empathy group had more
conflict with parents than the other groups. Note that both these key
findings are non-linear; modeling them in a single-group model would be
difficult. LCGA greatly simplifies model specification and
interpretation, albeit at a cost of some loss of information and,
potentially, generalizability. Appendix C is a tutorial on LCGA, based
on ongoing research on trajectories of depressive symptoms in military
service members after deployment.

The need for dimension reduction can be especially pertinent when using
ordinal indicators. For example, one study measured fifteen dichotomous
health indicators among injured military personnel
(MacGregor et al., 2021). Combinatorics informs us that fifteen
dichotomous items have \(2^{15}\) or \(32,768\) unique symptom combinations.
However, LCA was able to reduce response patterns to five clinically
meaningful latent classes.

\hypertarget{classification}{%
\subsubsection{Classification}\label{classification}}

Another use case is to estimate individuals' unobserved class membership
based on observed indicators. The shoe size example in Figure
\ref{fig:shoedens} is a rudimentary illustration of this application.
In applied research, LCA is often used for classification in clinical
contexts. For example, LCA was used to determine clinical cut-off scores
for diagnosing whooping cough (Baughman, Bisgard, Lynn, \& Meade, 2006). Another study
used LCA to not only classify the sample used to estimate the model, but
also future cases {[}REF Zegwaard et al, in prep{]}. The study identified
four types of care providers among those who supported close kin with
mental health problems. Importantly, the LCA model was embedded in an
interactive app that could be used by healthcare professionals to
determine the most likely class membership for new care providers, to
provide more targeted support. To obtain predicted class membership
based on an existing LCA model, the authors computed the likelihood of
data from a new participant under the multivariate normal distributions
estimated within all latent classes.
These applications of LCA aid the clinical decision making process, and
help decide whether additional support is warranted, or what type of
intervention is appropriate.

\hypertarget{violations-of-normality}{%
\subsubsection{Violations of normality}\label{violations-of-normality}}

Finally, LCA can be used to deal with data which violate certain
assumptions. For example, zero-inflated and hurdle models are special
cases of LCA for dealing data with a very high proportion of zeros
(extreme right skew) (\textbf{baughmanMixtureModelFramework2007?}). These models
estimate one class to explain the ``true zeros'', and another class with a
different distribution to explain all other values, which can include
zero. Note that this LCA has different models for both classes, and thus
falls beyond the scope of this paper.
For instance, the aforementioned study on marine plastic noted that many
prior studies had treated the distribution of particles as
(multivariate) normal, despite clear deviations from normality
(Alkema et al., 2022). A mixture of two normal distributions with
distinct variances described the observed distribution better.

\hypertarget{best-practices}{%
\section{Best Practices}\label{best-practices}}

The best practices in estimation, as outlined in Table
\ref{tab:checkest}, are rooted in existing recommendations for best
practices for estimating specific subtypes of LCA, including latent
class growth analysis (Van De Schoot, Sijbrandij, Winter, Depaoli, \& Vermunt, 2017) and latent
class analysis with ordinal indicators (e.g., Nylund-Gibson \& Choi, 2018).
These were generalized to be relevant to all types of LCA, and updated
to current best practices, as explained below.

\hypertarget{best-practices-in-estimation}{%
\subsection{Best Practices in Estimation}\label{best-practices-in-estimation}}

\begin{tabular}[t]{l|l}
\hline
\# & Item\\
\hline
1. & Examining Observed Data\\
\hline
2. & Handling Missing Data\\
\hline
3. & Alternative Model Specifications\\
\hline
4. & Software\\
\hline
5. & Algorithm\\
\hline
6. & Class Enumeration\\
\hline
7. & Model Fit Indices\\
\hline
8. & Classification Diagnostics\\
\hline
9. & Interpreting Class Solutions\\
\hline
10. & Label switching\\
\hline
\end{tabular}

\hypertarget{examining-observed-data}{%
\subsubsection{Examining Observed Data}\label{examining-observed-data}}

Examining observed data is essential for any analysis as it may reveal
patterns and violations of assumptions that had not been considered
prior to data collection. The function \texttt{descriptives(data)} provides
extensive descriptive statistics, including the number of unique values,
variance of continuous and categorical variables, missingness, skew and
kurtosis (peakedness of the distribution). However, sample-level
descriptive statistics are of limited value when the goal of a study is
to identify subsamples using LCA. Plots convey additional information;
specifically, density plots for continuous variables, and bar charts for
categorical ones. Density plots can also aid in the choice of the number
of classes, as explained in the section on visualization.

Particular points of attention include measurement level of the
indicators. Recall that \texttt{tidySEM} can estimate mixture models with
either numeric (type \texttt{numeric} and \texttt{integer}) or \texttt{ordered} indicators.
The column \texttt{type} should thus only contain these types. Also check that
the number of \texttt{unique} values match the intended measurement level.
Numeric variables with few unique values are typically better modeled as
ordinal. Our experience consulting on LCA methods and moderating the
\texttt{tidyLPA} Google group suggest that the misapplication of mixture models
to ordinal (e.g., Likert-type) indicators is the most common source of
user error. Whereas it has been argued that some parametric methods are
robust when scales with 7+ response categories are treated as continuous
(e.g., Norman, 2010), it is very unlikely that an ordinal
variable can be modeled as a \emph{mixture} of multiple normal distributions.
To illustrate the problem, consider what happens when estimating a
5-class solution on a single 5-point Likert scale indicator. Each
class-specific mean would describe a single response category, and a
class-specific variance component would be empirically underidentified.
In sum, variables with few unique values might not be suitable for
Gaussian mixture modeling; LCA with ordinal indicators is more
appropriate.

\hypertarget{data-preprocessing}{%
\subsubsection{Data preprocessing}\label{data-preprocessing}}

LCA can accommodate variables of different measurement levels:
continuous, binary, nominal, and ordinal. Numeric variables should have
an interval measurement level, such that it is sensible to estimate
means and (co)variances. For the remaining three measurement levels,
some preprocessing is required. The exact data types accepted by
\texttt{tidySEM} at the time of writing are \texttt{numeric} and \texttt{ordered}. To
correctly specify ordered variables such as Likert scales, use the
function \texttt{OpenMx::mxFactor()}. Binary variables, such as presence or
absence of some symptom, are also treated as ordinal, with one category
coded as zero and the other as one. Nominal variables are converted to
binary indicators of group membership via dummy coding, for example,
using \texttt{mx\_dummies()}.

A second consideration in data preprocessing is that models with
indicators on different scales may present with convergence issues. The
reason for this is that the parameter space is high-dimensional with
many potential local optima. If one indicator is on a much larger scale
than another, the parameter space may be more extensively explored for
that parameter. The problem can be resolved by transforming the
indicators to place them on (roughly) similar scales. The most common
transformation is standardization, which involves subtracting the mean
of each variable from each value and dividing by the standard deviation.
In R, this transformation is applied using the \texttt{scale()} function.
Standardization retains all available information, and model parameters
can be transformed back to the original scale of the indicators by
reversing the transformation. The transformed variable has a mean of
zero and a standard deviation of one. Note that standardization is
typically not necessary if indicators are already on the same, or
similar, scales.

A third consideration in preprocessing relates to the univariate
distributions of indicators. For example, the \texttt{plas\_depression} data
show extreme right skew in depressive symptoms: most respondents report
few symptoms, and few respondents report many symptoms. Similarly, the
\texttt{alkema\_microplastics} data are extremely right-skewed: Very small
plastic fragments abound in the oceans, and large pieces are rare. These
distributional characteristics must be taken into account either in
preprocessing, or model specification. In preprocessing, the data can be
transformed to reduce skew. After reducing skew, a simpler LCA model may
fit the data well. Alternatively, one can specify a model that
accommodates the distributional idiosyncrasies. For example, allowing
free variances across classes might result in one class with a near-zero
mean and small variance to account for the fact that most respondents
have few depressive symptoms, and one class with an elevated mean and
large variance to account for all respondents with non-zero symptoms.

\hypertarget{handling-missing-data}{%
\subsubsection{Handling Missing Data}\label{handling-missing-data}}

Previous work has emphasized the importance of examining the pattern of
missing data and reporting how missingness was handled
(Van De Schoot et al., 2017). While we concur that
investigating missingness is due diligence, it is important to emphasize
that missingness is adequately handled by default in \texttt{OpenMx}, which is
the computational backend of \texttt{tidySEM}. Its Full Information Maximum
Likelihood (FIML) estimator makes use of all available information
without imputing missing values, by estimating the model based on all
available information for each subject
(Enders, 2022; \textbf{endersRelativePerformanceFull2001?}). FIML is a
best-practice solution for handling missing data; at least on par with
multiple imputation (Lee \& Shi, 2021). Multiple imputation is less
suitable to LCA for two reasons. First, because LCA methods are often
computationally intensive, and conducting them on several imputed
datasets multiplies the load. Second, because many multiple imputation
methods assume normality; an assumption likely to be violated when data
come from a heterogeneous population.

Three types of missingness have been distinguished in the literature
(Enders, 2022; Rubin, 1976): random missingness
(MCAR); missingness contingent on \emph{observed} variables that can be known
and controlled for (MAR); and missingness related to unobserved
variables, which can be neither known nor controlled for (MNAR). A
so-called ``MCAR'' test is often reported (Jamshidian \& Jalal, 2010), but
note that this name is somewhat misleading, as a significant test
indicates that missingness is MAR, but a non-significant test statistic
does not distinguish between MCAR or MNAR. As the classic MCAR test
relies on the comparison of variances across groups with different
patterns of missing data, it assumes normality (Little, 1988). This
assumption is tenuous in the context of LCA. A non-parametric MCAR test
may be more suitable (Jamshidian \& Jalal, 2010). For this tutorial, the
lead author has implemented this test in the \texttt{mice} package as
\texttt{mice::mcar()}. Our recommendation is to inspect and report the
proportion of missingness per variable (e.g., using
\texttt{tidySEM::descriptives()}), and report the \texttt{mice::mcar()} test. As FIML
estimation assumes that missingness is either MCAR or MAR, one would
proceed with FIML regardless of the outcome.

One minor concern is that \texttt{tidySEM} uses a different algorithm to
determine starting values for complete data (K-means) and for missing
data (hierarchical clustering). This should not be a problem, however,
as the global optimizer used to estimate the model should be independent
of starting values (\textbf{coranaMinimizingMultimodalFunctions1987?}). A second
concern is that FIML is not suitable for cases with complete missing
data. If such cases must be included, their values must be imputed
first.

\hypertarget{model-specification}{%
\subsubsection{Model Specification}\label{model-specification}}

In principle, LCA can be based on any parametric model. At present,
\texttt{tidySEM} only facilitates estimation of models that can be specified in
\texttt{lavaan} syntax. The available model parameters depend on the
measurement level of the indicators. Continuous indicators can be
parametrized in terms of means, variances, and covariances. LCGA and GMM
have these same parameters, but with respect to the latent growth
variables. Ordinal indicators can be parametrized differently; the
default approach in \texttt{tidySEM} assumes that each ordinal indicator
reflects an underlying standard normal distribution,
\(N(\mu = 0, \sigma = 1)\). The parameters are ``thresholds'' that
correspond to quantiles of this distribution, estimated based on the
proportion of individuals in each of the response categories of the
indicator variable. For example, a binary indicator has a single
threshold that distinguishes the two response categories. If responses
are distributed 50/50, then the corresponding threshold would be
\(t_1 = 0.00\). If the responses are distributed 60/40, then the resulting
threshold would be \(t_1 = 0.25\). All of the aforementioned parameters
can be freely estimated, or constrained across classes, or fixed (e.g.,
to zero).

Prior literature emphasized the importance of considering alternative
model specifications
(\textbf{vandeschootGRoLTSChecklistGuidelinesReporting2017?}). What has remained
underemphasized is that alternative model specifications should be
approached differently for exploratory versus confirmatory analyses (see
Use Cases). Most prior literature has emphasized exploratory LCA (e.g., Nylund, Asparouhov, \& Muthén, 2007). In this context, model specification typically
consists of an exhaustive grid search along several model specifications
and varying numbers of classes. A ``final'' model is then selected based
on a combination of fit indices, significance tests, and
interpretability (see Class Enumeration).

Confirmatory LCA does not require such an extensive search over the
model space. If hypothesized models have been specified a priori, and
possibly even preregistered, it makes sense to focus on these models,
rather than proceed in a purely data-driven manner. Nonetheless, the
theoretical model specification should still be compared to a few
alternatives because LCA lacks objective fit indices. The fit of a
theoretical model can thus only be assessed by comparison to other
candidates. A one-class model is a sensible benchmark to test whether
the theoretical model fits better than a model that assumes a
homogeneous population. Other candidates for comparison include
competing theoretical models and those with a few more or less classes
than the hypothesized number. If the theoretical model has much better
fit than other models, for example, based on IC weights, this provides
evidence for the theory. If another model fits substantially better,
this provides evidence against the theory. If fit differences are small,
the theoretical model may still be preferred, as theory is a relevant
consideration in model selection (\textbf{jungIntroductionLatentClass2008?}). The
fact that the theoretical model received little differential support
should of course be discussed.

A second consideration in model specification is the risk of overfitting
(\textbf{hastieElementsStatisticalLearning2009?}). Overfitting occurs when a
model has many parameters relative to the number of observations,
causing it to capture idiosyncratic noise in the sample and limiting its
generalizability to new samples. As explained before, the number of
parameters in LCA scales with the number of estimated classes.
Consequently, LCA with many classes have a potentially very high number
of parameters. In Bayesian mixture modeling, it was observed that if the
estimated number of classes exceeded the true number, the number of
observations assigned to these classes tended towards zero as sample
size increased (\textbf{rousseauAsymptoticBehaviourPosterior2011?}). Although it
is not known whether frequentist LCA exhibits the same behavior, low
class counts are still an indicator of potential overfitting. The
section on Class Enumeration addresses ways to guard against
overfitting.

Models can be specified in \texttt{tidySEM} in different ways. The wrapper
functions \texttt{mx\_profiles()} and \texttt{mx\_lca()} construct LCA models for
continuous and ordinal indicators respectively, without needing to
specify model syntax. The function \texttt{mx\_profiles()} has two arguments to
free elements of the covariance matrix: \texttt{variances} and \texttt{covariances}.
By default, \texttt{mx\_profiles()} fixes covariances to zero, and thus performs
latent profile analysis. Finally, function \texttt{mx\_mixture()} accepts a more
flexible \texttt{model} argument, which can be a character string specifying a
model in \texttt{lavaan} syntax, which is translated to \texttt{OpenMx} models via the
function \texttt{as\_ram()}, or a list of \texttt{OpenMx} models.

\hypertarget{estimation-and-convergence}{%
\subsubsection{Estimation and Convergence}\label{estimation-and-convergence}}

LCA parameters and model fit statistics can be estimated in different
ways. The choice of estimator depends on the presence of missing values,
sample size, number of indicators, and available computational resources
(Weller, Bowen, \& Faubert, 2020). LCA in \texttt{tidySEM} uses maximum likelihood
estimation (ML), which can be conceptually understood as follows:
Imagine we are estimating two parameters, e.g.~the class-specific means
\(\mu_c\) on a continuous indicator for two classes (ignoring the variance
for now). ML attempts to find a combination of values for these two
parameters that minimizes minus two times the log likelihood (\(-2LL\)) of
the data.
To understand the ML optimization procedure, imagine a three-dimensional
landscape: The X and Y dimensions represent potential values of the
class-specific means, \(X = \mu_1\) and \(Y = \mu_2\), and the Z-dimension
is determined by \(Z = -2LL\). The optimizer must find the deepest
``valley'' in this landscape, which reflects the combination of \(\mu_1\)
and \(\mu_2\) that minimizes the \(-2LL\). The ML optimizer behaves somewhat
like a marble dropped in this landscape. It is dropped at some random
point in space, and rolls into the nearest valley (a combination of
values for \(\mu_1\) and \(\mu_2\) that has a low \(-2LL\)). The challenge is
that LCA models are often highly parametrized. This means that the
landscape is not three-dimensional, but (very) high-dimensional, with
many valleys and sparse data to provide information about which valley
is deepest. If the marble rolls into one of these valleys, it will
settle there and not climb out again. The estimator has ``converged''.

A unique challenge in LCA is that not only model parameters, but also
class membership must be estimated. Parameters in all classes are
estimated from the same dataset, but for each class, observations are
weighted by their probability of being in that class - which is in turn
computed based on the class-specific parameters. The resulting
chicken-and-egg problem is solved by the expectation-maximization (EM)
algorithm: It alternates between estimating model parameters,
calculating posterior class probabilities from those parameters,
re-estimating model parameters using those new class probabilities as
weights, and so on, until both have converged.

A risk with this approach is that the optimizer might get stuck in a
shallower valley (a ``local optimum''), and never discovers the deepest
valley (the ``global optimum'', or best solution). One solution to this
problem is to drop many marbles at random places, compare their final
\(-2LL\) values, choose the solution with the lowest \(-2LL\), and make sure
that several marbles replicated this solution. This is the ``random
starts'' approach implemented in, e.g., Mplus.

One problem with the random starts approach is that it is
computationally expensive to run this many replications. Moreover,
because the algorithm begins with random starting values, it is
inefficient because many marbles are likely to start far away from any
valley. Additionally, if all sets of starting values are close to a
specific local optimum, there is a risk that the global optimum is never
found. An alternative solution that overcomes these challenges is to use
a global optimizer, like simulated annealing (SA, \textbf{coranaMinimizingMultimodalFunctions1987?}). SA does not behave quite like
the marbles of the random starts approach; instead, it iteratively
considers some ``destination'' in the landscape, and compares its \(-2LL\)
to the current one. If the destination \(-2LL\) is lower, the estimator
moves there. However, if the destination \(-2LL\) is \emph{higher}, it still
moves there occasionally, with a certain probability. Think of this as
occasionally flicking the marble, to see if it rolls back into the
current valley, or finds another, even deeper one. This property allows
SA it to escape local optima and find the global optimum.

Since SA is a global optimizer, it should be independent of starting
values. With this in mind, it is efficient to start estimation from a
``reasonable solution'', instead of a random starting point. For example,
if we assume that the different classes are likely to have different
mean values on the indicators, then a nonparametric clustering
algorithm - like K-means or hierarchical clustering - can be used to
determine these cluster centroids (see also \textbf{biernackiChoosingStartingValues2003?}). By default, \texttt{tidySEM} derives
starting values using K-means clustering {[}REF{]}, computing the mixture
model starting values by treating the K-means solution as multi-group
model. Next, SA is used to find the global optimum solution. Finally, SA
is followed up with a short run of the ML algorithm, as ML inherently
produces an asymptotic covariance matrix for the parameters that can be
used to compute standard errors. Note that these defaults can be
manually overridden.

In practice, models do not always converge. Reasons may include local
optima, bad starting values, incorrect model specification (the model
may be too complex for the data, or conversely, too simple to capture
important patterns), or poor data quality (small data sets, or
distributions that are difficult to model). Convergence problems may be
indicated by errors and warnings from the estimator, or they may be
suspected, for example, when results change if the analyses are
reproduced.

Nonconvergence may be due to bad model specification. Consider, for
example, an LCA with multiple ordinal indicators. If certain
combinations of responses never occur together, in general or within one
of the classes, this can result in extreme conditional item
probabilities (exactly 0 or 1). Such extreme values impede model
convergence. This problem could be addressed by trying alternative model
specifications (e.g., fewer classes), or by merging categories of the
ordinal indicators with very few responses. Similarly problems occur
with continuous indicators. For instance, consider a zero-inflated
indicator and a two class LPA. To account for the excess zeroes, one
class will have a mean of zero. It is difficult to estimate the standard
deviation for this class, however, because all of its members have the
same value. If standard deviations are constrained to be equal across
classes, this model will likely not converge. This problem can be
addressed by considering more appropriate model specifications (e.g.,
free class variances).

Nonconvergence due to bad starting values can be overcome by manually
specifying starting values: Specify the argument \texttt{run\ =\ FALSE} in the
call to \texttt{mx\_mixture()} and related functions. This returns the model
with default starting values, which can be updated. Nonconvergence and
local optima may also be overcome by re-estimating the model while
permuting the starting values. The \texttt{OpenMx} package contains a family of
functions for this purpose: \texttt{mxTryHard()} and \texttt{mxTryHardWideSearch()}
for continuous indicators, and \texttt{mxTryHardOrdinal()} for ordinal
indicators. When poor data quality is to blame, the only solution is to
collect more, or higher quality, data.

\hypertarget{class-enumeration}{%
\subsubsection{Class Enumeration}\label{class-enumeration}}

Class enumeration refers to the process of determining the appropriate
number of classes. It is approached differently for exploratory versus
confirmatory LCA. In exploratory LCA, the set of models typically
consists of a sequence from 1 to the maximum number of classes \(K\). The
number \(K\) may be chosen a-priori or dictated by the data. It is
sensible to limit \(K\) to the maximum number of classes that result in
valid and interpretable solutions (also see Classification Diagnostics).
Above a certain number of classes, convergence problems typically arise,
or the proportion of cases assigned to the smallest classes may be low
relative to the number of class-specific parameters. In Bayesian LCA, it
has been observed that, when the estimated number of classes exceeds the
true number, the proportion of cases assigned to the excessive classes
converges to zero as the sample size increases
(\textbf{rousseauAsymptoticBehaviourPosterior2011?}). It is not known whether
this also happens in frequentist LCA; nonetheless, near-zero class
memberships should be taken into account when determining the maximum
\(K\).

In confirmatory LCA, the set of models with different numbers of classes
to be compared is typically smaller. Its purpose is to provide context
for the relative fit of the theoretical model, or to compare multiple
theoretical models.

In class enumeration, the set of models is compared on several
statistical and substantive criteria. In exploratory LCA, these criteria
may be decisive in selecting a final model; in the confirmatory case,
the burden of proof is reversed: there must be convincing evidence
against the theoretical model in order to reject it. Statistical
criteria to consider in class enumeration include model fit indices and
likelihood ratio tests (Masyn, 2013; Weller et al., 2020).
Substantive criteria include convergence and model identification, class
separability, and interpretability. We address each of these criteria
below.

\hypertarget{model-fit-indices}{%
\subsubsection{Model Fit Indices}\label{model-fit-indices}}

As there are no absolute fit indices for LCA, information criteria are
typically used to assess relative fit. Information criteria are computed
from the \(-2LL\), which is the most basic fit measure, testing the
overall model fit if the model properly represents the data. Information
criteria add a penalty for model complexity to the \(-2LL\), and thus
balance fit and complexity. The lower the value of an information
criterion, the better the overall fit of the model. The original
information criterion is the Akaike Information Criterion (AIC),
computed as \(-2LL+2p\), where \(p\) is the number of parameters. The
Bayesian Information Criterion (BIC) multiplies the penalty for
complexity with a function of the sample size, \(p \ln(n)\), such that
complex models are penalized more in smaller samples. Finally, the
so-called \emph{sample size adjusted} BIC (saBIC) implements a different
ad-hoc penalty for sample size, \(\ln(\frac{n + 2}{24})\). Although a
simulation study showed that the saBIC performed well for class
enumeration, its penalty is not based on theory {[}REF{]}. BIC also performs
well and is theoretically substantiated; as such, it may be the most
appropriate information criterion to use for model comparison
(Masyn, 2013; Nylund-Gibson \& Choi, 2018). All three ICs are
available in the output of \texttt{table\_fit()}.

Information criteria may occasionally contradict each other, so it is
important to identify a suitable strategy to reconcile them. One option
is to select a specific fit index, such as the BIC, before seeing the
results. This choice can be preregistered to avoid concerns about cherry
picking. Another option is to always prefer the most parsimonious model
that has best fit according to any of the available fit indices. Other
approach is to select several competing model based on the ICs, and
evaluate the theoretical intepretation. Yet another option is to
incorporate information from multiple fit indices using the analytic
hierarchy process (Akogul \& Erisoglu, 2016), which is implemented in
\texttt{tidyLPA} (\textbf{rosenbergTidyLPAPackageEasily2018?}). Finally, one can use a
scree plot to determine the inflection point at which additional classes
contribute little to further decreases of the ICs
(Nylund-Gibson \& Choi, 2018). The preferred number of classes is at the
inflection point where the slope of the curve is clearly leveling off.
Adding this many classes results in substantial decreases of the IC, and
adding further classes results in relatively small decreases.

Given a specific IC, it is also possible to calculate so-called
``IC-weights'', which indicate the relative support the data provide for
each model in the set (\textbf{wagenmakersAICModelSelection2004?}). This is
particularly useful for confirmatory LCA, as it allows one to quantify
the support for the theoretical model. IC-weights are calculated by
calling \texttt{ic\_weights()} on the model fit table. Note, however, that
IC-weights will (strongly) favor the model with the lowest IC value; the
scree plot approach above may be more suitable to select the most
parsimonious model with adequate fit.

\hypertarget{nested-model-tests}{%
\subsubsection{Nested Model Tests}\label{nested-model-tests}}

When two models are nested, the significance of the difference in fit
between these models can be tested using a Likelihood Ratio test (LR).
The test statistic is computed as the -2 times the log of the ratio of
likelihoods, and its degrees of freedom are equal to the difference in
the degrees of freedom of the two compared models. Since a \(k-1\)-class
model is nested in a \(k\)-class model, it is possible to test the
difference in model fit using a LR test. One challenge is that the LR
test statistic is not asymptotically \(\chi^2\) distributed for LCA. The
so-called Lo-Mendell-Rubin adjusted LR test (LMR) overcomes this problem
by applying an ad-hoc correction to the LR {[}REF lo 2001{]}. This test is
implemented in \texttt{tidySEM}, and is part of the default output of
\texttt{table\_fit()}, comparing subsequent rows of the table.

Another solution is to derive an empirical sampling distribution for the
LR difference. The so-called bootstrapped Likelihood Ratio Test (BLRT)
does so by simulating data from the \(k-1\)-class model \(B\) times (e.g.,
1000), and fitting the \(k\)- and \(k-1\)-class models to each simulated
dataset {[}REF Nylund et al., 2007{]}. The p-value of this test is the
proportion of \(B\) in which the LR of the models fit to simulated data
exceeds the LR observed in the real data. Simulation studies suggest
that the BLRT has greater statistical power than the LRT, although both
performed comparably well at detecting the true number of classes {[}REF
Nylund 2007, REF Tein 2013{]}. A crucial limitation of the BLRT is that it
is very computationally expensive. With this in mind, \texttt{table\_fit()} does
not provide the BLRT by default. Specific models can be compared using
the function \texttt{BLRT()}.

Anecdotal experience as a statistical co-author on LCA studies suggests
that both these tests tend to be liberal, returning significant p-values
for models with more classes even when other indicators suggest that
those models are overfitted (e.g., near-zero class counts). Other
authors have reported similar experiences
(\textbf{sinhaPractitionerGuideLatent2021?}), which matches the differences
between LRT and ICs from information theory as LRT tests and compare the
models based on in sample predictive accuracy, and ICs evaluates and
compare the models out-of-sample predictive accuracy which helps to
control for overfitting (McElreath, 2020). This suggests
that, like existing ICs, LR tests might not adequately balance fit and
complexity in LCA. Better solutions for class enumeration are thus an
important target for future research.

A promising novel solution is the so-called \emph{lazy bootstrap}, which is
much more efficient than the BLRT because it is based on sample
statistics of model-implied data generated from the \(k-1\) and \(k\)-class
models, rather than on fit statistics of LCA models fit to those data
(Kollenburg, Mulder, \& Vermunt, 2018). This simulation-based approach to inference
is similar to the posterior predictive p-value in Bayesian statistics,
which describes how well each model is able to reproduce the observed
correlation matrix (see McElreath, 2020). One version of the
lazy bootstrap is already implemented in the commercial program
latentGOLD {[}REF{]}. At the time of writing, we are developing a more
general simulation-based likelihood ratio test (SBLRT) for class
enumeration. Future versions of \texttt{tidySEM} will likely incorporate this
test.

\hypertarget{classification-diagnostics}{%
\subsubsection{Classification Diagnostics}\label{classification-diagnostics}}

Classification diagnostics assess whether the classes identified by LCA
are distinct and clearly separable. These diagnostics are not model fit
indices and should not be used for class enumeration, because a model
can fit the data well and still show poor latent class separation
(Masyn, 2013). However, classification diagnostics can be used to
eliminate models: If one analysis goal is to interpret properties of
(members of) classes, it makes sense to only consider models that
classify well (Nylund-Gibson \& Choi, 2018). Classification diagnostics
should thus not be used for model selection, but they can help restrict
the search space of acceptable models.

All classification diagnostics are derived from the posterior
classification probabilities: an \(n \times k\) matrix called \(P\), with
the probability that every individual \(n\) belongs to latent class \(k\).
When classification accuracy is high, and classes are distinct, each
individual's posterior class probabilities are high for one class, and
low for the remaining classes. The individual classification
probabilities \(P\) are obtained by running
\texttt{class\_prob(res,\ type\ =\ "individual")}. Note that one column is added to
the matrix, with each individual's most likely class membership \(M\). \(M\)
is obtained by assigning each individual to the class corresponding to
their maximum posterior class probability.
Predicted class
membership \(M\) is often of interest to researchers who wish to carry out
follow up analyses. Note, however, that follow-up analyses should take
classification error into account. The function \texttt{BCH()} does this
automatically.

The matrix \(P\) can be summarized in different way. For example, the
estimated frequency table of the latent categorical variable can be
obtained by summing the posterior probabilities by class (i.e., the
column sums of \(P\)). This frequency table \(F_P\) is obtained by calling
\texttt{class\_prob(res,\ type\ =\ "sum.posterior")}. These numbers are fractional
because this table takes classification error into account. Individuals
can contribute partially to multiple classes.

Second, most likely class membership based on the highest posterior
class probability, \(F_M\), is obtained by running
\texttt{class\_prob(res,\ "sum.mostlikely")}. This is the frequency table of the
variable \(M\). Unlike \(F_P\), \(F_M\) ignores classification error. If every
observation is classified with perfect accuracy, \(F_M\) and \(F_P\) are
identical. Part of the output of \texttt{table\_fit()} for LCA analyses are the
minimum and maximum percentage of the sample assigned to a particular
class, \texttt{n\_min} and \texttt{n\_max}, derived from \(F_M\). This is reported by
default because the smallest class proportion, \texttt{n\_min}, is of particular
interest. When the size of the smallest class is not much larger than
the number of parameters per class, proceed with caution: The model
might not be locally identified in that class
(\textbf{depaoliMixtureClassRecovery2013?}). Existing rules of thumb have focused
on smallest class size in terms of percentage of the total sample size.
This is not particularly meaningful, however, as the primary concern is
that small classes may comprise too few observations to accurately
estimate class parameters. Thus, a better rule of thumb might be to
ascertain that the smallest class has many more observations than the
average number of parameters per class. Another good practice is to
ensure that the smallest class has a unique and meaningful substantive
interpretation. If the smallest class does not comprise at least a few
observations per parameter, or is redundant with another larger class, a
simpler solution is preferable.

To assess classification accuracy, we turn to the third diagnostic
table: The average posterior probabilities by most likely class
membership, \(P_{p_{(M = m)}}\),
\texttt{class\_prob(res,\ type\ =\ "avg.mostlikely")}: for participants assigned to
one particular class (rows) according to \(M\), this table gives the mean
probability of being assigned to any class (columns). This table is
obtained by calculating the column means of \(P\) for the \(k\) subsets of
observations with most likely class of \(M = 1 \ldots k\). The diagonal
represents the average predictive probability of a the subjects related
to their most likely class, indicating the certainty of class assignment
for the subjects. Some guidelines have being suggested for this
diagonal, like for them to be higher than 0.7 (\textbf{masyn2013?}), but we
suggest researchers to be cautions about these guidelines as they cannot
be generalized and should not be use as strict cutoffs to defined ``good''
or ``bad'' classification.

The table of classification probabilities for the most likely latent
class \(M\) by latent class, \(P_{p_{(M = m | C = c)}}\) is similar to the
aforementioned table \(P_{p_{(M = m)}}\), except that it accounts for
classification uncertainty. If \(C\) is the true class of an observation,
which is imperfectly measured, and \(M\) is the most likely class, then
this table shows the probability of an observation being assigned to
class \(m\) in \(M\), given that its true class is \(c\) in \(C\), or
\(p(M=m|C=c)\). Values on the diagonal can thus be conceptualized as the
reliability with which each level of the categorical latent variable is
measured by \(M\), and off-diagonal elements can be conceptualized as
measurement error. The minimum and maximum of the diagonal of this table
are given by default as part of the output of \texttt{table\_fit()} for LCA
analyses (\texttt{prob\_min} and \texttt{prob\_max}). These probabilities are of
particular interest because they conveniently summarize classification
accuracy per class. Both probabilities should be high, as this means
that all classes are reliably classified. If the minimum probability is
low, there is at least one class with poor classification accuracy.

Finally, class separability can be summarized in a single statistic: the
entropy criterion {[}REF{]}.
In
physics, high entropy refers to maximum randomness or uncertainty, and
low entropy corresponds to a strict arrangement of the units of study.
Applied to LCA, high entropy means that classes are completely
indistinguishable; every individual has equal probability of being in
any class. Low entropy, by contrast, means that classes are fully
separated: every individual has a near-one probability of being in one
class only, and near-zero probability of being in any other class.
Crucially, in \texttt{tidySEM}, the entropy criterion is defined as 1-entropy
(a convention originating in Mplus). Its interpretation is reversed; 0
means the model classification is no better than chance, and 1 means
perfect classification. There are no rules of thumb for entropy, because
entropy is also sensitive to the sample size (entropy tends to be lower
in larger samples), number of latent classes (entropy tends to be lower
for higher number of classes), and number of indicators. Like other
classification diagnostics, entropy is not a model fit measure and
should not be used for model selection (Masyn, 2013). However, it
can be used to disqualify certain solutions if the goal is to identify
clearly distinct classes.

\hypertarget{labelling-classes}{%
\subsubsection{Labelling Classes}\label{labelling-classes}}

Class names should be chosen to accurately reflect theoretically
relevant characteristics of group membership. These class names should
aid the interpretation of the class solutions and serve as a shorthand
throughout the manuscript. For example, one study used a dual-trajectory
LCGA of cognitive and affective empathy development with gender as a
covariate. Although this model has dozens of parameters, the resulting
classes were labeled low, average, and high empathy because level
differences were the most distinguishing feature. However, overly
simplistic and general class names can be misleading, a phenomenon known
as the \emph{naming fallacy} (Weller et al., 2020). For example, it would be
misleading to call one group ``low empathy'' and another group ``medium
empathy'' if the empathy differences were non-significant, or if other
differences, like the effect of the covariate gender, were more
substantial than level differences. One way to avoid the naming fallacy
is comprehensively reporting model parameters, and explicating that
class names are just a shorthand.

When labeling classes, it is important to note that LCA models are
identified up to the class order.
This
means that the order of classes is arbitrary and may change if the
analysis is replicated. A simple solution to ensure replicable results
is setting a random seed before fitting the model, using \texttt{set.seed()}.
This will ensure that the classes stay in the same order upon
replication, but that may not be the preferred order for, e.g.,
interpreting the results. To address this problem, the function
\texttt{mx\_switch\_labels()} can be used. By default, the function orders
classes from largest to smallest class count - but classes can also be
ordered by values of a specific parameter (e.g., the mean intercept in a
latent class growth model). Note that label switching is a much bigger
problem when using Bayesian estimation, where it can occur during model
estimation. This is beyond the scope of the present paper.

\hypertarget{follow-up-analyses}{%
\subsubsection{Follow-up Analyses}\label{follow-up-analyses}}

LCA is rarely performed in isolation; often, researchers wish to relate
class membership to auxiliary variables, or even auxiliary models. It is
important to keep in mind that model-implied class membership is subject
to classification error. Treating class membership as an observed
variable in follow-up analyses disregards this error. The results will
be biased, except when class separation is very high - as indicated by
high (near 1) entropy and minimum posterior classification probability.
It is therefore important to account for classification error.

There are several different ways to do follow-up analyses in latent
class analysis (LCA) while accounting for classification error. At
present, the so-called BCH method is considered one of the best
approaches {[}after its creators, REF{]}. It has been implemented in
\texttt{tidySEM}. The advantage of the BCH method is that it not only allows
for the comparison of individual variables, but also entire auxiliary
models across latent classes.

\hypertarget{best-practices-in-reporting}{%
\subsection{Best Practices in Reporting}\label{best-practices-in-reporting}}

Among studies using LCA, reporting practices vary significantly
(Weller et al., 2020). Various authors have tried to improve and
standardize ways of reporting LCA (e.g. Masyn, 2013; Weller et al., 2020), but more work is needed.

\hypertarget{open-and-reproducible-science}{%
\subsubsection{Open and Reproducible Science}\label{open-and-reproducible-science}}

Journals and funding bodies increasingly require research to be open and
reproducible. For instance, many journals (and the APA) have adopted the
TOP guidelines, which recommend comprehensive citation of literature,
software, and data sources; sharing data and analysis code required to
exactly reproduce analyses, and pre-registering study plans prior to
data collection or analysis (\textbf{nosekTransparencyOpennessPromotion2016a?}).
A key advantage of using open source software for LCA is that, compared
to commercial software, it enables much greater adherence to open
science principles. Sharing code has limited utility if it can only be
evaluated in commercial software, as this restricts the ability to
reproduce analyses to license holders, and prevents auditing research
because the underlying source code is proprietary. The \texttt{tidySEM} package
enables performing LCA in free open source software, thus increasing the
utility of shared code.

Adopting FOSS makes it possible to conduct a fully reproducible LCA
analysis compliant with best practices in open science. The Workflow for
Open Reproducible Code in Science describes how to create a fully
reproducible research archive, and the \texttt{worcs} R-package automates many
of the steps involved (\textbf{vanlissaWORCSWorkflowOpen2021a?}).
This workflow is based on three principles: First, the paper (or
analysis section) is written as a \emph{dynamic document} that contains both
prose and analysis code. Results are automatically rendered as text,
tables, and plots. Second, the research archive is \emph{version controlled},
which means that the entire historical record of the project is
retained. This version controlled archive can be made public. Third,
\emph{dependency management} ensures that all software used is documented,
including its version number and provenance. Without dependency
management, code might not work (for example, because it is not clear
where certain functions come from), or its results may change when some
dependencies are updated. By combining \texttt{tidySEM} and \texttt{worcs}, it is
possible to conduct a fully reproducible LCA analysis compliant with
best practices in open science.

An additional argument for sharing data in the context of LCA is that
reproducing LCA analyses requires individual participant data, not just
sufficient statistics like many other methods (i.e., a correlation or
covariance matrix). If the original data cannot be shared, sharing
synthetic data at least enables reproducing and reusing analysis code.
The \texttt{worcs} package provides a generic, non-parametric method to create
a synthetic copy of data when data cannot be shared. It is also possible
to synthesize data directly from the LCA model using \texttt{mxGenerateData()},
which often provides a better basis for reproducing the LCA.

\hypertarget{preregistration}{%
\subsubsection{Preregistration}\label{preregistration}}

LCA analyses inherently involve some subjectivity, because of the
potential for disagreement between criteria for class enumeration and
the lack of objective fit indices. In our experience, reviewers (and
authors) often disagree about which criteria should be used.
Preregistering the class enumeration criteria as part of the analysis
plan reduces disagreement and eliminates concerns about cherry picking
fit measures after the results are known. It is particularly important
to pre-specify the criteria that will be used to determine the maximum
number of classes, because there is a risk that information criteria and
likelihood ratio tests will suggest a larger number of classes, even if
there are concerns about overfitting, classification accuracy, or model
interpretability.

The aforementioned Workflow for Open Reproducible Code in Science
facilitates creating a preregistration. More interestingly, it also
enables creating a \emph{Preregistration As Code} (PAC): A draft paper with
reproducible analysis section based on fake (synthetic or simulated)
data (\textbf{peikertReproducibleResearchTutorial2021?}). When the real data are
accessed or collected, the analysis is simply reproduced with this new
data file, and the results in the paper are automatically updated. PACs
are quite unambiguous; the selection of the final model can even be
automated by hard-coding the criteria into the analysis section. For an
example of PAC, see (\textbf{vanlissaComplementingPreregisteredConfirmatory2022?}).

\hypertarget{reporting-model-fit}{%
\subsubsection{Reporting Model Fit}\label{reporting-model-fit}}

The function \texttt{table\_fit()} reports an overview of fit indices,
customized for LCA. In reporting model fit, it is important, firstly, to
report the effective sample size, which should be the same for all
models in order to compare relative fit measures. Secondly, it should be
clearly reported, for example, in the strategy of analyses, which
parameters are estimated across and within classes. In case of doubt, a
full overview of the parameters is given by the \texttt{coef()} and
\texttt{table\_results()} functions. The number of parameters for each model
should be included in the fit table. Thirdly, it is important that
readers can assess the relative fit of different solutions. The most
fundamental fit measure is the \(-2LL\), as this is the quantity optimized
during estimation. From these statistics, information criteria can be
computed. By default, \texttt{table\_fit()} includes the AIC, BIC, and saBIC. We
suggest to report at least the BIC, as it is ubiquitous, grounded in
theory, and performs well in class enumeration.

Classification diagnostics should be reported, as they may determine
whether models are excluded from consideration. Of particular interest
is the diagonal of either the average posterior probabilities by most
likely class membership, or of the most likely class membership by true
class membership. The former indicates the average probability that
individuals belong to the class they were assigned to, and the latter
indicates the probability that individuals are assigned to a class if
they actually belong to that class. By default, \texttt{table\_fit()} reports
the latter as \texttt{prob\_min} and \texttt{prob\_max}. The entropy is also of interest
as an overall measure of class separation. Finally, the range of
proportions of cases assigned to each class is important; the minimum
proportion of cases per class can be used to diagnose classes that are
too small to be locally identified or practically relevant (see \textbf{depaoliMixtureClassRecovery2013?}). By default, \texttt{table\_fit()} reports the
minimum and maximum proportion of cases, based on most likely class
membership. Fifth, the LMR or BLRT can be reported if it is part of the
class enumeration strategy, using \texttt{lr\_lmr()} and \texttt{BLRT()}. By default,
\texttt{table\_fit()} reports the VLM for subsequent models in the table.

\hypertarget{reporting-results}{%
\subsubsection{Reporting results}\label{reporting-results}}

Reporting LCA results involves reporting the class proportions, obtained
via \texttt{class\_prob(res,\ "sum.posterior")}, and model parameters, obtained
via \texttt{table\_results(res)}. The class proportions are obtained by
standardizing the class weights: A \(k\)-class model has \(k-1\) free class
weights, and one weight constrained to 1 for model identification. The
free weights are included in the model results, but are redundant with
the aforementioned class proportions. Other model parameters may include
means, (co)variances, regression coefficients (rarely), and thresholds
for ordinal indicators. It is good practice to report the point estimate
and confidence interval of these parameters. Although it is conventional
to report p-values as well, note that the default p-values tests the
hypothesis that the parameter is equal to zero. This hypothesis may be
irrelevant in LCA when the focus is on parameter differences between
classes. Reporting standard errors allows readers to compute p-values
for other more informative hypotheses. Thresholds are more readily
interpretable when converted to conditional item probabilities, using
the function \texttt{table\_prob()}. These indicate the probability of an item
being endorsed given that the observation belongs to a particular latent
(Geiser, 2012).

\hypertarget{inference}{%
\subsubsection{Inference}\label{inference}}

As mentioned before, the results table included null-hypothesis tests
for all parameters by default. Whether these tests are informative
depends on the study goals and scale of the data. For example, if LCA
indicators are mean-centered, then the p-values of class-specific means
indicate whether the mean in that class differs significantly from the
sample average. It is also possible to test more informative hypotheses,
using the function \texttt{wald\_test()}. Equality-constrained informative
hypotheses are specified using the \texttt{bain} syntax, and evaluated on the
model object. This enables testing theoretical hypotheses - for example,
if specific values for class-specific means had been hypothesized. The
same function allows testing parameters for equality, either within or
across classes.

\hypertarget{visualization}{%
\subsubsection{Visualization}\label{visualization}}

In an informal review of published graphics for LCA, we found that most
visualizations focused on the estimated model parameters, displaying
class means as points connected by lines, with point shape and linetype
indicating class membership. Note that the lines introduce spurious
information. Furthermore, they omit uncertainty about class means and do
not illustrate how well the model describes the data.

The Grammar of Graphics addresses best practices in visualization, which
we summarize here with specific emphasis on the application to LCA. Most
importantly, good graphics start with the \emph{data} in mind. Visualizing
raw data allows one to assess class separability, and which aspects of
the data are captured well or poorly by the model. A second
consideration are which \emph{aesthetics} to use. In a color plot, class
membership can be indicated by color. In a black and white plot, it can
be indicated by linetype and point shape. A third consideration is which
geometric objects (geoms) to use. As mentioned before, common practice
is to plot parameter estimates as points, and connect these points with
lines. Points are suitable geoms to visualize point estimates and raw
data. For continuous indicators, additional relevant geoms include error
bars for standard errors and standard deviations, and density plots to
visualize the distribution of raw data (smoothed with a kernel density
function). In two dimensions, (co)variances can be visualized as
ellipses. Finally, when plotting multiple class solutions, it makes
sense to \emph{facet} a plot into multiple panels by class solution. The
plots in \texttt{tidySEM} are based on best practices as outlined in the
grammar of graphics, and are implemented in \texttt{ggplot2}. As such, these
plots can be further modified just like any other \texttt{ggplot2} object. One
remaining limitation is that these best practices have only been
implemented for LCA with continuous indicators, in \texttt{plot\_profiles()},
\texttt{plot\_density()} and \texttt{plot\_bivariate()}, and for LCGA in
\texttt{plot\_growth()}. Plots for LCA with ordinal indicators are still
rudimentary, more work remains to be done in this area.

\hypertarget{potential-pitfalls}{%
\subsection{Potential Pitfalls}\label{potential-pitfalls}}

Based on our experience in statistical consultations, there are a few
common pitfalls and misunderstandings about LCA. First, perhaps the most
common problem we see in consultations is treating ordinal data as
continuous. As explained previously, LCA for ordinal indicators should
be used for Likert-type scales and other data with few unique values. A
second common pitfall is omitting the 1-class solution from class
enumeration. If a one-class solution has better or similar fit to models
with a higher number of classes, LCA may not be appropriate. A one-class
solution should thus always be included as a benchmark. Third, although
it is widely understood that classification diagnostics (e.g., entropy,
posterior probability, class counts) are not class enumeration criteria,
they still have a role in class enumeration. Specifically, they can be
used to exclude solutions that do not meet predefined criteria from
consideration. Fourth, previous reporting standards called for reporting
of sample descriptive statistics (see \textbf{schreiberLatentClassAnalysis2017?}). While we agree that this is due
dilligence, sample descriptives have two limitations in the context of
LCA: they are not sufficient to check or reproduce the analysis, and
many use cases of LCA assume that the population is heterogeneous, which
limits the utility of whole-sample descriptives. These two limitations
can be addressed by data sharing and reporting descriptive statistics
for each latent class separately. Fifth, a recent publication mistakenly
stated that an assumption of mixture models is that observed indicators
are normally distributed (Spurk, Hirschi, Wang, Valero, \& Kauffeld, 2020). This is incorrect, as
evident from the shoe size example: When the number of classes is
greater than one, Gaussian mixture models assume that the distribution
of observed indicators is a \emph{mixture of multiple normal distributions}.
If the indicators were normally
distributed, there would be no classes to extract as the whole
population would belong to a single (homogeneous) class. Finally, the
same paper suggested that robust standard errors should be used when
indicators are not normally distributed (Spurk et al., 2020). This is
also incorrect; to our knowledge, there is no literature on the merits
of robust standard errors for LCA. Finally, we caution against step-wise
class enumeration, whereby additional models are estimated until the
selected class enumeration criteria no longer show improvement of model
fit (e.g., see Spurk et al., 2020). This practice is problematic,
because it is possible that an even higher number of classes would lead
to a (significant) improvement in fit. Consider this thought experiment:
there exist two groups of three relatively similar classes each (six in
total). Step-wise class enumeration shows improvement in model fit going
from a 1-class to 2-class model: the two classes capture the difference
between the two groups of three classes. Then, no noteworthy improvement
in model fit is found for adding more classes, until the 6-class
solution is reached. In the step-wise approach, the analysis would be
terminated after the 2-class model - even though the correct number of
classes is 6.

\hypertarget{smart-lca-checklist}{%
\subsection{SMART-LCA Checklist}\label{smart-lca-checklist}}

The best practices discussed in this paper are summarized in the
SMART-LCA Checklist: Standards for More Accuracy in Reporting of
different Types of Latent Class Analysis.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Report item descriptives

  \begin{itemize}
  \tightlist
  \item
    Are all items on similar scales?
  \item
    Is the measurement level correctly coded?
  \item
    Any other reasons why the observed (multivariate) distribution
    could not be described by the LCA (e.g., extreme skew)?
  \end{itemize}
\item
  Report missingness

  \begin{itemize}
  \tightlist
  \item
    Proportion of missingness per variable
  \item
    MAR test
  \item
    Missing data method for handling missingness (typically FIML)
  \end{itemize}
\item
  Report model specification

  \begin{itemize}
  \tightlist
  \item
    In exploratory LCA, compare all sensible alternative
    specifications
  \item
    In confirmatory LCA, justify alternative models on theoretical
    grounds
  \item
    Check that the observed (multivariate) distribution could
    theoretically be captured by the specified model (see \#1).
  \end{itemize}
\item
  Class enumeration

  \begin{itemize}
  \tightlist
  \item
    Justify the maximum number of classes \(K\)
  \item
    In exploratory LCA, estimate 1:\(K\) classes
  \item
    In confirmatory LCA, estimate the theoretical number of classes,
    a 1-class model, and optionally other numbers of classes for
    comparison
  \end{itemize}
\item
  Justify the (preregistered) criteria used for class enumeration,
  which can include:

  \begin{itemize}
  \tightlist
  \item
    Information criteria
  \item
    LR tests
  \item
    Theoretical interpretability of the solution
  \end{itemize}
\item
  Justify the (preregistered) criteria used to eliminate models from
  consideration, including:

  \begin{itemize}
  \tightlist
  \item
    Classification diagnostics
  \item
    Local identifiability (acceptable number of observations per
    parameter in each class)
  \item
    Model convergence
  \item
    Theoretical interpretability of the solution
  \end{itemize}
\item
  Report the following:

  \begin{itemize}
  \tightlist
  \item
    Fit of all models under consideration
  \item
    Classification diagnostics for all models under consideration
  \item
    All parameters of the chosen model (means, (co)variances,
    thresholds)
  \item
    Proportion of cases per class
  \end{itemize}
\item
  Provide an informative plot of the chosen solution

  \begin{itemize}
  \tightlist
  \item
    Show raw data
  \item
    Show parameters
  \item
    Show confidence bounds if possible
  \item
    Show multivariate distributions for multivariate models, if
    possible
  \end{itemize}
\item
  Account for classification inaccuracy in follow-up analyses

  \begin{itemize}
  \tightlist
  \item
    If entropy \textgreater{} .95 and posterior classification probability for
    each class \textgreater{} .95, this is less crucial
  \end{itemize}
\end{enumerate}

\hypertarget{conclusions}{%
\section{Conclusions}\label{conclusions}}

Write me

\newpage

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-akogul_comparison_2016}{}}%
Akogul, S., \& Erisoglu, M. (2016). A {Comparison} of {Information} {Criteria} in {Clustering} {Based} on {Mixture} of {Multivariate} {Normal} {Distributions}. \emph{Mathematical and Computational Applications}, \emph{21}(3), 34. \url{https://doi.org/10.3390/mca21030034}

\leavevmode\vadjust pre{\hypertarget{ref-alkema_maximizing_2022}{}}%
Alkema, L. M., Van Lissa, C. J., Kooi, M., \& Koelmans, A. A. (2022). Maximizing {Realism}: {Mapping} {Plastic} {Particles} at the {Ocean} {Surface} {Using} {Mixtures} of {Normal} {Distributions}. \emph{Environmental Science \& Technology}, \emph{56}(22), 15552--15562. \url{https://doi.org/10.1021/acs.est.2c03559}

\leavevmode\vadjust pre{\hypertarget{ref-baughman_mixture_2006}{}}%
Baughman, A. L., Bisgard, K. M., Lynn, F., \& Meade, B. D. (2006). Mixture model analysis for establishing a diagnostic cut-off point for pertussis antibody levels. \emph{Statistics in Medicine}, \emph{25}(17), 2994--3010. \url{https://doi.org/10.1002/sim.2442}

\leavevmode\vadjust pre{\hypertarget{ref-collins_latent_2009}{}}%
Collins, L. M., \& Lanza, S. T. (2009). \emph{Latent {Class} and {Latent} {Transition} {Analysis}: {With} {Applications} in the {Social}, {Behavioral}, and {Health} {Sciences}}. John Wiley \& Sons.

\leavevmode\vadjust pre{\hypertarget{ref-enders_applied_2022}{}}%
Enders, C. K. (2022). \emph{Applied {Missing} {Data} {Analysis}}. Guilford Publications.

\leavevmode\vadjust pre{\hypertarget{ref-figueiredo_unsupervised_2002}{}}%
Figueiredo, M. A. T., \& Jain, A. K. (2002). Unsupervised learning of finite mixture models. \emph{IEEE Transactions on Pattern Analysis and Machine Intelligence}, \emph{24}(3), 381--396. \url{https://doi.org/10.1109/34.990138}

\leavevmode\vadjust pre{\hypertarget{ref-geiser_data_2012}{}}%
Geiser, C. (2012). \emph{Data {Analysis} with {Mplus}}. Guilford Press.

\leavevmode\vadjust pre{\hypertarget{ref-geiser_is_2014}{}}%
Geiser, C., \& Wurpts, I. (2014). Is adding more indicators to a latent class analysis beneficial or detrimental? {Results} of a {Monte} {Carlo} study. \emph{Frontiers in Psychology: Quantitative Psychology and Measurement}, \emph{5}. https://doi.org/\url{https://doi.org/10.3389/fpsyg.2014.00920}

\leavevmode\vadjust pre{\hypertarget{ref-gromatsky_characteristics_2022}{}}%
Gromatsky, M., Edwards, E. R., Sullivan, S. R., Van Lissa, C. J., Lane, R., Spears, A. P., \ldots{} Goodman, M. (2022). Characteristics of suicide attempts associated with lethality and method: {A} latent class analysis of the {Military} {Suicide} {Research} {Consortium}. \emph{Journal of Psychiatric Research}, \emph{149}, 54--61.

\leavevmode\vadjust pre{\hypertarget{ref-hennig_handbook_2015}{}}%
Hennig, C., Meila, M., Murtagh, F., \& Rocci, R. (2015). \emph{Handbook of {Cluster} {Analysis}}. 28.

\leavevmode\vadjust pre{\hypertarget{ref-jamshidian_tests_2010}{}}%
Jamshidian, M., \& Jalal, S. (2010). Tests of {Homoscedasticity}, {Normality}, and {Missing} {Completely} at {Random} for {Incomplete} {Multivariate} {Data}. \emph{Psychometrika}, \emph{75}(4), 649--674. \url{https://doi.org/10.1007/s11336-010-9175-3}

\leavevmode\vadjust pre{\hypertarget{ref-jung_introduction_2008}{}}%
Jung, T., \& Wickrama, K. a. S. (2008). An {Introduction} to {Latent} {Class} {Growth} {Analysis} and {Growth} {Mixture} {Modeling}. \emph{Social and Personality Psychology Compass}, \emph{2}(1), 302--317. \url{https://doi.org/10.1111/j.1751-9004.2007.00054.x}

\leavevmode\vadjust pre{\hypertarget{ref-van_kollenburg_lazy_2018}{}}%
Kollenburg, G. H. van, Mulder, J., \& Vermunt, J. K. (2018). \emph{The {Lazy} {Bootstrap}. {A} {Fast} {Resampling} {Method} for {Evaluating} {Latent} {Class} {Model} {Fit}}. 23.

\leavevmode\vadjust pre{\hypertarget{ref-lee_comparison_2021}{}}%
Lee, T., \& Shi, D. (2021). A comparison of full information maximum likelihood and multiple imputation in structural equation modeling with missing data. \emph{Psychological Methods}, No Pagination Specified--No Pagination Specified. \url{https://doi.org/10.1037/met0000381}

\leavevmode\vadjust pre{\hypertarget{ref-little_test_1988}{}}%
Little, R. J. A. (1988). A {Test} of {Missing} {Completely} at {Random} for {Multivariate} {Data} with {Missing} {Values}. \emph{Journal of the American Statistical Association}, \emph{83}(404), pp. 1198--1202. \url{https://doi.org/10.2307/2290157}

\leavevmode\vadjust pre{\hypertarget{ref-macgregor_symptom_2021}{}}%
MacGregor, A. J., Dougherty, A. L., D'Souza, E. W., McCabe, C. T., Crouch, D. J., Zouris, J. M., \ldots{} Fraser, J. J. (2021). Symptom profiles following combat injury and long-term quality of life: A latent class analysis. \emph{Quality of Life Research}, \emph{30}(9), 2531--2540. \url{https://doi.org/10.1007/s11136-021-02836-y}

\leavevmode\vadjust pre{\hypertarget{ref-maene_perceived_2022}{}}%
Maene, C., D'hondt, F., Van Lissa, C. J., Thijs, J., \& Stevens, P. A. J. (2022). Perceived {Teacher} {Discrimination} and {Depressive} {Feelings} in {Adolescents}: {The} {Role} of {National}, {Regional}, and {Heritage} {Identities} in {Flemish} {Schools}. \emph{Journal of Youth and Adolescence}, \emph{51}(12), 2281--2293. \url{https://doi.org/10.1007/s10964-022-01665-7}

\leavevmode\vadjust pre{\hypertarget{ref-marcia_development_1966}{}}%
Marcia, J. E. (1966). Development and validation of ego-identity status. \emph{Journal of Personality and Social Psychology}, \emph{3}, 551--558. \url{https://doi.org/10.1037/h0023281}

\leavevmode\vadjust pre{\hypertarget{ref-masyn_latent_2013}{}}%
Masyn, K. E. (2013). Latent {Class} {Analysis} and {Finite} {Mixture} {Modeling}. In \emph{The {Oxford} {Handbook} of {Quantitative} {Methods}}: \emph{Vol.} \emph{2: Statistical Analysis} (p. 551). Oxford University Press.

\leavevmode\vadjust pre{\hypertarget{ref-mcelreath_statistical_2020}{}}%
McElreath, R. (2020). \emph{Statistical {Rethinking} {\textbar} {A} {Bayesian} {Course} with {Examples} in {R} and {STAN}} (2nd ed.). Chapman; Hall/CRC. Retrieved from \url{https://www.taylorfrancis.com/books/mono/10.1201/9780429029608/statistical-rethinking-richard-mcelreath}

\leavevmode\vadjust pre{\hypertarget{ref-neale_openmx_2016}{}}%
Neale, M. C., Hunter, M. D., Pritikin, J. N., Zahery, M., Brick, T. R., Kirkpatrick, R. M., \ldots{} Boker, S. M. (2016). {OpenMx} 2.0: {Extended} {Structural} {Equation} and {Statistical} {Modeling}. \emph{Psychometrika}, \emph{81}(2), 535--549. \url{https://doi.org/10.1007/s11336-014-9435-8}

\leavevmode\vadjust pre{\hypertarget{ref-norman_likert_2010}{}}%
Norman, G. (2010). Likert scales, levels of measurement and the {``laws''} of statistics. \emph{Advances in Health Sciences Education}, \emph{15}(5), 625--632. \url{https://doi.org/10.1007/s10459-010-9222-y}

\leavevmode\vadjust pre{\hypertarget{ref-nylund_deciding_2007}{}}%
Nylund, K. L., Asparouhov, T., \& Muthén, B. O. (2007). Deciding on the {Number} of {Classes} in {Latent} {Class} {Analysis} and {Growth} {Mixture} {Modeling}: {A} {Monte} {Carlo} {Simulation} {Study}. \emph{Structural Equation Modeling: A Multidisciplinary Journal}, \emph{14}(4), 535--569. \url{https://doi.org/10.1080/10705510701575396}

\leavevmode\vadjust pre{\hypertarget{ref-nylund-gibson_ten_2018}{}}%
Nylund-Gibson, K., \& Choi, A. Y. (2018). Ten frequently asked questions about latent class analysis. \emph{Translational Issues in Psychological Science}, \emph{4}(4), 440--461. \url{https://doi.org/10.1037/tps0000176}

\leavevmode\vadjust pre{\hypertarget{ref-pavlopoulos_measuring_2015}{}}%
Pavlopoulos, D., \& Vermunt, J. K. (2015). \emph{Measuring temporary employment. {Do} survey or register data tell the truth?} 37.

\leavevmode\vadjust pre{\hypertarget{ref-rosseel_lavaan_2012}{}}%
Rosseel, Y. (2012). Lavaan: {An} {R} {Package} for {Structural} {Equation} {Modeling}. \emph{Journal of Statistical Software}, \emph{48}, 1--36. \url{https://doi.org/10.18637/jss.v048.i02}

\leavevmode\vadjust pre{\hypertarget{ref-rubin_inference_1976}{}}%
Rubin, D. B. (1976). Inference and {Missing} {Data}. \emph{Biometrika}, \emph{63}(3), 581--592. \url{https://doi.org/10.2307/2335739}

\leavevmode\vadjust pre{\hypertarget{ref-scrucca_mclust_2016}{}}%
Scrucca, L., Fop, M., Murphy, T. B., \& Raftery, A. E. (2016). \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5096736}{Mclust 5: {Clustering}, {Classification} and {Density} {Estimation} {Using} {Gaussian} {Finite} {Mixture} {Models}}. \emph{The R Journal}, \emph{8}(1), 289--317.

\leavevmode\vadjust pre{\hypertarget{ref-spurk_latent_2020}{}}%
Spurk, D., Hirschi, A., Wang, M., Valero, D., \& Kauffeld, S. (2020). Latent profile analysis: {A} review and {``how to''} guide of its application within vocational behavior research. \emph{Journal of Vocational Behavior}, \emph{120}, 103445. \url{https://doi.org/10.1016/j.jvb.2020.103445}

\leavevmode\vadjust pre{\hypertarget{ref-tsaousis_measurement_2020}{}}%
Tsaousis, I., Sideridis, G. D., \& AlGhamdi, H. M. (2020). Measurement {Invariance} and {Differential} {Item} {Functioning} {Across} {Gender} {Within} a {Latent} {Class} {Analysis} {Framework}: {Evidence} {From} a {High}-{Stakes} {Test} for {University} {Admission} in {Saudi} {Arabia}. \emph{Frontiers in Psychology}, \emph{11}. Retrieved from \url{https://www.frontiersin.org/articles/10.3389/fpsyg.2020.00622}

\leavevmode\vadjust pre{\hypertarget{ref-van_de_schoot_grolts-checklist_2017}{}}%
Van De Schoot, R., Sijbrandij, M., Winter, S. D., Depaoli, S., \& Vermunt, J. K. (2017). The {GRoLTS}-{Checklist}: {Guidelines} for {Reporting} on {Latent} {Trajectory} {Studies}. \emph{Structural Equation Modeling: A Multidisciplinary Journal}, \emph{24}(3), 451--467. \url{https://doi.org/10.1080/10705511.2016.1247646}

\leavevmode\vadjust pre{\hypertarget{ref-van_lissa_divergence_2015}{}}%
Van Lissa, C. J., Hawk, S. T., Branje, S. J. T., Koot, H. M., Van Lier, P. A. C., \& Meeus, W. H. J. (2015). Divergence {Between} {Adolescent} and {Parental} {Perceptions} of {Conflict} in {Relationship} to {Adolescent} {Empathy} {Development}. \emph{Journal of Youth and Adolescence}, \emph{44}(1), 48--61. \url{https://doi.org/10.1007/s10964-014-0152-5}

\leavevmode\vadjust pre{\hypertarget{ref-vermunt_longitudinal_2010}{}}%
Vermunt, Jeroen K. (2010). Longitudinal {Research} {Using} {Mixture} {Models}. In K. Montfort, J. H. Oud, \& A. Satorra (Eds.), \emph{Longitudinal {Research} with {Latent} {Variables}} (pp. 119--152). Berlin/Heidelberg: Springer. \url{https://doi.org/10.1007/978-3-642-11760-2_4}

\leavevmode\vadjust pre{\hypertarget{ref-vermunt_latent_2004}{}}%
Vermunt, Jeroen K., Magidson, J., Lewis-Beck, M., Bryman, A., Liao, T. F., \& Department of Methodology and Statistics. (2004). Latent class analysis. In \emph{The {Sage} encyclopedia of social sciences research methods} (pp. 549--553). Sage. Retrieved from \url{https://research.tilburguniversity.edu/en/publications/0caedd00-27c1-42bd-bb4e-d1dcb0864956}

\leavevmode\vadjust pre{\hypertarget{ref-depmixS4}{}}%
Visser, I., \& Speekenbrink, M. (2010). {depmixS4}: An {R} package for hidden markov models. \emph{Journal of Statistical Software}, \emph{36}(7), 1--21. Retrieved from \url{https://www.jstatsoft.org/v36/i07/}

\leavevmode\vadjust pre{\hypertarget{ref-weller_latent_2020}{}}%
Weller, B. E., Bowen, N. K., \& Faubert, S. J. (2020). Latent {Class} {Analysis}: {A} {Guide} to {Best} {Practice}. \emph{Journal of Black Psychology}, \emph{46}(4), 287--311. \url{https://doi.org/10.1177/0095798420930932}

\end{CSLReferences}

\end{document}