### 2023-01-24 appendix_c.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man,floatsintext]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\usepackage{csquotes}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Appendix C: Latent Class Growth Analysis},
  pdflang={en-EN},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Appendix C: Latent Class Growth Analysis}
\author{\phantom{0}}
\date{}

\shorttitle{BEST PRACTICES LATENT CLASS ANALYSIS}

\affiliation{\phantom{0}}

\begin{document}
\maketitle

This vignette illustrated \texttt{tidySEM}'s ability to perform latent class growth analysis, or growth mixture modeling.
The simulated data used for this example are inspired by work in progress by Plas and colleagues,
on heterogeneity in depression trajectories among Dutch military personnel who were deployed to Afghanistan.
The original data were collected as part of the \emph{Prospection in Stress-related Military Research (PRISMO)} study,
which examined of psychological problems after deployment in more than 1,000 Dutch military personnel from 2005-2019.

First, we load all required packages:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(tidySEM)}
\FunctionTok{library}\NormalTok{(ggplot2)}
\FunctionTok{library}\NormalTok{(MASS)}
\end{Highlighting}
\end{Shaded}

\hypertarget{data-preprocessing}{%
\subsection{Data preprocessing}\label{data-preprocessing}}

We first examined the descriptive statistics for the sum score scales:

\begin{table}[tbp]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:unnamed-chunk-2}Item descriptives}

\begin{tabular}{llllllll}
\toprule
name & \multicolumn{1}{c}{mean} & \multicolumn{1}{c}{median} & \multicolumn{1}{c}{sd} & \multicolumn{1}{c}{min} & \multicolumn{1}{c}{max} & \multicolumn{1}{c}{skew\_2se} & \multicolumn{1}{c}{kurt\_2se}\\
\midrule
scl.1 & 20.20 & 20.00 & 2.36 & 17.00 & 38.00 & 14.52 & 40.39\\
scl.2 & 20.42 & 19.00 & 3.51 & 16.00 & 64.00 & 25.90 & 111.96\\
scl.3 & 20.49 & 20.00 & 3.41 & 17.00 & 59.00 & 26.41 & 107.36\\
scl.4 & 20.61 & 20.00 & 3.36 & 16.00 & 50.00 & 18.23 & 54.25\\
scl.5 & 20.93 & 20.00 & 4.06 & 16.00 & 64.00 & 24.54 & 93.40\\
scl.6 & 21.07 & 20.00 & 4.10 & 16.00 & 58.00 & 20.44 & 65.76\\
\bottomrule
\end{tabular}

\end{threeparttable}
\end{center}

\end{table}

Note that all variables were extremely right-skewed due to censoring at the lower end of the scale.

We can examine these distributions visually as well:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{df\_plot }\OtherTok{\textless{}{-}} \FunctionTok{reshape}\NormalTok{(df, }\AttributeTok{direction =} \StringTok{"long"}\NormalTok{, }\AttributeTok{varying =} \FunctionTok{names}\NormalTok{(df))}
\FunctionTok{ggplot}\NormalTok{(df\_plot, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ scl)) }\SpecialCharTok{+} \FunctionTok{geom\_density}\NormalTok{() }\SpecialCharTok{+} \FunctionTok{facet\_wrap}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{time) }\SpecialCharTok{+}
    \FunctionTok{theme\_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{figure}
\includegraphics[width=0.5\linewidth]{appendices/plot_dist} \caption{ }\label{fig:unnamed-chunk-5}
\end{figure}

As this type of skew can result in convergence problems in LCGA,
we compared several transformations to reduce skew:
The square and cube root, log, inverse, and Box-Cox transformations.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{df\_scores }\OtherTok{\textless{}{-}}\NormalTok{ df\_plot}
\CommentTok{\# Store original range of SCL}
\NormalTok{rng\_scl }\OtherTok{\textless{}{-}} \FunctionTok{range}\NormalTok{(df\_scores}\SpecialCharTok{$}\NormalTok{scl)}
\CommentTok{\# Log{-}transform}
\NormalTok{df\_scores}\SpecialCharTok{$}\NormalTok{log }\OtherTok{\textless{}{-}}\NormalTok{ scales}\SpecialCharTok{::}\FunctionTok{rescale}\NormalTok{(}\FunctionTok{log}\NormalTok{(df\_scores}\SpecialCharTok{$}\NormalTok{scl), }\AttributeTok{to =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}
    \DecValTok{1}\NormalTok{))}
\CommentTok{\# Square root transform}
\NormalTok{df\_scores}\SpecialCharTok{$}\NormalTok{sqrt }\OtherTok{\textless{}{-}}\NormalTok{ scales}\SpecialCharTok{::}\FunctionTok{rescale}\NormalTok{(}\FunctionTok{sqrt}\NormalTok{(df\_scores}\SpecialCharTok{$}\NormalTok{scl), }\AttributeTok{to =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}
    \DecValTok{1}\NormalTok{))}
\CommentTok{\# Cube root transform}
\NormalTok{df\_scores}\SpecialCharTok{$}\NormalTok{qrt }\OtherTok{\textless{}{-}}\NormalTok{ scales}\SpecialCharTok{::}\FunctionTok{rescale}\NormalTok{(df\_scores}\SpecialCharTok{$}\NormalTok{scl}\SpecialCharTok{\^{}}\FloatTok{0.33}\NormalTok{, }\AttributeTok{to =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}
    \DecValTok{1}\NormalTok{))}
\CommentTok{\# Reciprocal transform}
\NormalTok{df\_scores}\SpecialCharTok{$}\NormalTok{reciprocal }\OtherTok{\textless{}{-}}\NormalTok{ scales}\SpecialCharTok{::}\FunctionTok{rescale}\NormalTok{(}\DecValTok{1}\SpecialCharTok{/}\NormalTok{df\_scores}\SpecialCharTok{$}\NormalTok{scl, }\AttributeTok{to =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}
    \DecValTok{1}\NormalTok{))}
\CommentTok{\# Define function for Box{-}Cox transformation}
\NormalTok{bc }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(x, lambda) \{}
\NormalTok{    (((x}\SpecialCharTok{\^{}}\NormalTok{lambda) }\SpecialCharTok{{-}} \DecValTok{1}\NormalTok{)}\SpecialCharTok{/}\NormalTok{lambda)}
\NormalTok{\}}
\CommentTok{\# Inverse Box{-}Cox transformation}
\NormalTok{invbc }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(x, lambda) \{}
\NormalTok{    ((x }\SpecialCharTok{*}\NormalTok{ lambda) }\SpecialCharTok{+} \DecValTok{1}\NormalTok{)}\SpecialCharTok{\^{}}\NormalTok{(}\DecValTok{1}\SpecialCharTok{/}\NormalTok{lambda)}
\NormalTok{\}}
\CommentTok{\# Box{-}Cox transform}
\NormalTok{b }\OtherTok{\textless{}{-}}\NormalTok{ MASS}\SpecialCharTok{::}\FunctionTok{boxcox}\NormalTok{(}\FunctionTok{lm}\NormalTok{(df\_scores}\SpecialCharTok{$}\NormalTok{scl }\SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{), }\AttributeTok{plotit =} \ConstantTok{FALSE}\NormalTok{)}
\NormalTok{lambda }\OtherTok{\textless{}{-}}\NormalTok{ b}\SpecialCharTok{$}\NormalTok{x[}\FunctionTok{which.max}\NormalTok{(b}\SpecialCharTok{$}\NormalTok{y)]}
\NormalTok{df\_scores}\SpecialCharTok{$}\NormalTok{boxcox }\OtherTok{\textless{}{-}} \FunctionTok{bc}\NormalTok{(df\_scores}\SpecialCharTok{$}\NormalTok{scl, lambda)}
\CommentTok{\# Store range of Box{-}Cox transformed data}
\NormalTok{rng\_bc }\OtherTok{\textless{}{-}} \FunctionTok{range}\NormalTok{(df\_scores}\SpecialCharTok{$}\NormalTok{boxcox)}
\NormalTok{df\_scores}\SpecialCharTok{$}\NormalTok{boxcox }\OtherTok{\textless{}{-}}\NormalTok{ scales}\SpecialCharTok{::}\FunctionTok{rescale}\NormalTok{(df\_scores}\SpecialCharTok{$}\NormalTok{boxcox, }\AttributeTok{to =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}
    \DecValTok{1}\NormalTok{))}
\CommentTok{\# Rescale SCL}
\NormalTok{df\_scores}\SpecialCharTok{$}\NormalTok{scl }\OtherTok{\textless{}{-}}\NormalTok{ scales}\SpecialCharTok{::}\FunctionTok{rescale}\NormalTok{(df\_scores}\SpecialCharTok{$}\NormalTok{scl, }\AttributeTok{to =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

We can plot these transformations:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Make plot data}
\NormalTok{df\_plot }\OtherTok{\textless{}{-}} \FunctionTok{do.call}\NormalTok{(rbind, }\FunctionTok{lapply}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\StringTok{"scl"}\NormalTok{, }\StringTok{"log"}\NormalTok{, }\StringTok{"sqrt"}\NormalTok{, }\StringTok{"qrt"}\NormalTok{,}
    \StringTok{"boxcox"}\NormalTok{), }\ControlFlowTok{function}\NormalTok{(n) \{}
    \FunctionTok{data.frame}\NormalTok{(df\_scores[}\FunctionTok{c}\NormalTok{(}\StringTok{"time"}\NormalTok{, }\StringTok{"id"}\NormalTok{)], }\AttributeTok{Value =}\NormalTok{ df\_scores[[n]],}
        \AttributeTok{Transformation =}\NormalTok{ n)}
\NormalTok{\}))}
\CommentTok{\# Plot}
\FunctionTok{ggplot}\NormalTok{(df\_plot, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Value, }\AttributeTok{colour =}\NormalTok{ Transformation)) }\SpecialCharTok{+} \FunctionTok{geom\_density}\NormalTok{() }\SpecialCharTok{+}
    \FunctionTok{facet\_wrap}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{time) }\SpecialCharTok{+} \FunctionTok{scale\_y\_sqrt}\NormalTok{() }\SpecialCharTok{+} \FunctionTok{xlab}\NormalTok{(}\StringTok{"scl (rescaled to 0{-}1)"}\NormalTok{) }\SpecialCharTok{+}
    \FunctionTok{theme\_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{figure}
\includegraphics[width=0.5\linewidth]{appendices/plot_trans} \caption{ }\label{fig:unnamed-chunk-9}
\end{figure}

Evidently, the Box-Cox transformation reduced skew the most.
Consequently, we proceeded with the Box-Cox transformed scores for analysis.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{dat }\OtherTok{\textless{}{-}}\NormalTok{ df\_scores[, }\FunctionTok{c}\NormalTok{(}\StringTok{"id"}\NormalTok{, }\StringTok{"time"}\NormalTok{, }\StringTok{"boxcox"}\NormalTok{)]}
\NormalTok{dat }\OtherTok{\textless{}{-}} \FunctionTok{reshape}\NormalTok{(dat, }\AttributeTok{direction =} \StringTok{"wide"}\NormalTok{, }\AttributeTok{v.names =} \StringTok{"boxcox"}\NormalTok{, }\AttributeTok{timevar =} \StringTok{"time"}\NormalTok{,}
    \AttributeTok{idvar =} \StringTok{"id"}\NormalTok{)}
\FunctionTok{names}\NormalTok{(dat) }\OtherTok{\textless{}{-}} \FunctionTok{gsub}\NormalTok{(}\StringTok{"boxcox."}\NormalTok{, }\StringTok{"scl"}\NormalTok{, }\FunctionTok{names}\NormalTok{(dat))}
\end{Highlighting}
\end{Shaded}

\hypertarget{latent-class-growth-analysis}{%
\section{Latent Class Growth Analysis}\label{latent-class-growth-analysis}}

Next, we estimated a latent class growth model
for SCL.
The model included an overall intercept, centered at T1, \texttt{i}.
To model the potential effect of deployment on
depresion,
we also included a dummy variable that was zero before
deployment, and 1 after deployment, \texttt{step}.
Finally, to model potential change (or recovery) in depression post-deployment,
we included a linear slope from T2-T6, \texttt{s}.
All variances of growth parameters were fixed to zero due to the sparse nature of the data.
In this vignette,
we do not consider more than 5 classes,
because the analyses are computationally very intensive and the data were simulated from a 3-class model.

\textbf{NOTE: The time scales in this model are not correct; it currently assumes that all measurements are equidistant. Feel free to experiment with adjusting this.}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{set.seed}\NormalTok{(}\DecValTok{27796}\NormalTok{)}
\NormalTok{dat[[}\StringTok{"id"}\NormalTok{]] }\OtherTok{\textless{}{-}} \ConstantTok{NULL}
\NormalTok{res\_step }\OtherTok{\textless{}{-}} \FunctionTok{mx\_growth\_mixture}\NormalTok{(}\AttributeTok{model =} \StringTok{"}
\StringTok{  i =\textasciitilde{} 1*scl1 + 1*scl2 + 1*scl3 +1*scl4 +1*scl5 +1*scl6}
\StringTok{  step =\textasciitilde{} 0*scl1 + 1*scl2 + 1*scl3 +1*scl4 +1*scl5 +1*scl6}
\StringTok{  s =\textasciitilde{} 0*scl1 + 0*scl2 + 1*scl3 +2*scl4 +3*scl5 +4*scl6}
\StringTok{  scl1 \textasciitilde{}\textasciitilde{} vscl1*scl1}
\StringTok{  scl2 \textasciitilde{}\textasciitilde{} vscl2*scl2}
\StringTok{  scl3 \textasciitilde{}\textasciitilde{} vscl3*scl3}
\StringTok{  scl4 \textasciitilde{}\textasciitilde{} vscl4*scl4}
\StringTok{  scl5 \textasciitilde{}\textasciitilde{} vscl5*scl5}
\StringTok{  scl6 \textasciitilde{}\textasciitilde{} vscl6*scl6}
\StringTok{  i \textasciitilde{}\textasciitilde{} 0*i}
\StringTok{  step \textasciitilde{}\textasciitilde{} 0*step}
\StringTok{  s \textasciitilde{}\textasciitilde{} 0*s}
\StringTok{  i \textasciitilde{}\textasciitilde{} 0*s}
\StringTok{  i \textasciitilde{}\textasciitilde{} 0*step}
\StringTok{  s \textasciitilde{}\textasciitilde{} 0*step"}\NormalTok{,}
    \AttributeTok{classes =} \DecValTok{1}\SpecialCharTok{:}\DecValTok{5}\NormalTok{, }\AttributeTok{data =}\NormalTok{ dat)}
\CommentTok{\# Additional iterations because of convergence problems for}
\CommentTok{\# model 1:}
\NormalTok{res\_step[[}\DecValTok{1}\NormalTok{]] }\OtherTok{\textless{}{-}} \FunctionTok{mxTryHardWideSearch}\NormalTok{(res\_step[[}\DecValTok{1}\NormalTok{]], }\AttributeTok{extraTries =} \DecValTok{50}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Note that the first model showed convergence problems, throwing the error:
\emph{The model does not satisfy the
first-order optimality conditions to
the required accuracy, and no improved
point for the merit function could be
found during the final linesearch.}
To address this problem, we performed
additional iterations to\\
find a better solution, using \texttt{OpenMx::mxTryHardWideSearch()}.
This also illustrates that \texttt{tidySEM} mixture models inherit from \texttt{OpenMx}'s \texttt{MxModel},
and thus, different \texttt{OpenMx} functions can be used to act on models specified via \texttt{tidySEM}.

The fifth model also evidenced convergence problems, but this (as we will see) is because the solution is overfitted.

\hypertarget{class-enumeration}{%
\subsection{Class enumeration}\label{class-enumeration}}

To determine the correct number of classes, we considered the following criteria:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  We do not consider classes with, on average, fewer than 5 participants per parameter in a class due to potential local underidentification
\item
  Lower values for information criteria (AIC, BIC, saBIC) indicate better fit
\item
  Significant Lo-Mendell-Rubin LRT test indicates better fit for \(k\) vs \(k-1\) classes
\item
  We do not consider solutions with entropy \textless{} .90 because poor class separability compromises interpretability of the results
\item
  We do not consider solutions with minimum posterior classification probability \textless{} .90 because poor class separability compromises interpretability of the results
\end{enumerate}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Get fit table fit}
\NormalTok{tab\_fit }\OtherTok{\textless{}{-}} \FunctionTok{table\_fit}\NormalTok{(res\_step)}
\CommentTok{\# Select columns}
\NormalTok{tab\_fit }\OtherTok{\textless{}{-}}\NormalTok{ tab\_fit[, }\FunctionTok{c}\NormalTok{(}\StringTok{"Name"}\NormalTok{, }\StringTok{"Classes"}\NormalTok{, }\StringTok{"LL"}\NormalTok{, }\StringTok{"Parameters"}\NormalTok{,}
    \StringTok{"BIC"}\NormalTok{, }\StringTok{"Entropy"}\NormalTok{, }\StringTok{"prob\_min"}\NormalTok{, }\StringTok{"n\_min"}\NormalTok{, }\StringTok{"warning"}\NormalTok{, }\StringTok{"lmr\_p"}\NormalTok{)]}
\NormalTok{tab\_fit}
\end{Highlighting}
\end{Shaded}

\begin{table}[tbp]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:tabfit}Fit of LCGA models}

\begin{tabular}{llllllllll}
\toprule
Name & \multicolumn{1}{c}{Classes} & \multicolumn{1}{c}{LL} & \multicolumn{1}{c}{p} & \multicolumn{1}{c}{BIC} & \multicolumn{1}{c}{Ent.} & \multicolumn{1}{c}{p\_min} & \multicolumn{1}{c}{n\_min} & \multicolumn{1}{c}{warn} & \multicolumn{1}{c}{lmr\_p}\\
\midrule
1 & 1.00 & 2,592.32 & 9 & -5,122.67 & 1.00 & 1.00 & 1.00 & NA & NA\\
2 & 2.00 & 3,797.78 & 13 & -7,506.04 & 0.92 & 0.97 & 0.31 & NA & 0.00\\
3 & 3.00 & 4,264.43 & 17 & -8,411.80 & 0.95 & 0.96 & 0.10 & NA & 0.00\\
4 & 4.00 & 4,264.45 & 21 & -8,384.30 & 0.72 & 0.00 & 0.00 & NA & 1.00\\
5 & 5.00 & 4,264.45 & 25 & -8,356.75 & 0.55 & 0.00 & 0.00 & TRUE & 1.00\\
\bottomrule
\end{tabular}

\end{threeparttable}
\end{center}

\end{table}

According to the Table, increasing the number of classes keeps increasing model fit according to all ICs except the BIC, which increased after 3 classes.

The first two LMR tests are significant,
indicating that a 2- and 3-class solution were a significant improvement over a 1- and 2-class solution, respectively.
However, solutions with \textgreater3 classes had entropy and minimum posterior classification probability below the pre-specified thresholds.
Models with \textgreater3 solutions also had fewer than five observations per parameter.
This suggests that the preferred model should be selected from 1-3 classes.

\hypertarget{scree-plot}{%
\subsubsection{Scree plot}\label{scree-plot}}

A scree plot indicates that
the largest decrease in ICs occurs from 1-2 classes,
and the inflection point for all ICs is at 3 classes.
Moreover, the BIC increased after 3 classes.
A three-class solution thus appears to be the most parsimonious
solution with good fit.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot}\NormalTok{(tab\_fit, }\AttributeTok{statistics =} \FunctionTok{c}\NormalTok{(}\StringTok{"AIC"}\NormalTok{, }\StringTok{"BIC"}\NormalTok{, }\StringTok{"saBIC"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{figure}
\includegraphics[width=0.5\linewidth]{appendices/lcga_plot_fit} \caption{ }\label{fig:unnamed-chunk-15}
\end{figure}

Based on the aforementioned criteria,
we selected a 3-class model for further analyses.
First, to prevent label switching,
we re-order these classes by the value of the intercept \texttt{i}.
Then, we report the estimated parameters.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{res\_final }\OtherTok{\textless{}{-}} \FunctionTok{mx\_switch\_labels}\NormalTok{(res\_step[[}\DecValTok{3}\NormalTok{]], }\AttributeTok{param =} \StringTok{"M[1,7]"}\NormalTok{,}
    \AttributeTok{decreasing =} \ConstantTok{FALSE}\NormalTok{)}
\NormalTok{tab\_res }\OtherTok{\textless{}{-}} \FunctionTok{table\_results}\NormalTok{(res\_final, }\AttributeTok{columns =} \ConstantTok{NULL}\NormalTok{)}
\CommentTok{\# Select rows and columns}
\NormalTok{tab\_res }\OtherTok{\textless{}{-}}\NormalTok{ tab\_res[tab\_res}\SpecialCharTok{$}\NormalTok{Category }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\StringTok{"Means"}\NormalTok{, }\StringTok{"Variances"}\NormalTok{),}
    \FunctionTok{c}\NormalTok{(}\StringTok{"Category"}\NormalTok{, }\StringTok{"lhs"}\NormalTok{, }\StringTok{"est"}\NormalTok{, }\StringTok{"se"}\NormalTok{, }\StringTok{"pval"}\NormalTok{, }\StringTok{"confint"}\NormalTok{, }\StringTok{"name"}\NormalTok{)]}
\NormalTok{tab\_res}
\end{Highlighting}
\end{Shaded}

\begin{center}
\begin{ThreePartTable}

\begin{longtable}{llllllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
\caption{\label{tab:tabres}Results from 3-class LCGA model}\\
\toprule
 & \multicolumn{1}{c}{Category} & \multicolumn{1}{c}{lhs} & \multicolumn{1}{c}{est} & \multicolumn{1}{c}{se} & \multicolumn{1}{c}{pval} & \multicolumn{1}{c}{confint} & \multicolumn{1}{c}{name}\\
\midrule
\endfirsthead
\caption*{\normalfont{Table \ref{tab:tabres} continued}}\\
\toprule
 & \multicolumn{1}{c}{Category} & \multicolumn{1}{c}{lhs} & \multicolumn{1}{c}{est} & \multicolumn{1}{c}{se} & \multicolumn{1}{c}{pval} & \multicolumn{1}{c}{confint} & \multicolumn{1}{c}{name}\\
\midrule
\endhead
16 & Means & i & 0.35 & 0.00 & 0.00 & {}[0.35, 0.36] & class1.M[1,7]\\
17 & Means & step & -0.03 & 0.01 & 0.00 & {}[-0.04, -0.02] & class1.M[1,8]\\
18 & Means & s & 0.01 & 0.00 & 0.00 & {}[0.00, 0.01] & class1.M[1,9]\\
19 & Variances & scl1 & 0.01 & 0.00 & 0.00 & {}[0.01, 0.01] & class1.S[1,1]\\
20 & Variances & scl2 & 0.01 & 0.00 & 0.00 & {}[0.01, 0.01] & class1.S[2,2]\\
21 & Variances & scl3 & 0.01 & 0.00 & 0.00 & {}[0.01, 0.01] & class1.S[3,3]\\
22 & Variances & scl4 & 0.01 & 0.00 & 0.00 & {}[0.01, 0.01] & class1.S[4,4]\\
23 & Variances & scl5 & 0.01 & 0.00 & 0.00 & {}[0.01, 0.01] & class1.S[5,5]\\
24 & Variances & scl6 & 0.01 & 0.00 & 0.00 & {}[0.01, 0.02] & class1.S[6,6]\\
40 & Means & i & 0.46 & 0.01 & 0.00 & {}[0.44, 0.47] & class2.M[1,7]\\
41 & Means & step & 0.03 & 0.01 & 0.00 & {}[0.01, 0.04] & class2.M[1,8]\\
42 & Means & s & 0.01 & 0.00 & 0.00 & {}[0.01, 0.02] & class2.M[1,9]\\
43 & Variances & scl1 & 0.01 & 0.00 & 0.00 & {}[0.01, 0.01] & class2.S[1,1]\\
44 & Variances & scl2 & 0.01 & 0.00 & 0.00 & {}[0.01, 0.01] & class2.S[2,2]\\
45 & Variances & scl3 & 0.01 & 0.00 & 0.00 & {}[0.01, 0.01] & class2.S[3,3]\\
46 & Variances & scl4 & 0.01 & 0.00 & 0.00 & {}[0.01, 0.01] & class2.S[4,4]\\
47 & Variances & scl5 & 0.01 & 0.00 & 0.00 & {}[0.01, 0.01] & class2.S[5,5]\\
48 & Variances & scl6 & 0.01 & 0.00 & 0.00 & {}[0.01, 0.02] & class2.S[6,6]\\
64 & Means & i & 0.63 & 0.01 & 0.00 & {}[0.61, 0.65] & class3.M[1,7]\\
65 & Means & step & 0.07 & 0.01 & 0.00 & {}[0.05, 0.10] & class3.M[1,8]\\
66 & Means & s & 0.01 & 0.00 & 0.13 & {}[-0.00, 0.01] & class3.M[1,9]\\
67 & Variances & scl1 & 0.01 & 0.00 & 0.00 & {}[0.01, 0.01] & class3.S[1,1]\\
68 & Variances & scl2 & 0.01 & 0.00 & 0.00 & {}[0.01, 0.01] & class3.S[2,2]\\
69 & Variances & scl3 & 0.01 & 0.00 & 0.00 & {}[0.01, 0.01] & class3.S[3,3]\\
70 & Variances & scl4 & 0.01 & 0.00 & 0.00 & {}[0.01, 0.01] & class3.S[4,4]\\
71 & Variances & scl5 & 0.01 & 0.00 & 0.00 & {}[0.01, 0.01] & class3.S[5,5]\\
72 & Variances & scl6 & 0.01 & 0.00 & 0.00 & {}[0.01, 0.02] & class3.S[6,6]\\
\bottomrule
\end{longtable}

\end{ThreePartTable}
\end{center}

As evident from these results,
Class 1 started at a relatively lower level of depressive symptoms,
experienced a decrease after deployment,
followed by increase over time.
Class 2 started at a moderate level of depressive symptoms,
experienced an increase after deployment,
followed by significant increase over time from T2-T6.
Class 3 started at a relatively higher level,
experienced an increase after deployment, followed by stability.

\hypertarget{wald-tests}{%
\subsection{Wald tests}\label{wald-tests}}

To test whether parameters are significantly different between classes,
we can use Wald tests.
Wald tests can be specified for all parameters in the model,
using the hypothesis syntax from the \texttt{bain} package for informative hypothesis testing.

To identify the names of parameters in the model,
we can use the \texttt{name} column of the results table above.
Alternatively, to see all parameters in the model, run:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{names}\NormalTok{(}\FunctionTok{coef}\NormalTok{(res\_final))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
#>  [1] "mix3.weights[1,2]" "mix3.weights[1,3]" "vscl1"            
#>  [4] "vscl2"             "vscl3"             "vscl4"            
#>  [7] "vscl5"             "vscl6"             "class1.M[1,7]"    
#> [10] "class1.M[1,8]"     "class1.M[1,9]"     "class2.M[1,7]"    
#> [13] "class2.M[1,8]"     "class2.M[1,9]"     "class3.M[1,7]"    
#> [16] "class3.M[1,8]"     "class3.M[1,9]"
\end{verbatim}

Next, specify equality constrained hypotheses.
For example, a hypothesis that states that the mean intercept is equal across groups is specified as follows:

\texttt{"class1.M{[}1,7{]}\ =\ class2.M{[}1,7{]}\ \&\ class1.M{[}1,7{]}\ =\ class3.M{[}1,7{]}}

It is also possible to consider comparisons between two classes at a time.
When conducting many significance tests,
consider correcting for multiple comparisons however.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{wald\_tests }\OtherTok{\textless{}{-}} \FunctionTok{wald\_test}\NormalTok{(res\_final, }\StringTok{"}
\StringTok{                   class1.M[1,7] = class2.M[1,7]\&}
\StringTok{                   class1.M[1,7] = class3.M[1,7];}
\StringTok{                   class1.M[1,8] = class2.M[1,8]\&}
\StringTok{                   class1.M[1,8] = class3.M[1,8];}
\StringTok{                   class1.M[1,9] = class2.M[1,9]\&}
\StringTok{                   class1.M[1,9] = class3.M[1,9]"}\NormalTok{)}
\CommentTok{\# Rename the hypothesis}
\NormalTok{wald\_tests}\SpecialCharTok{$}\NormalTok{Hypothesis }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\StringTok{"Mean i"}\NormalTok{, }\StringTok{"Mean step"}\NormalTok{, }\StringTok{"Mean slope"}\NormalTok{)}
\NormalTok{papaja}\SpecialCharTok{::}\FunctionTok{apa\_table}\NormalTok{(wald\_tests, }\AttributeTok{digits =} \DecValTok{2}\NormalTok{, }\AttributeTok{caption =} \StringTok{"Wald tests"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}[tbp]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:waldtests}Wald tests}

\begin{tabular}{llll}
\toprule
Hypothesis & \multicolumn{1}{c}{df} & \multicolumn{1}{c}{chisq} & \multicolumn{1}{c}{p}\\
\midrule
Mean i & 2 & 628.38 & 0.00\\
Mean step & 2 & 65.32 & 0.00\\
Mean slope & 2 & 13.03 & 0.00\\
\bottomrule
\end{tabular}

\end{threeparttable}
\end{center}

\end{table}

All Wald tests are significant, indicating that there are significant differences between the intercepts, step function, and slopes of the three classes.

\hypertarget{trajectory-plot}{%
\subsection{Trajectory plot}\label{trajectory-plot}}

Finally, we can plot the growth trajectories.
This can help interpret the results better,
as well as the residual heterogeneity around class trajectories.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{p }\OtherTok{\textless{}{-}} \FunctionTok{plot\_growth}\NormalTok{(res\_step[[}\DecValTok{3}\NormalTok{]], }\AttributeTok{rawdata =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{alpha\_range =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}
    \FloatTok{0.05}\NormalTok{))}
\CommentTok{\# Add Y{-}axis breaks in original scale}
\NormalTok{brks }\OtherTok{\textless{}{-}} \FunctionTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{, }\AttributeTok{length.out =} \DecValTok{5}\NormalTok{)}
\NormalTok{labs }\OtherTok{\textless{}{-}} \FunctionTok{round}\NormalTok{(}\FunctionTok{invbc}\NormalTok{(scales}\SpecialCharTok{::}\FunctionTok{rescale}\NormalTok{(brks, }\AttributeTok{from =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{), }\AttributeTok{to =}\NormalTok{ rng\_bc),}
\NormalTok{    lambda))}
\NormalTok{p }\OtherTok{\textless{}{-}}\NormalTok{ p }\SpecialCharTok{+} \FunctionTok{scale\_y\_continuous}\NormalTok{(}\AttributeTok{breaks =} \FunctionTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{, }\AttributeTok{length.out =} \DecValTok{5}\NormalTok{),}
    \AttributeTok{labels =}\NormalTok{ labs) }\SpecialCharTok{+} \FunctionTok{ylab}\NormalTok{(}\StringTok{"SCL (rescaled from Box{-}Cox)"}\NormalTok{)}
\NormalTok{p}
\end{Highlighting}
\end{Shaded}

\begin{figure}
\includegraphics[width=0.5\linewidth]{appendices/plot_traj} \caption{ }\label{fig:unnamed-chunk-22}
\end{figure}

Note that the observed individual trajectories show very high variability within classes.

\end{document}